### 2023-11-12 ms.qmd
Intro & Methods

---
docx: 
reference-doc: "./bib/tmpl.docx"
bibliography: "./bib/refs.bib"
execute:
  echo: false
  error: false
  cache: false
  warning: false
crossref:  
  fig-title: Figure    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Tab.   # (default is "Table")
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup

pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest)

source(here("R", "func.R"))
```

## Introduction

Hypothesis PRESENTATION:

Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity. In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals, which are more sensitive to stressors. As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter. At the same time I expect these individuals to have increased phisiological costs (mass lost) when submitted to a 'chronic' stressor.

## Methods

### Subjects

We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.

### Subjects treatments, collection, and housing

The skinks used in our experiments came from a breeding colony established in the lab since 2019. The colony is compound by a total total of 270 adults of *L. delicata* and 180 adults of *L.* guichenoti that are housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. These enclosures are provided with non-stick matting, shelter, and several small water dishes. Lizards are given water daily, and fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius; but to ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.

Eggs were collected between mid-October 2022 to the end of February 2023. We placed in the commonal enclosure small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside so females had a place to lay the eggs. Three times a week, we checked these boxes for the presence of eggs. Once a clutch was found, we measured eggs' length and width with a digital caliper to the nearest 0.1 mm and mass with a (OHAUS, Model spx123) digital scale ± 0.001g error. To test the effect of prenatal CORT and temperature, we manipulated CORT levels in eggs and then incubated them at one of two temperature regimes (‘Cold’ – 23ºC ± 3ºC or ‘Hot’ – 30ºC ± 3ºC) in a 2x2 factorial design. CORT was increased by topical aplication of a solution of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at concentrations 2 standard deviations above the mean natural hormone concentration (non-published data); while a 'Control' group received an equal volume of the vehicle. Then, eggs were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite) covered with cling wrap to retain moisture, and were left in LATWIT 2X5D-R1160 incubators at two different temperatures ('Cold' or 'Hot') until hatchling.

INSERT Fig 1

The incubators were checked three times a week for hatchlings. When hatchligns were found, we recorded Snout-vent length (SVL) and tail length (TL) with a rule to the nearest mm while mass was measured with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then, hatchlings were enclosed in individual tubs (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).

### Experimental arena and design
The experiments were done:
The sample size was:
The experimental arena was:
The experimental design was:
The lizards used in this experiment were subjected before to a personality test and two types of learning tasks. These experiments are not considered stresful, and, as such, we do not expect them to affect the results of the present study. However, we left three weeks between those experiments and the present one to control for possible effects of the previous tests.
After 3 months, we measured again SVL, TL, and mass of the juveniles and then they were moved to the experimental arena (see below).
The day before starting the experiment, lizards were measured again, and then moved to a new enclosure the same size and conditions before, but in the room where the CCTV system is held () experimental arena (Fig. 1c) for acclimatation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other. These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below). The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum.* We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The rooms temperature was set to between 22-24 Celsius.

### CORT measurement and validation

### Statistical analyses

## Results

```{r, dataload}
#| label: dataload
#| echo: false
#| warning: false

# Load data
	data  <-  read.csv("./data/.csv")
```

```{r, dataclean}
#| label: dataclean
#| echo: false
#| warning: false

# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
data_associative <- data %>%
  group_by(lizard_id) %>%
    filter(sum(is.na(FC_associative)) <= 15) %>%
    filter(trial_associative <= 35) %>%
  ungroup()  %>% 
    mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
    mutate(group = factor(group, levels = c("R_B", "B_R"))) %>%
  data.frame()

# Standarize data by trial (i.e. make the first trial where each individual participated their trial 1)

data_associative <- data_associative %>%
  group_by(lizard_id) %>%
  mutate(
    first_non_na = min(which(!is.na(FC_associative))),
    Associative_Trial = ifelse(!is.na(first_non_na),trial_associative - first_non_na + 1, trial_associative))%>%
    filter(Associative_Trial >= 1) %>%
   ungroup() %>% data.frame()

write.csv(data_associative, file= "./output/Checking/data_associative.csv")

# Split data by species
deli_associative <- data_associative %>% filter (species == "delicata")
guich_associative <- data_associative %>% filter (species == "guichenoti")

```

```{r, models1}
#| label: models1
#| echo: false
#| warning: false

# Generalized Mixed Models delicata
GMM_deli <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = binomial(link = "logit"))
summary(GMM_deli)

emmeans(GMM_deli, pairwise ~ temp*Associative_Trial)
emmeans(GMM_deli, pairwise ~ cort*Associative_Trial)

# Generalized Mixed Models guichenoti
GMM_guich <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = binomial(link = "logit"))
summary(GMM_guich)

emmeans(GMM_guich, pairwise ~ temp*Associative_Trial)
emmeans(GMM_guich, pairwise ~ cort*Associative_Trial)
emmeans(GMM_guich, pairwise ~ temp*cort*Associative_Trial)

```

```{r, models2}
#| label: models2
#| echo: false
#| warning: false

# Bayesian model delicata
  deli_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(deli_mod)
  plot(deli_mod)

# Bayesian model guichenoti
  guich_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(deli_mod)
# R2
bayes_R2(mod)

# Extract posteriors
posterior <- posterior_samples(mod, pars = "^b")
plyr::ldply(lapply(posterior, mean))
plyr::ldply(lapply(posterior, function(x) quantile(x, c(0.025, 0.975))))
temp23 <- c(posterior[,"b_Intercept"], (posterior[,"b_Intercept"]+ posterior[,"b_trtB_23"]))
temp28 <- c((posterior[,"b_Intercept"] + posterior[,"b_trtB_28"]), posterior[,"b_trtB_23"])

pmcmc(temp23 - temp28)
dharma(GMM_deli2)

```

```{r, plots}
#| label: plots
#| echo: false
#| warning: false

# Plot delicata
plot_deli <- ggplot(deli_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
  facet_grid(as.factor(deli_associative$group)) + 
  labs(y = "Colour Association Task", x = "Trial")  
print(plot_deli)

# Plot guichenoti
plot_guich <- ggplot(guich_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
  facet_grid(as.factor(guich_associative$group)) + 
  labs(y = "Colour Association Task", x = "Trial") 
print(plot_guich)
```

We calculated $R^2$ as follows

$$
R^2 = \frac{SS_{reg}}{SS_{tot}}
$$ {#eq-r2}

@eq-r2

```{r, fig1}
#| label: fig-fig1
#| fig-cap: "Activity budget of the focal group of chacma baboons (Papio ursinus) at the Cape Peninsula, South Africa, during the study period. The data are presented as the mean percentage of time spent in each activity category per hour (± SE)."

data %>% ggplot(aes(x = Lizard_id))  + geom_histogram()

```

We have a total of `r length(unique(data$Lizard_id))`

```{r,tb1}
#| label: tbl-tb1
#| tbl-cap: "Summary of the data"

tab <- data %>% 
  group_by(Lizard_id) %>% 
  summarise(
    n = n()
  ) 
flextable(tab[2:3,])
```

# References

### 2023-11-12 ms.qmd
Changes analyses

@@
 
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS)
 
@@
 # Load data
-	data  <-  read.csv("./data/.csv")
+	data  <-  read.csv("./data/Stressor.csv")
 ```
 
-```{r, dataclean}
-#| label: dataclean
+```{r, dataexplore}
+#| label: dataexplore
 #| echo: false
@@
 
-# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
-data_associative <- data %>%
-  group_by(lizard_id) %>%
-    filter(sum(is.na(FC_associative)) <= 15) %>%
-    filter(trial_associative <= 35) %>%
-  ungroup()  %>% 
+# Divide Treatment in CORT and Temperature
+data <- data %>%
       mutate(temp = gsub("[AB]_", "", trt),
           cort = gsub("_[2][38]", "", trt))  %>% 
-    mutate(group = factor(group, levels = c("R_B", "B_R"))) %>%
   data.frame()
 
-# Standarize data by trial (i.e. make the first trial where each individual participated their trial 1)
-
-data_associative <- data_associative %>%
-  group_by(lizard_id) %>%
-  mutate(
-    first_non_na = min(which(!is.na(FC_associative))),
-    Associative_Trial = ifelse(!is.na(first_non_na),trial_associative - first_non_na + 1, trial_associative))%>%
-    filter(Associative_Trial >= 1) %>%
-   ungroup() %>% data.frame()
-
-write.csv(data_associative, file= "./output/Checking/data_associative.csv")
+# Transform mass into mg
+data$mass <- data$mass*1000
 
 # Split data by species
-deli_associative <- data_associative %>% filter (species == "delicata")
-guich_associative <- data_associative %>% filter (species == "guichenoti")
+deli_stressor <- data %>% filter (species == "delicata")
+guich_stressor <- data %>% filter (species == "guichenoti")
+
+# Explore distributions of variables
+
+# A) L. deli
+    #a.1) Mass 
+plot_mass <- ggplot(deli_stressor, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_mass)
+
+    #a.2) Latency to move
+plot_freeze <- ggplot(deli_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_freeze)
+
+# I will both distributions with saphiro.test and lat_to_move with a qqplot
+mass_normal_test<-by(deli_stressor$mass, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test((x)))
+print(mass_normal_test)
+
+freeze_normal_test<-by(deli_stressor$lat_to_move, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test(log(x+1)))
+print(freeze_normal_test)
+# Once checked the distributions, I will transform lat_to_move with log(x+1) to make it normal
+
+# B) L. guich
+    #b.1) Mass 
+plot_mass <- ggplot(guich_stressor, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_mass)
+
+    #a.2) Latency to move
+plot_freeze <- ggplot(guich_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_freeze)
+
+# I will both distributions with saphiro.test
+mass_normal_test<-by(guich_stressor$mass, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test((x)))
+print(mass_normal_test)
+
+freeze_normal_test<-by(guich_stressor$lat_to_move, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test(log(x+1)))
+print(freeze_normal_test)
+# Once checked the distributions, I will transform lat_to_move with log(x+1) to make it normal
 

### 2023-11-16 .vdoc.r
Updates

@@ -1,78 +1,78 @@
----
-docx: 
-reference-doc: "./bib/tmpl.docx"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS)
 
 source(here("R", "func.R"))
-```
-
-## Introduction
-
-Hypothesis PRESENTATION:
-
-Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity. In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals, which are more sensitive to stressors. As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter. At the same time I expect these individuals to have increased phisiological costs (mass lost) when submitted to a 'chronic' stressor.
-
-
-## Methods
-
-### Subjects
-
-We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.
-
-### Subjects treatments, collection, and housing
-
[…]

### 2023-11-16 ms.qmd
Updates

--- 2023-11-12 ms.qmd
@@
 
-# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
-data_associative <- data %>%
-  group_by(lizard_id) %>%
-    filter(sum(is.na(FC_associative)) <= 15) %>%
-    filter(trial_associative <= 35) %>%
-  ungroup()  %>% 
+# Divide Treatment in CORT and Temperature
+data <- data %>%
       mutate(temp = gsub("[AB]_", "", trt),
           cort = gsub("_[2][38]", "", trt))  %>% 
-    mutate(group = factor(group, levels = c("R_B", "B_R"))) %>%
   data.frame()
 
-# Standarize data by trial (i.e. make the first trial where each individual participated their trial 1)
+# MASS
+# Create new database to analyse increases in mass (i.e. having mass after and before in different columns)
+data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested), names_from = Day, values_from = mass)
+data_2 <- data_2 %>% mutate(delta_mass=After-Before) 
+data_2 <- data_2 %>% mutate(rescaled_delta_mass = delta_mass - min(delta_mass) + 1)
+write.csv(data_2, "./data/Stressor_2.csv")
+# Transform mass into mg in new database
+data_2$mass <- data$mass*1000
 
-data_associative <- data_associative %>%
-  group_by(lizard_id) %>%
-  mutate(
-    first_non_na = min(which(!is.na(FC_associative))),
-    Associative_Trial = ifelse(!is.na(first_non_na),trial_associative - first_non_na + 1, trial_associative))%>%
-    filter(Associative_Trial >= 1) %>%
-   ungroup() %>% data.frame()
+# Estimate mass increase
 
-write.csv(data_associative, file= "./output/Checking/data_associative.csv")
+# Split data by species in new database
+deli_stressor_2 <- data_2 %>% filter (species == "delicata")
+guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")
 
-# Split data by species
-deli_associative <- data_associative %>% filter (species == "delicata")
-guich_associative <- data_associative %>% filter (species == "guichenoti")
+# Explore distribution of mass increase
+  # A) L.deli
+plot_mass <- ggplot(deli_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_mass)
+
+mass_normal_test<-by(deli_stressor_2$mass, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test((x)))
+print(mass_normal_test)
+
+  #B) L.guich
+plot_mass <- ggplot(guich_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_mass)
+
+mass_normal_test<-by(guich_stressor$mass, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test((x)))
+print(mass_normal_test)
+
+# BEHAVIOUR
+# Split data by species in original database
+deli_stressor <- data %>% filter (species == "delicata")
+guich_stressor <- data %>% filter (species == "guichenoti")
+# Explore distribution of latency to move/time frozen
+  # A) L. deli
+plot_freeze <- ggplot(deli_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_freeze)
+
+freeze_normal_test<-by(deli_stressor$lat_to_move, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test(log(x+1)))
+print(freeze_normal_test)
+
+ #B) L.guich
+plot_freeze <- ggplot(guich_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
+  theme_minimal() +
+  facet_grid(cort*temp~Day, scales = "free_y")
+print(plot_freeze)
+
+freeze_normal_test<-by(guich_stressor$lat_to_move, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test(log(x+1)))
+print(freeze_normal_test)
 

### 2023-11-16 .vdoc.r

@@ -1,265 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS)
-
-source(here("R", "func.R"))
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: dataload
-#| echo: false
-#| warning: false
- 
-# Load data
-	data  <-  read.csv("./data/Stressor.csv")
-#
-#
-#
-#| label: dataexplore
-#| echo: false
-#| warning: false
-
-# Divide Treatment in CORT and Temperature
-data <- data %>%
-      mutate(temp = gsub("[AB]_", "", trt),
-          cort = gsub("_[2][38]", "", trt))  %>% 
-  data.frame()
-
-# MASS
-# Create new database to analyse increases in mass (i.e. having mass after and before in different columns)
-data_2 <- pivot_wider(data, id_cols=lizard_id, names_from = Day, values_from = mass)
-write.csv(data_2, "./data/Stressor_2.csv")
-# Transform mass into mg in new database
-data_2$mass <- data$mass*1000
-
-# Estimate mass increase
-
-# Split data by species in new database
-deli_stressor_2 <- data_2 %>% filter (species == "delicata")
-guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")
-
-# Explore distribution of mass increase
-  # A) L.deli
-plot_mass <- ggplot(deli_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
-  theme_minimal() +
-  facet_grid(cort*temp~Day, scales = "free_y")
-print(plot_mass)
-
[…]

### 2023-11-16 .vdoc.r
Updates

@@ -1,78 +1,78 @@
----
-docx: 
-reference-doc: "./bib/tmpl.docx"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr)
 
 source(here("R", "func.R"))
-```
-
-## Introduction
-
-Hypothesis PRESENTATION:
-
-Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity. In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals, which are more sensitive to stressors. As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter. At the same time I expect these individuals to have increased phisiological costs (mass lost) when submitted to a 'chronic' stressor.
-
-
-## Methods
-
-### Subjects
-
-We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.
-
-### Subjects treatments, collection, and housing
-
[…]

### 2023-11-17 ms.qmd
Updates and graphs

@@
 
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
 
@@
 
-# Divide Treatment in CORT and Temperature
+# Divide Treatment in CORT and Temperature Transform mass into mg
 data <- data %>%
@@
           cort = gsub("_[2][38]", "", trt))  %>% 
+      mutate(mass=mass*1000) %>%
 data.frame()
+data$cort <- factor(data$cort, 
+        levels = c("A" ,"B"),
+        labels = c("Control", "CORT"))
+data$temp <- factor(data$temp,
+        levels = c("23", "28"),
+        labels = c("Cold", "Hot"))
+data$Day <- factor(data$Day,
+        levels = c("Before", "After"))
 
 # MASS
-# Create new database to analyse increases in mass (i.e. having mass after and before in different columns)
-data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested), names_from = Day, values_from = mass)
-data_2 <- data_2 %>% mutate(delta_mass=After-Before) 
-data_2 <- data_2 %>% mutate(rescaled_delta_mass = delta_mass - min(delta_mass) + 1)
-write.csv(data_2, "./data/Stressor_2.csv")
-# Transform mass into mg in new database
-data_2$mass <- data$mass*1000
+# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
+data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
+ names_from = Day, values_from = mass)
 
-# Estimate mass increase
+# Estimate mass increase and rescale it to avoid negative values
+data_2 <- data_2 %>% 
+ mutate(delta_mass=After-Before) %>%
+ mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
+  NA, 
+  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
+write.csv(data_2, file = "./data/Stressor_2.csv")
 
-# Split data by species in new database
+# Split new database by species
 deli_stressor_2 <- data_2 %>% filter (species == "delicata")
@@
   # A) L.deli
-plot_mass <- ggplot(deli_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+plot_mass <- ggplot(deli_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
+  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
+  labs(x = "Mass", y = "Density", fill = "temp ~ cort") +
+  theme_minimal() +
+  facet_grid(as.factor(deli_stressor_2$cort)~as.factor(deli_stressor_2$temp), scales = "free_y")
+print(plot_mass)
+
+mass_normal_test<-by(deli_stressor_2$rescaled_delta_mass, interaction (deli_stressor_2$cort, deli_stressor_2$temp), function(x) shapiro.test((x)))
+print(mass_normal_test)
+
+  #B) L.guich
+plot_mass <- ggplot(guich_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(guich_stressor_2$temp),as.factor(guich_stressor_2$cort), sep="_"))) +
   geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
@@
   theme_minimal() +
-  facet_grid(cort*temp~Day, scales = "free_y")
+  facet_grid(guich_stressor_2$cort~guich_stressor_2$temp, scales = "free_y")
 print(plot_mass)
 
-mass_normal_test<-by(deli_stressor_2$mass, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test((x)))
-print(mass_normal_test)
-
-  #B) L.guich
-plot_mass <- ggplot(guich_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
-  theme_minimal() +
-  facet_grid(cort*temp~Day, scales = "free_y")
-print(plot_mass)
-
-mass_normal_test<-by(guich_stressor$mass, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test((x)))
+mass_normal_test<-by(guich_stressor_2$rescaled_delta_mass, interaction (guich_stressor_2$cort, guich_stressor_2$temp), function(x) shapiro.test((x)))
 print(mass_normal_test)
@@
 
-```{r, models1}
-#| label: models1
+```{r, models}
+#| label: models
 #| echo: false
@@
 
-# Generalized Mixed Models delicata
-GMM_deli <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = binomial(link = "logit"))
-summary(GMM_deli)
+# Delta mass L. delicata
+  deli_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  plot(deli_mod)
+  summary(deli_mod)
+# Delta mass L. guichenoti
+  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  summary(guich_mod)
+bayes_R2(guich_mod)
 
-emmeans(GMM_deli, pairwise ~ temp*Associative_Trial)
-emmeans(GMM_deli, pairwise ~ cort*Associative_Trial)
+      # Seems that only food ingestion is significant in both species
 
-# Generalized Mixed Models guichenoti
-GMM_guich <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = binomial(link = "logit"))
-summary(GMM_guich)
-
-emmeans(GMM_guich, pairwise ~ temp*Associative_Trial)
-emmeans(GMM_guich, pairwise ~ cort*Associative_Trial)
-emmeans(GMM_guich, pairwise ~ temp*cort*Associative_Trial)
-
-```
-
-```{r, models2}
-#| label: models2
-#| echo: false
-#| warning: false
-
-# Bayesian model delicata
-  deli_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+# Frozen behaviour L. delicata
+  deli_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = deli_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
   summary(deli_mod)
   plot(deli_mod)
+bayes_R2(deli_mod)
 
-# Bayesian model guichenoti
-  guich_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(deli_mod)
-# R2
-bayes_R2(mod)
+# Frozen behaviour L. delicata
+ guich_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = guich_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  summary(guich_mod)
+  plot(guich_mod)
+bayes_R2(guich_mod)
 
@@
 
-```{r, plots}
-#| label: plots
+```{r, plotsMass}
+#| label: plotsMass
 #| echo: false
@@
 
-# Plot delicata
-plot_deli <- ggplot(deli_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
-  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
-  facet_grid(as.factor(deli_associative$group)) + 
-  labs(y = "Colour Association Task", x = "Trial")  
+# Plot delta Mass L. delicata
+plot_deli <- ggplot(deli_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
+  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
+  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
+    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
+  ) +
+  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
+  theme_classic() +
+  theme(
+    axis.title = element_text(size = 18, family = "sans"),  
+    axis.text = element_text(size = 10, family = "sans"),  
+    legend.title = element_text(size = 14, family = "sans"), 
+    legend.text = element_text(size = 12, family = "sans")  
+  )
 print(plot_deli)
+ggsave("./output/figures/fig_deli_Mass.png", plot=plot_deli, width = 18, height = 9, units = "cm", dpi = 3000)
+# Plot delta Mass L. guichenoti
+plot_guich <- ggplot(guich_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
+  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
+  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
+    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
+  ) +
+  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
+  theme_classic() +
+  theme(
+    axis.title = element_text(size = 18, family = "sans"),  
+    axis.text = element_text(size = 10, family = "sans"),  
+    legend.title = element_text(size = 14, family = "sans"), 
+    legend.text = element_text(size = 12, family = "sans")  
+  )
+print(plot_guich)
+ggsave("./output/figures/fig_guich_Mass.png", plot=plot_guich, width = 18, height = 9, units = "cm", dpi = 3000)
+```
 
-# Plot guichenoti
-plot_guich <- ggplot(guich_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
-  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
-  facet_grid(as.factor(guich_associative$group)) + 
-  labs(y = "Colour Association Task", x = "Trial") 
-print(plot_guich)
+ ```{r, plotsFrozenbehaviour}
+#| label: plotsFrozenbehaviour
+#| echo: false
+#| warning: false
+
+# Plot time frozen L. delicata
+plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
+  geom_boxplot(outlier.shape = NA) +
+  ylim(0, 450) +
+  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
+    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
+  ) +
+  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
+  theme_classic() +
+  theme(
+    axis.title = element_text(size = 18, family = "sans"),  
+    axis.text = element_text(size = 12, family = "sans"),  
+    legend.title = element_text(size = 14, family = "sans"), 
+    legend.text = element_text(size = 12, family = "sans")  
+  )
+print(plot_deli_BEH)
+ggsave("./output/figures/fig_deli_BEH.png", plot=plot_deli_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
+
+# Plot time frozen L. guichenoti
+plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
+  geom_boxplot(outlier.shape = NA) +
+  ylim(0, 450) +
+  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
+    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
+  ) +
+  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
+  theme_classic() +
+  theme(
+    axis.title = element_text(size = 18, family = "sans"),  
+    axis.text = element_text(size = 12, family = "sans"),  
+    legend.title = element_text(size = 14, family = "sans"), 
+    legend.text = element_text(size = 12, family = "sans")  
+  )
+print(plot_guich_BEH)
+ggsave("./output/figures/fig_guich_BEH.png", plot=plot_guich_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
+
 ```

### 2023-11-17 .vdoc.r

@@ -1,271 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr)
-
-source(here("R", "func.R"))
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: dataload
-#| echo: false
-#| warning: false
- 
-# Load data
-	data  <-  read.csv("./data/Stressor.csv")
-#
-#
-#
-#| label: dataexplore
-#| echo: false
-#| warning: false
-
-# Divide Treatment in CORT and Temperature Transform mass into mg
-data <- data %>%
-      mutate(temp = gsub("[AB]_", "", trt),
-          cort = gsub("_[2][38]", "", trt))  %>% 
-      mutate(mass=mass*1000) %>%
-      mutate(cort = factor(data$cort,
-        levels = c("A" ,"B")
-        labels = c("Control", "CORT"))) %>%
-      mutate (temp = factor(data$temp,
-        levels = c("23", "28"),
-        labels = c("Cold", "Hot"))) %>%
-  data.frame()
-     
-# MASS
-# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
-data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
- names_from = Day, values_from = mass)
-
-# Stimate mass increase and rescale it to avoid negative values
-data_2 <- data_2 %>% 
- mutate(delta_mass=After-Before) %>%
- mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
-  NA, 
-  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
-write.csv(data_2, file = "./data/Stressor_2.csv")
-
-# Split new database by species
-deli_stressor_2 <- data_2 %>% filter (species == "delicata")
-guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")
-
-# Explore distribution of mass increase
-  # A) L.deli
[…]

### 2023-11-17 .vdoc.r
Updates images

@@ -1,78 +1,78 @@
----
-docx: 
-reference-doc: "./bib/tmpl.docx"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
 
 source(here("R", "func.R"))
-```
-
-## Introduction
-
-Hypothesis PRESENTATION:
-
-Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity. In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals, which are more sensitive to stressors. As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter. At the same time I expect these individuals to have increased phisiological costs (mass lost) when submitted to a 'chronic' stressor.
-
-
-## Methods
-
-### Subjects
-
-We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.
-
-### Subjects treatments, collection, and housing
-
[…]

### 2025-05-24 .vdoc.5cea6b43-ef86-4f29-aced-9b01630b6493.r
updates

@@ -1,0 +1,225 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
+
+source(here("R", "func.R"))
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: models
+#| echo: false
+#| warning: false
+
+# Delta mass L. delicata
+  deli_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  plot(deli_mod)
+  summary(deli_mod)
+# Delta mass L. guichenoti
+  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  summary(guich_mod)
+bayes_R2(guich_mod)
+
+      # Seems that only food ingestion is significant in both species
+
+
+# Frozen behaviour L. delicata
+  deli_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = deli_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  summary(deli_mod)
+  plot(deli_mod)
+bayes_R2(deli_mod)
+
+# Frozen behaviour L. delicata
+ guich_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = guich_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+  summary(guich_mod)
+  plot(guich_mod)
+bayes_R2(guich_mod)
+
+
+# Extract posteriors
[…]

### 2025-05-24 R/func.r

@@ -1,12 +1,318 @@
+#################### 
+####################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom,
+tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest)
+#
+####################
+####################
+# Extract sample size per group
+#' @title sample
+#' @description Estract sample size per group
+#' @param df Database
+#' @param corti To select cort treatment ("CORT"/"Control")
+#' @param therm To select temp treatment ("Cold"/"Hot")
+sample <- function(df, corti, therm){
+  sample_size <- df %>%
+    filter(cort == corti, temp == therm) %>%
+    distinct(lizard_id) %>%
+    summarise(total_count = n()) %>%
+    pull(total_count)
+  return(sample_size)
+}
+####################
+####################
+# Function to generate four Q-Q plots for one variable
+#' @title qq_plots_single 
+#' @description Generate four Q-Q plots for one variable
+#' @param df Database
+#' @param vle Variable chosen
+#' @param label Label assigned to each variable
+#' @return fig_qq
+qq_plots_single <- function(df, vle, label) {
+  # Remove NAs
+  plot_df <- data.frame(value = na.omit(df[[vle]])) 
 
+    # Normal Q-Q Plot
+    p1 <- ggplot(plot_df, aes(sample = value)) +
+      stat_qq() + 
+      stat_qq_line(color = "red") +
+      ggtitle(paste(label, "- Normal Q-Q Plot")) +
+      theme_classic()
+
+    # Log-Normal Q-Q Plot
+    p2 <- ggplot(plot_df, aes(sample = log(value +1))) +
+      stat_qq() + 
+      stat_qq_line(color = "red") +
+      ggtitle(paste(label, "- Log-Normal Q-Q Plot")) +
+      theme_classic()
+
+    # Exponential Q-Q Plot
+    p3 <- ggplot(plot_df, aes(sample = value)) +
+      stat_qq(distribution = qexp, dparams = list(rate = 1/mean(plot_df$value))) +
+      stat_qq_line(distribution = qexp, dparams = list(rate = 1/mean(plot_df$value)), color = "red") +
+      ggtitle(paste(label, "- Exponential Q-Q Plot")) +
+      theme_classic()
+
+    # Arrange plots in a grid
[…]

### 2025-05-24 ms.qmd
updates

@@
 #| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+```
 
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
+```{r, cleandata}
+#| label: cleandata
+# Obtain the main df using "./R/1_data_process.R"
+source(here("R", "data_process.R"))
+```
 
+```{r, sampleSize}
+#| label: sampleSize
+# List with the sample sizes from the main database.
 source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = mass_clean, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+```
+
+```{r, countclutches}
+#| label: countclutches
+# Count the number of clutches per species
+#
+clutches <- mass_clean %>% 
+  distinct(clutch) %>% 
+  nrow()
 ```
@@
 
-### CORT measurement and validation
-
 ### Statistical analyses
@@
 ## Results
-
-```{r, dataload}
-#| label: dataload
-#| echo: false
-#| warning: false
-
-# Load data
-	data  <-  read.csv("./data/Stressor.csv")
-```
-
-```{r, dataexplore}
-#| label: dataexplore
-#| echo: false
-#| warning: false
-
-# Divide Treatment in CORT and Temperature Transform mass into mg
-data <- data %>%
-      mutate(temp = gsub("[AB]_", "", trt),
-          cort = gsub("_[2][38]", "", trt))  %>% 
-      mutate(mass=mass*1000) %>%
-data.frame()
-data$cort <- factor(data$cort, 
-        levels = c("A" ,"B"),
-        labels = c("Control", "CORT"))
-data$temp <- factor(data$temp,
-        levels = c("23", "28"),
-        labels = c("Cold", "Hot"))
-data$Day <- factor(data$Day,
-        levels = c("Before", "After"))
-
-# MASS
-# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
-data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
- names_from = Day, values_from = mass)
-
-# Estimate mass increase and rescale it to avoid negative values
-data_2 <- data_2 %>% 
- mutate(delta_mass=After-Before) %>%
- mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
-  NA, 
-  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
-write.csv(data_2, file = "./data/Stressor_2.csv")
-
-# Split new database by species
-deli_stressor_2 <- data_2 %>% filter (species == "delicata")
-guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")
-
-# Explore distribution of mass increase
-  # A) L.deli
-plot_mass <- ggplot(deli_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "temp ~ cort") +
-  theme_minimal() +
-  facet_grid(as.factor(deli_stressor_2$cort)~as.factor(deli_stressor_2$temp), scales = "free_y")
-print(plot_mass)
-
-mass_normal_test<-by(deli_stressor_2$rescaled_delta_mass, interaction (deli_stressor_2$cort, deli_stressor_2$temp), function(x) shapiro.test((x)))
-print(mass_normal_test)
-
-  #B) L.guich
-plot_mass <- ggplot(guich_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(guich_stressor_2$temp),as.factor(guich_stressor_2$cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
-  theme_minimal() +
-  facet_grid(guich_stressor_2$cort~guich_stressor_2$temp, scales = "free_y")
-print(plot_mass)
-
-mass_normal_test<-by(guich_stressor_2$rescaled_delta_mass, interaction (guich_stressor_2$cort, guich_stressor_2$temp), function(x) shapiro.test((x)))
-print(mass_normal_test)
-
-# BEHAVIOUR
-# Split data by species in original database
-deli_stressor <- data %>% filter (species == "delicata")
-guich_stressor <- data %>% filter (species == "guichenoti")
-# Explore distribution of latency to move/time frozen
-  # A) L. deli
-plot_freeze <- ggplot(deli_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
-  theme_minimal() +
-  facet_grid(cort*temp~Day, scales = "free_y")
-print(plot_freeze)
-
-freeze_normal_test<-by(deli_stressor$lat_to_move, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test(log(x+1)))
-print(freeze_normal_test)
-
- #B) L.guich
-plot_freeze <- ggplot(guich_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
-  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
-  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
-  theme_minimal() +
-  facet_grid(cort*temp~Day, scales = "free_y")
-print(plot_freeze)
-
-freeze_normal_test<-by(guich_stressor$lat_to_move, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test(log(x+1)))
-print(freeze_normal_test)
-
-```
-
-```{r, models}
+```{r, models_mass}
 #| label: models
@@
 # Plot time frozen L. delicata
-plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
+plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
   geom_boxplot(outlier.shape = NA) +
-  ylim(0, 450) +
+  ylim(0, 10) +
   scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
@@
   ) +
-  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
+  labs(y = "Time frozen (log-transformed)", x = "Day", fill="Treatment") +
   theme_classic() +
@@
 # Plot time frozen L. guichenoti
-plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
+plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
   geom_boxplot(outlier.shape = NA) +
-  ylim(0, 450) +
+  ylim(0, 8) +
   scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),

### 2025-05-24 R/1_data_process.R

@@ -1,21 +1,0 @@
-####################################
-# 1_data_process
-####################################
-
-## Load packages
-	intstall.packages("pacman")
-	pacman::p_load(tidyverse, readxl)
-
-## Load data	
-	              activ_files <- list.files("./data/Activity/")
-	       data_activity_list <- lapply(activ_files, function(x) read_xlsx(paste0("./data/Activity/", x))[-c(1:3),])
-	names(data_activity_list) <- activ_files
-
-## Merge
-	data_activity <- dplyr::bind_rows(data_activity_list, .id = "id")
-
-
-
-# At the end of processing you write the file
-	write_csv(data_activity, "./output/data/data_activity.csv")
-

### 2025-05-24 R/data_process.R

@@ -1,0 +1,40 @@
+####################################
+# Data_process
+####################################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2,
+               lme4, zoo, lmerTest, broom, tidybayes, forcats)
+#
+source(here("R", "func.R"))
+# A) Cleaning the general_data.csv for all individuals
+clean_ind <- read.csv(here("data", "general_data.csv")) %>%
+  mutate(lizard_id = as.character(lizard_id)) %>%
+  mutate(temp = gsub("[AB]_", "", trt),
+         cort = gsub("_[2][38]", "", trt))  %>%
+  mutate(temp = factor(temp,
+                       levels = c("23", "28"),
+                       labels = c("23" = "Cold", "28" = "Hot"))) %>%
+  mutate(cort = factor(cort,
+                       levels = c("A", "B"),
+                       labels = c("A" = "Control", "B" = "CORT"))) %>%
+  dplyr::select(-notes, -trt) %>%
+data.frame()
+
+####################################
+# B) Merging mass.csv with clean individual data (clean_ind)
+mass <- read.csv(here("data", "mass.csv")) %>%
+  mutate(lizard_id = as.character(lizard_id))%>%
+data.frame() #Read in mass data and transform lizard_id to character
+mass_clean <- merge(clean_ind, mass, by = "lizard_id", all = TRUE) %>%
+  mutate(delta_mass_rescaled = (delta_mass - min(delta_mass, na.rm = TRUE))) %>%
+data.frame() #Merge mass data with clean_ind data and rescale delta_mass
+
+write.csv(mass_clean, here("./output/data_clean/mass_clean.csv")) 
+
+####################################
+# C) Merging behaviour.csv with clean individual data (clean_ind)
+behav <- read.csv(here("data", "behaviour.csv"))%>%
+  mutate(lizard_id = as.character(lizard_id))%>%
+data.frame() #Read in behaviour data and transform lizard_id to character
+behav_clean <- merge(clean_ind, behav, by = "lizard_id", all = TRUE) 
+
+write.csv(behav_clean, here("./output/data_clean/behav_clean.csv"))

### 2025-05-24 R/exploratory.R

@@ -1,0 +1,29 @@
+####################################
+# Exploratory analyses
+####################################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes,
+ggh4x, cowplot, fitdistrplus, MASS, goftest)
+#
+source(here("R", "func.R"))
+
+####################################
+# A) Exploring distribution of the main variables for mass_clean 
+data_expl <- mass_clean %>%
+       mutate(treatment = paste(cort, temp, sep = "-"))
+# Histogram
+bin_width <- (max(data_expl$delta_mass_rescaled, na.rm = TRUE) - min(data_expl$delta_mass_rescaled, na.rm = TRUE)) / sqrt(length(data_expl$delta_mass_rescaled))
+hist_plot_mass <- ggplot(data_expl, aes(x = delta_mass_rescaled)) +
+  geom_histogram(binwidth = bin_width, fill = "#062d00", color = "black", alpha = 0.5) +
+  theme_classic() +
+  labs(x = "delta_mass", y = "Counts") +
+  theme(axis.title = element_text(size = 12, family = "Times"),
+        axis.text = element_text(size = 10, family = "Times"))
+ggsave(here("./output/figures/exploratory/hist_mass.png"), hist_plot_mass, width = 20, height = 20, dpi = 300)
+
+# Distribution
+qq_plot_mass <- qq_plots_single(data_expl, "delta_mass_rescaled", "mass")
+ggsave(here("./output/figures/exploratory/qq_plot_mass.png"), qq_plot_mass, width = 20, height = 20, dpi = 300)
+#### Delta mass considered as log-normal distribution
+
+####################################
+# B) Exploring distribution of the main variables for behav_clean

### 2025-05-25 R/exploratory.R
Update exploratory.R

@@ -27,3 +27,51 @@ ggsave(here("./output/figures/exploratory/qq_plot_mass.png"), qq_plot_mass, widt
 
 ####################################
 # B) Exploring distribution of the main variables for behav_clean
+data_expl <- behav_clean %>%
+  mutate(treatment = paste(cort, temp, sep = "-")) %>%
+  mutate(
+    lat_move = as.numeric(lat_move),
+    lat_shelter = as.numeric(lat_shelter)
+  )
+
+# Variables to explore
+var <- c("lat_move", "lat_shelter")
+
+# Histograms
+hist_list <- list()
+for (i in var) {
+  bin_width <- (max(data_expl[[i]], na.rm = TRUE) - min(data_expl[[i]], na.rm = TRUE)) / sqrt(length(data_expl[[i]]))
+  # Labels
+  if (i == "lat_move"){
+     lab <- "Latency to move (s)"
+  } else if (i == "lat_shelter") {
+     lab <- "Latency to shelter (s)"
+  }
+  # Create histogram
+  hist_plot <- ggplot(data_expl, aes(x = .data[[i]])) +
+    geom_histogram(binwidth = bin_width, fill = "#062d00", color = "black", alpha = 0.5) +
+    theme_classic() +
+    labs(x = lab, y = "Counts") +
+    theme(axis.title = element_text(size = 12, family = "Times"),
+          axis.text = element_text(size = 10, family = "Times"))
+  hist_list[[i]] <- hist_plot
+}
+hist_behav <- plot_grid(plotlist = hist_list, ncol = 3)
+ggsave(here("./output/figures/exploratory/hist_behav.png"), hist_behav, width = 20, height = 20, dpi = 300)
+
+# Distribution
+qq_plots_list <- lapply(var, function(v) {
+  if (i == "lat_move"){
+     label <- "Latency to move (s)"
+  } else if (i == "lat_shelter") {
+     label <- "Latency to shelter (s)"
+  } else {
+    v  # Fallback to variable name if not found
+  }
+  qq_plots_single(data_expl, v, label)
+})
+#
+behav_qq_plot <- plot_grid(plotlist = qq_plots_list, ncol = 1)
+ggsave(here("./output/figures/exploratory/behav_qq_plot.png"), behav_qq_plot, width = 20, height = 20, dpi = 300)
+#
+#

### 2025-05-25 ms.qmd
Update ms.qmd

@@
 ```{r, models_mass}
-#| label: models
-#| echo: false
-#| warning: false
-
-# Delta mass L. delicata
-  deli_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
+#| label: models_mass
+#
+# A) L. delicata
+deli_mod_mass <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
   plot(deli_mod)
@@
 bayes_R2(guich_mod)
-
-      # Seems that only food ingestion is significant in both species
-
-# Frozen behaviour L. delicata
-  deli_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = deli_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(deli_mod)
-  plot(deli_mod)
-bayes_R2(deli_mod)
-
-# Frozen behaviour L. delicata
- guich_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = guich_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(guich_mod)
-  plot(guich_mod)
-bayes_R2(guich_mod)
-
-# Extract posteriors
-posterior <- posterior_samples(mod, pars = "^b")
-plyr::ldply(lapply(posterior, mean))
-plyr::ldply(lapply(posterior, function(x) quantile(x, c(0.025, 0.975))))
-temp23 <- c(posterior[,"b_Intercept"], (posterior[,"b_Intercept"]+ posterior[,"b_trtB_23"]))
-temp28 <- c((posterior[,"b_Intercept"] + posterior[,"b_trtB_28"]), posterior[,"b_trtB_23"])
-
-pmcmc(temp23 - temp28)
-dharma(GMM_deli2)
-
 ```

### 2025-06-02 ms.qmd
Update ms.qmd

@@
 
+```{r, models_behaviour}
+#| label: models_behaviour
+#
+refit <- FALSE
+# A) L. delicata
+deli_beh <- behav_clean %>% 
+  filter(species == "delicata")
+if(refit){
+  deli_m_emergence <- brm(emergence ~ cort*temp*day + sex + (1 + day|lizard_id) + (1| clutch),
+                        data = deli_beh,
+                        family = bernoulli(link = "logit"),
+                        chains = 4, cores = 4, iter = 4000, warmup = 1000,
+                        control = list(adapt_delta = 0.99, max_treedepth =12))
+  saveRDS(deli_m_emergence, file = here(#ADD FOLDER))
+} else {
+  deli_m_emergence <- readRDS(here(#ADD FOLDER))
+}
+summary(deli_m_emergence)
+plot(deli_m_emergence)
+```
+
 ```{r, plotsMass}

### 2025-06-09 ms.qmd
updates

--- 2025-05-25 ms.qmd
@@
 ---
+title: "Effects of prenatal CORT and temperature on the response to a stressor"
+format:
   docx: 
+    bibliography: "./bib/refs.bib"
+    csl: "./bib/animal-behaviour.csl"
     reference-doc: "./bib/tmpl.docx"
-bibliography: "./bib/refs.bib"
 execute:
@@
   warning: false
+link-citations: true
 crossref:  
-  fig-title: Figure    # (default is "Figure")
+  fig-title: Fig    # (default is "Figure")
   title-delim: —     # (default is ":")
   fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
+  tbl-prefix: Table  # (default is "Table")
 editor_options: 
@@
 guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]).
-This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.
+**This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.**
 
@@
 ## Results
+
+```{r, models_behaviour}
+#| label: models_behaviour
+source(here("R", "func.R"))
+#
+refit <- FALSE
+#
+# A) L. delicata
+deli_beh <- behav_clean %>% 
+  filter(species == "delicata") %>%
+  mutate(lat_move = lat_move + 1, 
+         lat_shelter = lat_shelter + 1)
+#
+deli_move <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_move")
+deli_shelter <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_shelter")
+deli_emergence <- fit_m(df = deli_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "deli", var = "emergence")
+#
+#
+# B) L. guichenoti
+guich_beh <- behav_clean %>%
+  filter(species == "guichenoti") %>%
+  mutate(lat_move = lat_move + 1, 
+         lat_shelter = lat_shelter + 1)
+#
+guich_move <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_move")
+guich_shelter <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_shelter")
+guich_emergence <- fit_m(df = guich_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "guich", var = "emergence")
+```
+
 ```{r, models_mass}
 #| label: models_mass
+source(here("R", "func.R"))
+#
+refit <- FALSE
 #
 # A) L. delicata
-deli_mod_mass <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  plot(deli_mod)
-  summary(deli_mod)
-# Delta mass L. guichenoti
-  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(guich_mod)
-bayes_R2(guich_mod)
+deli_mass <- mass_clean %>%
+  filter(species == "delicata")
+#
+deli_mass <- fit_m(df = deli_mass, fam = lognormal(), type = "mass", sp = "deli", var = "delta_mass_rescaled")
+# 
+#
+# B) L. guichenoti
+guich_mass <- mass_clean %>%
+  filter(species == "guichenoti")
+#
+guich_mass <- fit_m(df = guich_mass, fam = lognormal(), type = "mass", sp = "guich", var = "delta_mass_rescaled")
+```
+
+```{r, clean_posteriors}
+#| label: clean_posteriors
+source(here("R", "func.R"))
+#
+variables <- c("move", "shelter", "emergence", "mass")
+species <- c("deli", "guich")
+# Initialize empty lists
+posterior_list <- list(deli = list(), guich = list())
+# Loop over species and variables
+for(i in species){
+  for(j in variables){
+    model_object <- get(paste0(i, "_", j))
+    result <- tidy_post(df = model_object)
+    posterior_list[[i]][[j]] <- result
+  }
+}
+#
+deli_post <- dplyr::bind_rows(posterior_list$deli, .id = "variable")
+guich_post <- dplyr::bind_rows(posterior_list$guich, .id = "variable")
 ```
@@
 
-We calculated $R^2$ as follows
+## Discussion
 
-$$
-R^2 = \frac{SS_{reg}}{SS_{tot}}
-$$ {#eq-r2}
+## References
 
-@eq-r2
+## Supplementary material
 
-```{r, fig1}
-#| label: fig-fig1
-#| fig-cap: "Activity budget of the focal group of chacma baboons (Papio ursinus) at the Cape Peninsula, South Africa, during the study period.
-The data are presented as the mean percentage of time spent in each activity category per hour (± SE)."
+_Table **S1**.
+Estimates of each predictor for each variable modelled for L.
+delicata.
+Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
+In bold, the values that are significant at pMCMC < 0.05._
 
-data %>% ggplot(aes(x = Lizard_id))  + geom_histogram()
-
+```{r, tablesummary_deli}
+#| label: tablesummary_deli
+#
+# Table defaults
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+# Make Table
+table_summ_deli <- flextable(deli_post) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  align(align = "center", j = c(3:6), part = "body") %>%
+  align(align = "center", j = c(1:6), part = "header") %>%
+  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(10,20,30), part = "body") %>% # Add horizontal lines
+  autofit()
+#
+table_summ_deli
 ```
 
-We have a total of `r length(unique(data$Lizard_id))`
-
-```{r,tb1}
-#| label: tbl-tb1
-#| tbl-cap: "Summary of the data"
-
-tab <- data %>% 
-  group_by(Lizard_id) %>% 
-  summarise(
-    n = n()
-  ) 
-flextable(tab[2:3,])
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
 ```
 
-# References
+_Table **S1**.
+Estimates of each predictor for each variable modelled for L.
+delicata.
+Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
+In bold, the values that are significant at pMCMC < 0.05._
+
+```{r, tablesummary_guich}
+#| label: tablesummary_guich
+#
+# Table defaults
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+# Make Table
+table_summ_guich <- flextable(guich_post) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  align(align = "center", j = c(3:6), part = "body") %>%
+  align(align = "center", j = c(1:6), part = "header") %>% 
+  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(10,20,30), part = "body") %>% # Add horizontal lines
+  autofit()
+#
+table_summ_guich
+```

### 2025-06-09 .vdoc.26166f2e-632f-4878-bc18-27c3d828c6b6.r

@@ -1,225 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
-
-source(here("R", "func.R"))
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: models
-#| echo: false
-#| warning: false
-
-# Delta mass L. delicata
-  deli_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  plot(deli_mod)
-  summary(deli_mod)
-# Delta mass L. guichenoti
-  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(guich_mod)
-bayes_R2(guich_mod)
-
-      # Seems that only food ingestion is significant in both species
-
-
-# Frozen behaviour L. delicata
-  deli_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = deli_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(deli_mod)
-  plot(deli_mod)
-bayes_R2(deli_mod)
-
-# Frozen behaviour L. delicata
- guich_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = guich_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
-  summary(guich_mod)
-  plot(guich_mod)
-bayes_R2(guich_mod)
-
-
-# Extract posteriors
[…]

### 2025-06-09 .vdoc.r

@@ -1,332 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)
-
-source(here("R", "func.R"))
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: dataload
-#| echo: false
-#| warning: false
- 
-# Load data
-	data  <-  read.csv("./data/Stressor.csv")
-#
-#
-#
-#| label: dataexplore
-#| echo: false
-#| warning: false
-
-# Divide Treatment in CORT and Temperature Transform mass into mg
-data <- data %>%
-      mutate(temp = gsub("[AB]_", "", trt),
-          cort = gsub("_[2][38]", "", trt))  %>% 
-      mutate(mass=mass*1000) %>%
-data.frame()
-data$cort <- factor(data$cort, 
-        levels = c("A" ,"B"),
-        labels = c("Control", "CORT"))
-data$temp <- factor(data$temp,
-        levels = c("23", "28"),
-        labels = c("Cold", "Hot"))
-data$Day <- factor(data$Day,
-        levels = c("Before", "After"))  
-     
-# MASS
-# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
-data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
- names_from = Day, values_from = mass)
-
-# Estimate mass increase and rescale it to avoid negative values
-data_2 <- data_2 %>% 
- mutate(delta_mass=After-Before) %>%
- mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
-  NA, 
-  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
-write.csv(data_2, file = "./data/Stressor_2.csv")
-
-# Split new database by species
-deli_stressor_2 <- data_2 %>% filter (species == "delicata")
-guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")
-
-# Explore distribution of mass increase
-  # A) L.deli
[…]

### 2025-06-09 R/data_process.R

@@ -25,7 +25,7 @@ mass <- read.csv(here("data", "mass.csv")) %>%
   mutate(lizard_id = as.character(lizard_id))%>%
 data.frame() #Read in mass data and transform lizard_id to character
 mass_clean <- merge(clean_ind, mass, by = "lizard_id", all = TRUE) %>%
-  mutate(delta_mass_rescaled = (delta_mass - min(delta_mass, na.rm = TRUE))) %>%
+  mutate(delta_mass_rescaled = (delta_mass - min(delta_mass, na.rm = TRUE) + 1)) %>%
 data.frame() #Merge mass data with clean_ind data and rescale delta_mass
 
 write.csv(mass_clean, here("./output/data_clean/mass_clean.csv")) 

### 2025-06-09 R/exploratory.R

@@ -56,7 +56,7 @@ for (i in var) {
           axis.text = element_text(size = 10, family = "Times"))
   hist_list[[i]] <- hist_plot
 }
-hist_behav <- plot_grid(plotlist = hist_list, ncol = 3)
+hist_behav <- plot_grid(plotlist = hist_list, ncol = 2)
 ggsave(here("./output/figures/exploratory/hist_behav.png"), hist_behav, width = 20, height = 20, dpi = 300)
 
 # Distribution
@@ -71,7 +71,7 @@ qq_plots_list <- lapply(var, function(v) {
   qq_plots_single(data_expl, v, label)
 })
 #
-behav_qq_plot <- plot_grid(plotlist = qq_plots_list, ncol = 1)
+behav_qq_plot <- plot_grid(plotlist = qq_plots_list, ncol = 2)
 ggsave(here("./output/figures/exploratory/behav_qq_plot.png"), behav_qq_plot, width = 20, height = 20, dpi = 300)
 #
-#
+#### Latency to move and latency to go to shelter considered as log-normal distribution

### 2025-06-09 R/func.r

@@ -1,24 +1,7 @@
 #################### 
 ####################
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom,
-tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x)
 #
-####################
-####################
-# Extract sample size per group
-#' @title sample
-#' @description Estract sample size per group
-#' @param df Database
-#' @param corti To select cort treatment ("CORT"/"Control")
-#' @param therm To select temp treatment ("Cold"/"Hot")
-sample <- function(df, corti, therm){
-  sample_size <- df %>%
-    filter(cort == corti, temp == therm) %>%
-    distinct(lizard_id) %>%
-    summarise(total_count = n()) %>%
-    pull(total_count)
-  return(sample_size)
-}
 ####################
 ####################
 # Function to generate four Q-Q plots for one variable
@@ -60,31 +43,53 @@ qq_plots_single <- function(df, vle, label) {
 }
 ####################
 ####################
-# Fit models and extract posteriors both per each treatment:
-#' @title fit_m Function
+# Extract sample size per group
+#' @title sample
+#' @description Estract sample size per group
+#' @param df To select the df
+#' @param corti To select cort treatment ("CORT"/"Control")
+#' @param therm To select temp treatment ("Cold"/"Hot")
+sample <- function(df, corti, therm){
+  sample_size <- df %>%
+                filter(cort == corti, temp == therm) %>%
+                group_by(lizard_id) %>%
+                summarise(n = n()) %>%
+                summarise(total_count = sum(n)) %>%
+                pull(total_count)
+  return(sample_size)
+}
+####################
+####################
+# Fit models and extract posteriors.
+#' @title fit_m_beh Function
 #' @description Fit brm models for different the associative task
 #' @param df To select the df
-#' @param cat To label whether the models are the preliminary or not
-#' @param var To select the variable
-#' @param formula To select the most appropriate formula
[…]

### 2025-06-10 ms.qmd
updates

@@
 
-The skinks used in our experiments came from a breeding colony established in the lab since 2019.
-The colony is compound by a total total of 270 adults of *L.
-delicata* and 180 adults of *L.* guichenoti that are housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
+_Breeding colony_: The skinks used in our experiments came from a breeding colony established in the lab since 2019.
+The colony is compound by a total total of 270 adults of _L.
+delicata_ and 180 adults of _L.
+guichenoti_ that are housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
 These enclosures are provided with non-stick matting, shelter, and several small water dishes.
 Lizards are given water daily, and fed approx.
-40 mid-size crickets (*Acheta domestica*) per enclosure three days a week.
+40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
 Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius; but to ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
+Room temperatures are set to 22-24 ºC, but to ensure a temperature gradient we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
+Heat lamp and heat cord keep the warm side of enclosures at 32 ºC.
 
-Eggs were collected between mid-October 2022 to the end of February 2023.
+_Egg collection_: Eggs were collected between mid-October 2022 to the end of February 2023.
 We placed in the commonal enclosure small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside so females had a place to lay the eggs.
@@
 Once a clutch was found, we measured eggs' length and width with a digital caliper to the nearest 0.1 mm and mass with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-To test the effect of prenatal CORT and temperature, we manipulated CORT levels in eggs and then incubated them at one of two temperature regimes (‘Cold’ – 23ºC ± 3ºC or ‘Hot’ – 30ºC ± 3ºC) in a 2x2 factorial design.
-CORT was increased by topical aplication of a solution of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at concentrations 2 standard deviations above the mean natural hormone concentration (non-published data); while a 'Control' group received an equal volume of the vehicle.
-Then, eggs were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite) covered with cling wrap to retain moisture, and were left in LATWIT 2X5D-R1160 incubators at two different temperatures ('Cold' or 'Hot') until hatchling.
+To test the effect of prenatal CORT and temperature, we manipulated CORT levels in eggs and then incubated them at one of two temperature regimes (see ....
+below).Then, eggs were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite) covered with cling wrap to retain moisture, and were left in LATWIT 2X5D-R1160 incubators at two different temperatures ('Cold' or 'Hot') until hatchling.
 
-INSERT Fig 1
-
-The incubators were checked three times a week for hatchlings.
+_Hatchlings_: The incubators were checked three times a week for hatchlings.
 When hatchligns were found, we recorded Snout-vent length (SVL) and tail length (TL) with a rule to the nearest mm while mass was measured with a (OHAUS, Model spx123) digital scale ± 0.001g error.
@@
 The rooms temperature was set to between 22-24 Celsius.
+
+FIG METHODS
 
@@
   align(align = "center", j = c(1:6), part = "header") %>%
-  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
+  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor"),
        bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
   flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
@@
   align(align = "center", j = c(1:6), part = "header") %>% 
-  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
+  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor"),
        bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
   flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column

### 2025-06-10 R/func.r

@@ -74,10 +74,10 @@ fit_m <- function(df, fam, sp, type, var) {
   if(refit){
     if(type == "beh"){
       formula <- as.formula(paste(var, 
-                            "~ cort*temp*day + age_start + sex + (1 + day|lizard_id) + (1|clutch)"))
+                            "~ cort*day + temp + (1|lizard_id) + (1|clutch)"))
     } else if(type == "mass"){
       formula <- as.formula(paste(var, 
-                            "~ cort*temp*food_ingested + age_start + sex + (1|clutch)"))
+                            "~ cort*food_ingested + temp + (1|clutch)"))
     }
     # Fit the model
     model <- brm(formula = formula,

### 2025-06-11 ms.qmd
updates

@@
 #
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
+species <- c("delicata", "guichenoti")
+trt <- c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot")
 #
+# Loop over species
+for (s in species) {
   n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = mass_clean, corti = hormone[k], therm = temperature[l])
+
+  data_n <- mass_clean %>% 
+    filter(species == s)
+
+  for (k in trt) {
+    list_name <- paste0(k)
+    group_data <- data_n %>% filter(trt == k)
+    n_list[[list_name]] <- nrow(group_data)
     }
+  assign(paste0("n_list_", s), n_list)
 }
-#
 ```
@@
 
-We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L.
-delicata*).
+We used two species of skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
+guichenoti_).
 These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]).
@@
 
-_Breeding colony_: The skinks used in our experiments came from a breeding colony established in the lab since 2019.
-The colony is compound by a total total of 270 adults of _L.
+*Breeding colony* - Lizards came from a breeding colony established in the laboratory since 2019.
+This colony consisted of 270 adults of _L.
 delicata_ and 180 adults of _L.
-guichenoti_ that are housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-These enclosures are provided with non-stick matting, shelter, and several small water dishes.
-Lizards are given water daily, and fed approx.
+guichenoti_ housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with two males and four females per enclosure.
+Enclosures were provided with shelter, nonstick matting, and several small water dishes.
+The lizards were given water daily and were fed approx.
 40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
-Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperatures are set to 22-24 ºC, but to ensure a temperature gradient we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
-Heat lamp and heat cord keep the warm side of enclosures at 32 ºC.
+The crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
+Room temperature was set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures at 34 ºC.
 
-_Egg collection_: Eggs were collected between mid-October 2022 to the end of February 2023.
-We placed in the commonal enclosure small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside so females had a place to lay the eggs.
-Three times a week, we checked these boxes for the presence of eggs.
-Once a clutch was found, we measured eggs' length and width with a digital caliper to the nearest 0.1 mm and mass with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-To test the effect of prenatal CORT and temperature, we manipulated CORT levels in eggs and then incubated them at one of two temperature regimes (see ....
-below).Then, eggs were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite) covered with cling wrap to retain moisture, and were left in LATWIT 2X5D-R1160 incubators at two different temperatures ('Cold' or 'Hot') until hatchling.
+*Eggs collection and incubation* - Eggs were collected between mid-October 2022 to the end of February 2023.
+We placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures to provide females with a place to lay the eggs.
+These boxes were checked three days a week.
+After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighted the eggs with a (OHAUS, Model spx123) digital scale with an accuracy of ± 0.001 g error.
+Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.
 
-_Hatchlings_: The incubators were checked three times a week for hatchlings.
-When hatchligns were found, we recorded Snout-vent length (SVL) and tail length (TL) with a rule to the nearest mm while mass was measured with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-Then, hatchlings were enclosed in individual tubs (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish.
-During this period, they were sprayed water every day and received 3-6 small *A.
-domestica* crickets three times a week.
+*Hatchlings* - Incubators were checked three times a week for hatchlings.
+Lizards were measured and weighed immediately after hatching.
+Snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a (OHAUS, Model spx123) digital scale with an accuracy of ± 0.001 g.
+Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with nonstick matting and a small water dish.
 All care otherwise follows similar protocols to adults (see above).
 
-### Experimental arena and design
-The experiments were done:
-The sample size was:
-The experimental arena was:
-The experimental design was:
-The lizards used in this experiment were subjected before to a personality test and two types of learning tasks.
-These experiments are not considered stresful, and, as such, we do not expect them to affect the results of the present study.
-However, we left three weeks between those experiments and the present one to control for possible effects of the previous tests.
-After 3 months, we measured again SVL, TL, and mass of the juveniles and then they were moved to the experimental arena (see below).
-The day before starting the experiment, lizards were measured again, and then moved to a new enclosure the same size and conditions before, but in the room where the CCTV system is held () experimental arena (Fig.
-1c) for acclimatation.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other.
-These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below).
-The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum.* We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
-The rooms temperature was set to between 22-24 Celsius.
+#### CORT and Temperature manipulation
 
-FIG METHODS
+To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them at cold (23 ± 3 ºC) or hot (28 ± 3 ºC) conditions (see @fig-Methods A).
+We used a partial split clutch design where eggs from a given clutch were distributed equally across the four treatments when clutch sizes were larger than four and randomly across treatments when less than four.
+Eggs were topically supplied with either 5 µL of crystalline corticosterone (Sigma, Cat.
+No.
+C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or an equal volume of 100% Ethanol (Control treatment).
+We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
+Mean incubation temperatures represent the lower and upper limits of the natural range of nest temperatures in _L.
+delicata_ [mean nesting temperatures = 27.4 ºC; @cheetham2011embryonic].
+Nonetheless, the higher temperature treatment (28 ºC) was above the thermal optima estimated for _L.
+delicata_ [T~opt~ = 25 ºC; @pettersen2023maternal] and reached more stressful temperatures daily as fluctuations occurred (i.e., 31 ºC).
+
+#### Response to acute and long-term stressors
+
+Three weeks before the start of the behavioural tests, we adapted the lizard enclosures to the experimental setup.
+The enclosures used for the stress response trials were the same as those where the lizards had previously been housed, but modified to include only a single shelter (9 × 6 × 1.5 cm) placed at one end and a water dish in the centre.
+All other shelter materials and the matting substrate were removed to allow continuous monitoring.
+The enclosures were then relocated to the experimental rooms, each containing six racks equipped with their own CCTV system (model DVR-HP210475).
+The number of lizards per treatment was counterbalanced across racks to control for potential effects of room or rack position.
+In these three weeks, lizards were fed one calcium- and multivitamin-dusted cricket per day, and water was provided _ad libitum_.
+A temperature gradient was maintained using heat cords and heat lamps on a 12:12 h light–dark cycle, with room temperatures kept between 22–24 °C.
+
+To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack once daily over multiple days [**insert relevant reference here**].
+_The simulated attack involved chasing the lizard for 60 seconds using a soft paintbrush, after which its behaviour was recorded for one hour._ **DETAILS OF EACH TRIAL**.
+All videos were later analysed by **MATHEO**, who was blinded to the lizards’ treatment groups.
+From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile (frozen) immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
+A lizard was considered to have resumed movement if it moved continuously for at least five seconds, and to be under shelter if no limbs were visible.
+This procedure was repeated daily for eight consecutive days to assess the effects of mid-term exposure to a stressor.
+All trials were conducted between 1000–1400 h, when lizards are most active.
+
+Lizard mass was measured to the nearest 0.001 g using a digital scale (OHAUS, model SPX123) both before and after the completion of the behavioural tests.
+Change in mass was calculated as the difference between final and initial mass and then rescaled by subtracting the smallest mass change observed in the dataset, to avoid negative values.
+This rescaled change in mass was used as a proxy for the physiological cost of repeated stress exposure.
+To control for hunger levels and potential effects of food ingestion on mass change, we also recorded the number of crickets ingested by each lizard during the acclimation period.
+To do so, we placed a known number of crickets in each enclosure after each of the trials, and counted the number of crickets remaining after 24 hours.
+The total number of crickets ingested during the eight days of tests was then used as a covariate in the analyses of mass change.
+The experimental design is summarised in @fig-Methods B.
+
+```{r, fig-Methods}
+#| label: fig-Methods
+#| fig.cap: ""
+
+knitr::include_graphics("./Others/Methods.png")
+
+```
 
@@
 #
-refit <- FALSE
+refit <- TRUE
 #
@@
   bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(10,20,30), part = "body") %>% # Add horizontal lines
+  flextable::compose(i = c(2:5,7:10,12:15,17:20), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(5,10,15), part = "body") %>% # Add horizontal lines
   autofit()

### 2025-06-11 R/data_process.R

@@ -8,15 +8,13 @@ source(here("R", "func.R"))
 # A) Cleaning the general_data.csv for all individuals
 clean_ind <- read.csv(here("data", "general_data.csv")) %>%
   mutate(lizard_id = as.character(lizard_id)) %>%
-  mutate(temp = gsub("[AB]_", "", trt),
-         cort = gsub("_[2][38]", "", trt))  %>%
-  mutate(temp = factor(temp,
-                       levels = c("23", "28"),
-                       labels = c("23" = "Cold", "28" = "Hot"))) %>%
-  mutate(cort = factor(cort,
-                       levels = c("A", "B"),
-                       labels = c("A" = "Control", "B" = "CORT"))) %>%
-  dplyr::select(-notes, -trt) %>%
+  mutate(trt = factor(trt,
+                        levels = c("A_23", "B_23", "A_28", "B_28"),
+                        labels = c("A_23" = "Control-Cold",
+                                  "B_23" = "CORT-Cold",
+                                  "A_28" = "Control-Hot",
+                                  "B_28" = "CORT-Hot"))) %>%
+  dplyr::select(-notes) %>%
 data.frame()
 
 ####################################
@@ -34,6 +32,7 @@ write.csv(mass_clean, here("./output/data_clean/mass_clean.csv"))
 # C) Merging behaviour.csv with clean individual data (clean_ind)
 behav <- read.csv(here("data", "behaviour.csv"))%>%
   mutate(lizard_id = as.character(lizard_id))%>%
+  mutate(day = day - 1) %>% # Adjust day to start from 0
 data.frame() #Read in behaviour data and transform lizard_id to character
 behav_clean <- merge(clean_ind, behav, by = "lizard_id", all = TRUE) 
 

### 2025-06-11 R/func.r

@@ -74,10 +74,10 @@ fit_m <- function(df, fam, sp, type, var) {
   if(refit){
     if(type == "beh"){
       formula <- as.formula(paste(var, 
-                            "~ cort*day + temp + (1|lizard_id) + (1|clutch)"))
+                            "~ day*trt + (1|lizard_id) + (1|clutch)"))
     } else if(type == "mass"){
       formula <- as.formula(paste(var, 
-                            "~ cort*food_ingested + temp + (1|clutch)"))
+                            "~ food_ingested*trt + (1|clutch)"))
     }
     # Fit the model
     model <- brm(formula = formula,

### 2025-06-11 ms.qmd
updates

@@
 #
-clutches <- mass_clean %>% 
+clutches_deli <- mass_clean %>% 
+  filter(species == "delicata") %>%
+  distinct(clutch) %>% 
+  nrow()
+#
+clutches_guich <- mass_clean %>% 
+  filter(species == "guichenoti") %>%
   distinct(clutch) %>% 
@@
 guichenoti_).
-These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]).
-Both species have similar breeding periods, but reproductive output in *L.
-delicata* is one clutch of 1 to 6 eggs per season, while *L.
-guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]).
-**This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.**
+These are oviparous, generalist, small (∼35-55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia [@chapple_know_2011; @chapple_biology_2014].
+However, _L.
+delicata_ has been reported as a successful invader in New Zealand and Hawaii, while there are not such reports for _L.
+guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
+Invassivenes does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [**REFpers**], which may be linked to differences in their stress response [@koolhaas1999coping].
 
@@
 To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack once daily over multiple days [**insert relevant reference here**].
-_The simulated attack involved chasing the lizard for 60 seconds using a soft paintbrush, after which its behaviour was recorded for one hour._ **DETAILS OF EACH TRIAL**.
-All videos were later analysed by **MATHEO**, who was blinded to the lizards’ treatment groups.
+Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the shelter.
+The lizard was then chased for 60 seconds using a soft paintbrush, simulating a predatory attack.
+Immediately after the chase, the shelter — but not the water dish — was returned to its original position, marking the start of the behavioural observation period.
+The lizard’s behaviour was then recorded for one hour.
+All videos were later analysed by **MATHEO**, who was blinded to the lizards’ treatment.
 From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile (frozen) immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
@@
 
-Lizard mass was measured to the nearest 0.001 g using a digital scale (OHAUS, model SPX123) both before and after the completion of the behavioural tests.
-Change in mass was calculated as the difference between final and initial mass and then rescaled by subtracting the smallest mass change observed in the dataset, to avoid negative values.
+Both before and after the completion of the behavioural tests, lizard mass was measured to the nearest 0.001 g using a digital scale.
+Change in mass was calculated as the difference between final minus the initial mass and then rescaled by adding the smallest mass change observed in the dataset, to avoid negative values.
 This rescaled change in mass was used as a proxy for the physiological cost of repeated stress exposure.
@@
 #| label: fig-Methods
-#| fig.cap: ""
+#| fig.cap: "Scheme of our experimental design.
+In panel A, the early environment manipulations resulting in four experimental conditions.
+In panel B, the timeline of the postnatal stress-response assay.
+On Day 0, lizard mass was recorded.
+From Days 1–8, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
+Variables measured included freezing duration (latency to move), latency to shelter, and emergence from shelter.
+Final mass was measured on Day 9."
 
@@
 
+We performed the analyses for each species separately.
+We fitted all models using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.4.0 [@R].
+We ran four parallel MCMC chains of 10000 iterations for each model, with a warmup period of 4000 iterations.
+
+We used as the response variables the latency to move, latency to shelter, and emergence from the shelter after the simulated attack, as well as the change in mass after the eight days of stress exposure.
+We fitted separate models for each response variable.
+For the three behavioural variables, we included the prenatal condition (Control-Cold, CORT-Cold, Control-Hot, CORT-Hot), the day of the trial, and their interaction as fixed effects.
+To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e.
+the intercept reflects lizard behaviour in response to the first exposure to the stressor).
+For the variable mass, we included the prenatal condition, the total number of crickets ingested during the eight days of stress exposure, and their interaction as predictors.
+
+Lizard identity was included as a random factor in the behavioural models, but it was not included when mass was the response.
+For all models, we included the clutch as a random factor.
+_L.
+delicata_ lays one clutch per year, while _L.
+guichenoti_ lays two [@chapple_know_2011; @chapple2015deliinvLHI].
+Since eggs were collected during half of the breeding season, clutches likely come from different mothers.
+Additionally, previous research has shown that clutches are generally sired by a single male, but sperm storage can occur [@kar2023heritability].
+Given our partial split-clutch design, and the fact that maternal effects are expected to be stronger than paternal effects in these species, including the clutch as a random factor should account for the effects of parental identity.
+
+We used the posterior distributions of model parameters to test for differences between prenatal conditions.
+For behavioural variables, differences in the model intercepts between conditions were interpreted as differences in the acute response to a stressor (i.e.
+the first day of exposure), while differences in the slopes were interpreted as differences in behavioural adjustment across repeated exposures (i.e.
+chronic stress response).
+Similarly, for mass change, we interpreted differences in the condition-by-ingestion interaction (i.e.
+variation in the slopes of mass change across treatment groups) as reflecting differences in energy assimilation efficiency with increased food intake.
+_pMCMC_ was used to test the hypothesis that posterior distributions of slopes and slope contrasts differed from zero.
+We considered an effect statistically significant when _pMCMC_ <0.05.
+
 ## Results
+
+We started with 96 lizards, 48 per species and 12 per treatment per species.
+However, our final sample size was 84 due to natural mortality (n = 10).
+The final sample sizes per treatment and species are listed in @fig-results_beh_deli, @fig-results_beh_guich and @fig-results_mass.
+These animals came from a total of `r clutches_deli` clutches in _L.
+delicata_ and `r clutches_guich` in _L.
+guichenoti_.
 
@@
 #
-refit <- TRUE
+refit <- FALSE
 #
@@
 
-```{r, clean_posteriors}
+```{r, post_contrasts}
+#| label: post_contrasts
+#  
+source(here("R", "func.R"))
+#
+# Getting the values for each treatments' intercept and slope for the contrasts later
+## A) L. delicata
+deli_move_posteriors <- post_values(deli_move, var = "beh")
+deli_shelter_posteriors <- post_values(deli_shelter, var = "beh")
+deli_emergence_posteriors <- post_values(deli_emergence, var = "beh")
+deli_mass_posteriors <- post_values(deli_mass, var = "mass")
+#
+## B) L. guichenoti
+guich_move_posteriors <- post_values(guich_move, var = "beh")
+guich_shelter_posteriors <- post_values(guich_shelter, var = "beh")
+guich_emergence_posteriors <- post_values(guich_emergence, var = "beh")
+guich_mass_posteriors <- post_values(guich_mass, var = "mass")
+#
+#
+# Getting the contrasts for the posterior distributions
+## A) L. delicata
+deli_move_contrasts <- contrasts_func(sp = "deli", res = "move")
+deli_shelter_contrasts <- contrasts_func(sp = "deli", res = "shelter")
+deli_emergence_contrasts <- contrasts_func(sp = "deli", res = "emergence")
+deli_mass_contrasts <- contrasts_func(sp = "deli", res = "mass")
+#
+## B) L. guichenoti
+guich_move_contrasts <- contrasts_func(sp = "guich", res = "move")
+guich_shelter_contrasts <- contrasts_func(sp = "guich", res = "shelter")
+guich_emergence_contrasts <- contrasts_func(sp = "guich", res = "emergence")
+guich_mass_contrasts <- contrasts_func(sp = "guich", res = "mass")
+```
+
+#### Do prenatal conditions affect the behavioural response to an acute stressor?
+
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to move (see Table S1 in _Supplementary Material_).
+We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see Table S1 in _Supplementary Material_).
+
+_Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had any effect on the behavioural response to an acute stressor on _L.
+guichenoti_ (see Table S2 in _Supplementary Material_).
+
+```{r, fig-results_beh_deli}
+
+```
+
+#### Do prenatal conditions affect the behavioural response to a chronic stressor?
+
+_Lampropholis delicata_: **HOW WERE THE SLOPES OVER TIME??**.
+We did not find any effects of prenatal conditions in _L.
+delicata_ response to a chronic stressor, as there were no significant differences in the slopes of the different treatments for any of the behavioural variables (see Table S1 in _Supplementary Material_).
+
+_Lampropholis guichenoti_: **HOW WERE THE SLOPES OVER TIME??**.
+There were no significant differences in the behavioural response of _L.
+guichenoti_ to the repeated exposure to an stressor (see Table S2 in _Supplementary Material_).
+
+```{r, fig-results_beh_guich}
+```
+
+#### Do prenatal conditions influence the costs of repeated stress exposure?
+
+_Lampropholis delicata_:
+
+_Lampropholis guichenoti_:
+
+```{r, fig-results_mass}
+```
+
+## Discussion
+
+## References
+
+## Supplementary material
+
+_Table S1.
+Contrasts of the posterior distributions of the predictors for each behavioural variable modelled for L.
+delicata.
+The contrasts are between the intercepts and slopes of the different treatments.
+Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
+In bold, the values that are significant at pMCMC < 0.05._
+
+```{r, table_contrasts_deli}
+#| label: table_contrasts_deli
+#
+source(here("R", "func.R"))
+# 
+# Get the posterior contrasts for L. delicata and merge them into a single df
+contrast_deli_beh <- bind_rows(move = deli_move_contrasts,
+                              shelter = deli_shelter_contrasts,
+                              emergence = deli_emergence_contrasts,
+                              .id = "variable") %>%
+                    mutate(across(everything(), as.character)) %>%
+
+                    pivot_longer(cols = -variable,
+                                names_to = c("type", "factor", "stat"),
+                                names_pattern = "(mean|pMCMC)_(.*)_(int|slope)") %>%
+
+                    pivot_wider(names_from = c(type, stat),
+                                values_from = value) %>%
+data.frame()
+write.csv(contrast_deli_beh, here("./output/checking/contrasts_deli_beh.csv"))
+#
+#
+# Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+table_deli_behaviour <- flextable(contrast_deli_beh) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 2, value = as_paragraph("Predictor"), part = "header") %>% # Change the second column header
+  flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
+  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("pMCMC"), part = "header") %>% # Change the fourth column header
+  flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
+  flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move"), part = "body") %>% # Change the first row of the first column
+  flextable::compose(i = 4, j = 1, value = as_paragraph("Latency to shelter"), part = "body") %>% # Change the fourth row of the first column
+  flextable::compose(i = 7, j = 1, value = as_paragraph("Emergence"), part = "body") %>% # Change the seventh row of the first column
+  flextable::hline(i = c(3,6), part = "body") %>% # Add horizontal lines
+  flextable::vline(j = c(2,4), part = "all") %>% # Add vertical lines
+  align(align = "center", j = c(3:6), part = "body") %>%
+  align(align = "center", i = c(1,2), j = c(1:6), part = "header") %>%
+  flextable::bold(i = 1, j = c(3,4), bold = TRUE, part = "body") %>% 
+autofit()
+#
+table_deli_behaviour
+```
+
+Contrasts were done by:  
+    - _Temperature_: β~Hot~ - β~Cold~  
+    - _CORT_: β~CORT~ - β~Control~  
+    - _Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+_Table S2.
+Contrasts of the posterior distributions of the predictors for each behavioural variable modelled for L.
+guichenoti.
+The contrasts are between the intercepts and slopes of the different treatments.
+Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
+In bold, the values that are significant at pMCMC < 0.05._
+
+```{r, table_contrasts_guich}
+#| label: table_contrasts_guich
+#
+source(here("R", "func.R"))
+# Get the posterior contrasts for L. guichenoti and merge them into a single df
+contrast_guich_beh <- bind_rows(move = guich_move_contrasts,
+                              shelter = guich_shelter_contrasts,
+                              emergence = guich_emergence_contrasts,
+                              .id = "variable") %>%
+                    mutate(across(-variable, as.character)) %>%
+
+                    pivot_longer(cols = -variable,
+                                names_to = c("type", "factor", "stat"),
+                                names_pattern = "(mean|pMCMC)_(.*)_(int|slope)") %>%
+
+                    pivot_wider(names_from = c(type, stat),
+                                values_from = value) %>%
+data.frame()
+write.csv(contrast_guich_beh, here("./output/checking/contrasts_guich_beh.csv"))
+#
+#
+# Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+table_guich_behaviour <- flextable(contrast_guich_beh) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 2, value = as_paragraph("Predictor"), part = "header") %>% # Change the second column header
+  flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
+  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("pMCMC"), part = "header") %>% # Change the fourth column header
+  flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
+  flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move"), part = "body") %>% # Change the first row of the first column
+  flextable::compose(i = 4, j = 1, value = as_paragraph("Latency to shelter"), part = "body") %>% # Change the fourth row of the first column
+  flextable::compose(i = 7, j = 1, value = as_paragraph("Emergence"), part = "body") %>% # Change the seventh row of the first column
+  flextable::hline(i = c(3,6), part = "body") %>% # Add horizontal lines
+  flextable::vline(j = c(2,4), part = "all") %>% # Add vertical lines
+  align(align = "center", j = c(3:6), part = "body") %>%
+  align(align = "center", i = c(1,2), j = c(1:6), part = "header") %>% 
+autofit()
+#
+table_guich_behaviour
+```
+
+Contrasts were done by:  
+    - _Temperature_: β~Hot~ - β~Cold~  
+    - _CORT_: β~CORT~ - β~Control~  
+    - _Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage") 
+```
+
+_Table S3.
+Values of the slopes for the different treatments for each behavioural variable modelled for both species.
+The values are the posterior means and 95% credible intervals (CI).
+In bold, the values that are different from 0, tested by pMCMC < 0.05._
+
+**INSERT TABLE FOR SLOPES**
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage") 
+```
+
+_Table S4.
+Contrasts for mass modelled for both species.
+The contrasts are between the slopes of the different treatments.
+Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
+In bold, the values that are significant at pMCMC < 0.05._
+
+**INSERT TABLE FOR MASS**
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r, clean_posteriors_tables}
 #| label: clean_posteriors
@@
 ```
-
-```{r, plotsMass}
-#| label: plotsMass
-#| echo: false
-#| warning: false
-
-# Plot delta Mass L. delicata
-plot_deli <- ggplot(deli_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
-  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
-  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
-    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
-  ) +
-  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
-  theme_classic() +
-  theme(
-    axis.title = element_text(size = 18, family = "sans"),  
-    axis.text = element_text(size = 10, family = "sans"),  
-    legend.title = element_text(size = 14, family = "sans"), 
-    legend.text = element_text(size = 12, family = "sans")  
-  )
-print(plot_deli)
-ggsave("./output/figures/fig_deli_Mass.png", plot=plot_deli, width = 18, height = 9, units = "cm", dpi = 3000)
-# Plot delta Mass L. guichenoti
-plot_guich <- ggplot(guich_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
-  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
-  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
-    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
-  ) +
-  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
-  theme_classic() +
-  theme(
-    axis.title = element_text(size = 18, family = "sans"),  
-    axis.text = element_text(size = 10, family = "sans"),  
-    legend.title = element_text(size = 14, family = "sans"), 
-    legend.text = element_text(size = 12, family = "sans")  
-  )
-print(plot_guich)
-ggsave("./output/figures/fig_guich_Mass.png", plot=plot_guich, width = 18, height = 9, units = "cm", dpi = 3000)
-```
-
- ```{r, plotsFrozenbehaviour}
-#| label: plotsFrozenbehaviour
-#| echo: false
-#| warning: false
-
-# Plot time frozen L. delicata
-plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
-  geom_boxplot(outlier.shape = NA) +
-  ylim(0, 10) +
-  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
-    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
-  ) +
-  labs(y = "Time frozen (log-transformed)", x = "Day", fill="Treatment") +
-  theme_classic() +
-  theme(
-    axis.title = element_text(size = 18, family = "sans"),  
-    axis.text = element_text(size = 12, family = "sans"),  
-    legend.title = element_text(size = 14, family = "sans"), 
-    legend.text = element_text(size = 12, family = "sans")  
-  )
-print(plot_deli_BEH)
-ggsave("./output/figures/fig_deli_BEH.png", plot=plot_deli_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
-
-# Plot time frozen L. guichenoti
-plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
-  geom_boxplot(outlier.shape = NA) +
-  ylim(0, 8) +
-  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
-    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
-  ) +
-  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
-  theme_classic() +
-  theme(
-    axis.title = element_text(size = 18, family = "sans"),  
-    axis.text = element_text(size = 12, family = "sans"),  
-    legend.title = element_text(size = 14, family = "sans"), 
-    legend.text = element_text(size = 12, family = "sans")  
-  )
-print(plot_guich_BEH)
-ggsave("./output/figures/fig_guich_BEH.png", plot=plot_guich_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
-
-```
-
-## Discussion
-
-## References
-
-## Supplementary material
 
@@
   bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:5,7:10,12:15,17:20), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(5,10,15), part = "body") %>% # Add horizontal lines
+  flextable::compose(i = c(2:8,10:16,18:24,26:32), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
   autofit()
@@
 ```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-06-11 R/func.r

@@ -168,51 +168,112 @@ tidy_post <- function(df) {
     )
   return(summary_df)
 }
-
-
-
 ####################
 ####################
 # Function to get estimates to make the contrasts between treatments
 #' @title post_values
 #' @param df To select the df
-#' @param fac To indicate if there are other parameters to take into account
-post_values <- function(df, fac){
+#' @param var To indicate the kind of variable to use (beh or mass)
+post_values <- function(df, var){
   df <- as.data.frame(df)
-  #Getting the estimated values per each treatment
-  if(fac == "none"){
-    Control_Cold <- df$b_Intercept
-    Control_Hot <- df$b_Intercept + df$b_tempHot
-    CORT_Cold <- df$b_Intercept + df$b_cortCORT
-    CORT_Hot <- df$b_Intercept + df$b_cortCORT + df$b_tempHot + df$`b_cortCORT:tempHot`
-  } else if (fac == "sex"){
-    # Males
-    Control_Cold_males <- base::sample(df$b_Intercept, size = 12000, replace = FALSE)
-    Control_Hot_males <- base::sample(df$b_Intercept + df$b_tempHot, size = 12000, replace = FALSE)
-    CORT_Cold_males <- base::sample(df$b_Intercept + df$b_cortCORT, size = 12000, replace = FALSE)
-    CORT_Hot_males <- base::sample(df$b_Intercept + df$b_cortCORT + df$b_tempHot +
-                                      df$`b_cortCORT:tempHot`, size = 12000, replace = FALSE)
-    # Females
-    Control_Cold_fem <- base::sample(df$b_Intercept + df$b_sexFemale, size = 12000, replace = FALSE)
-    Control_Hot_fem <- base::sample(df$b_Intercept + df$b_tempHot + df$b_sexFemale, size = 12000, replace = FALSE)
-    CORT_Cold_fem <- base::sample(df$b_Intercept + df$b_cortCORT + df$b_sexFemale, size = 12000,
-                                    replace = FALSE)
-    CORT_Hot_fem <- base::sample(df$b_Intercept + df$b_cortCORT + df$b_tempHot +
-                                    df$`b_cortCORT:tempHot` + df$b_sexFemale, size = 12000, replace = FALSE)
-    # Pool
-    CORT_Cold <- c(CORT_Cold_males, CORT_Cold_fem)
[…]

### 2025-06-26 ms.qmd
updates

@@
 
+Animals facing stressors/disruptors and examples.  
+Mechanisms involved in the response to stressors/disruptors: endocrinology.
+Consequences of stressors/disruptors on animals: short-term and long-term.
+
+Predation as a stressor/disruptor.
+
+Can early environment influence the response to stressors/disruptors: physiological AND behavioural effects of GC exposure. Effects on antipredatory behaviour.
+
+Can early environment influence the response to stressors/disruptors: other aspects. Why is temperature important? How is related to antipredatory behaviour?
+
 Hypothesis PRESENTATION:
@@
 Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity.
-In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals, which are more sensitive to stressors.
+In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals [**REFs**], which are more sensitive to stressors.
 As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter.
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to move (see Table S1 in _Supplementary Material_).
-We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-deli and Table S1 in _Supplementary Material_).
+We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-deli and Table S1 in _Supplementary Material_).
 
 _Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had any effect on the behavioural response to an acute stressor on _L.
-guichenoti_ (see Table S2 in _Supplementary Material_).
+guichenoti_ (see @fig-guich and Table S2 in _Supplementary Material_).
 
@@
 ```
+
+#### Do both species differ in their behavioural response to an stressor?
+
+- Acute stressor: differences between intercepts (put everything together or compare between incubation conditions?)
+
+- Chronic stressor: differences between slopes
 

### 2025-06-26 R/func.r

@@ -270,7 +270,68 @@ contrasts_func <- function(sp, res){
                           pMCMC_Interaction_slope = pMCMC_int_slope)
   data_table <- dplyr::bind_rows(data_table, data_temp)
 }
+####################
+####################
+# Function to get estimates to make the contrasts between treatments
+#' @title slopes_func
+#' @param sp To select the species ("guich" or "deli")
+#' @param res To select the response variable ("move", "shelter", "emergence", "mass")
+slopes_func <- function(sp, res){
+  data_table <- data.frame()
+  df <- get(paste0(sp, "_", res, "_posteriors"))
+  # Intercept contrasts
+  Temperature_int <- format_dec(mean(c(df$int_CORT_Hot, df$int_Control_Hot)) 
+                                - mean(c(df$int_CORT_Cold, df$int_Control_Cold)), 3)
 
+  pMCMC_temp_int <- format_p(pmcmc(c(df$int_CORT_Hot, df$int_Control_Hot) 
+                                  - c(df$int_CORT_Cold, df$int_Control_Cold)), 3, equal = FALSE)
+  #
+  CORT_int <- format_dec(mean(c(df$int_Control_Hot, df$int_Control_Cold)) 
+                        - mean(c(df$int_CORT_Hot, df$int_CORT_Cold)), 3)
+
+  pMCMC_cort_int <- format_p(pmcmc(c(df$int_Control_Hot, df$int_Control_Cold) 
+                                  - c(df$int_CORT_Hot, df$int_CORT_Cold)), 3, equal = FALSE)
+  #
+  Interaction_int <- format_dec((mean(df$int_Control_Hot) - mean(df$int_CORT_Hot)) 
+                              - (mean(df$int_Control_Cold) - mean(df$int_CORT_Cold)), 3)
+
+  pMCMC_int_int <- format_p(pmcmc((df$int_Control_Hot - df$int_CORT_Hot) 
+                              - (df$int_Control_Cold - df$int_CORT_Cold)), 3, equal = FALSE)
+  # Slope contrasts
+  Temperature_slope <- format_dec(mean(c(df$slope_CORT_Hot, df$slope_Control_Hot)) 
+                                - mean(c(df$slope_CORT_Cold, df$slope_Control_Cold)), 3)
+
+  pMCMC_temp_slope <- format_p(pmcmc(c(df$slope_CORT_Hot, df$slope_Control_Hot) 
+                                    - c(df$slope_CORT_Cold, df$slope_Control_Cold)), 3, equal = FALSE)
+  #
[…]

### 2025-07-04 ms.qmd
updates

@@
 title: "Effects of prenatal CORT and temperature on the response to a stressor"
-format:
 docx: 
@@
 crossref:  
-  fig-title: Fig    # (default is "Figure")
+  fig-title: Fig.    # (default is "Figure")
   title-delim: —     # (default is ":")
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-deli and Table S1 in _Supplementary Material_).
-We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-deli and Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli and Table S1 in _Supplementary Material_).
+We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli and Table S1 in _Supplementary Material_).
 
 _Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had any effect on the behavioural response to an acute stressor on _L.
-guichenoti_ (see @fig-guich and Table S2 in _Supplementary Material_).
+guichenoti_ (see @fig-results_beh_guich and Table S2 in _Supplementary Material_).
 
 ```{r, fig-results_beh_deli}
+#| label: fig-results_beh_deli
+#| fig.cap: "Behavioural response of Lampropholis delicata to an stressor."
 
+#### Panels A and D show the predicted probability of choosing the correct ramp in the first trial (Decision first trial).
+In panels B and E, the distribution of the estimates of slopes per each treatment.
+In all A, B, D, and E the x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
+The different colors indicate the different treatments.
+Points and bars represent the mean and standard deviation of the mean of the estimates, respectively.
+Dashed lines indicate value 0.33 (the probability of choosing correctly by chance) in panels A, D, and 0 in panels B, E.
+Asterisks indicate significant results (p~mcmc~ < 0.05).
+Panels C and F show the predicted probability of choosing the correct ramp first over trials.
+The lines represent the mean predicted probability of choosing the correct ramp for each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained using the slope and intercept estimates from the posterior distributions.
+The different colors indicate the different treatments.
+source(here("R", "func.R"))
+# First step, create the dfs for all variables
+## ACUTE STRESSORS
+deli_move_acute <- plotting(sp = "deli", res = "move", type = "int") 
+deli_move_chronic <- plotting(sp = "deli", res = "move", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+deli_move_trials <- plotting(sp = "deli", res = "move", type = "trials")
+#
+deli_shelter_acute <- plotting(sp = "deli", res = "shelter", type = "int")
+deli_shelter_chronic <- plotting(sp = "deli", res = "shelter", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+deli_shelter_trials <- plotting(sp = "deli", res = "shelter", type = "trials")
+#
+deli_emergence_acute <- plotting(sp = "deli", res = "emergence", type = "int")
+deli_emergence_chronic <- plotting(sp = "deli", res = "emergence", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+deli_emergence_trials <- plotting(sp = "deli", res = "emergence", type = "trials")
+#
+# Getting the legend
+legend_plot <- plotting(sp = "deli", res = "move", type = "int") + 
+  theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplotGrob(legend_plot)
+legend <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+# Merge the plots
+deli_beh <- plot_grid(deli_move_acute, deli_move_chronic, deli_move_trials,
+                      deli_shelter_acute, deli_shelter_chronic, deli_shelter_trials,
+                      deli_emergence_acute, deli_emergence_chronic, deli_emergence_trials,
+                      ncol = 3, nrow = 3,
+                      rel_widths = c(0.75, 0.75, 1))
+#
+fig_deli_beh <- ggdraw() +
+  # Make background white
+  theme(plot.background = element_rect(fill = "white", color = NA)) +
+  # Insert legend 
+  draw_grob(legend, x = 0.49, y = 0.015, width = 0.01, height = 0.01) +
+  # Add the final plot but with reduced size for the legend and "title"
+  draw_plot(deli_beh, x = 0, y = 0.04, width = 1, height = 0.91) +
+  # Insert title for each plot
+  annotate("text", x = 0.02, y = 0.948, label = "A", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.948, label = "B", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.948, label = "C", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.02, y = 0.645, label = "D", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.645, label = "E", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.645, label = "F", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.02, y = 0.341, label = "G", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.341, label = "H", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.341, label = "I", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  # Insert name species
+  annotate("text", x = 0.58, y = 0.99, label = "Lampropholis delicata", hjust = 1, vjust = 1, size = 6, fontface = "italic")
+#
+ggsave("./output/figures/main/fig_deli_beh.png", plot=fig_deli_beh, width = 21, height = 25, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/main/fig_deli_beh.png")
 ```
@@
 ```{r, fig-results_beh_guich}
+#| label: fig-results_beh_guich
+#| fig.cap: "Behavioural response of Lampropholis guichenoti to an stressor." 
+#
+source(here("R", "func.R"))
+# First step, create the dfs for all variables
+## ACUTE STRESSORS
+guich_move_acute <- plotting(sp = "guich", res = "move", type = "int") 
+guich_move_chronic <- plotting(sp = "guich", res = "move", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+guich_move_trials <- plotting(sp = "guich", res = "move", type = "trials")
+#
+guich_shelter_acute <- plotting(sp = "guich", res = "shelter", type = "int")
+guich_shelter_chronic <- plotting(sp = "guich", res = "shelter", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+guich_shelter_trials <- plotting(sp = "guich", res = "shelter", type = "trials")
+#
+guich_emergence_acute <- plotting(sp = "guich", res = "emergence", type = "int")
+guich_emergence_chronic <- plotting(sp = "guich", res = "emergence", type = "slop") +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black")
+guich_emergence_trials <- plotting(sp = "guich", res = "emergence", type = "trials")
+#
+# Getting the legend
+legend_plot <- plotting(sp = "guich", res = "move", type = "int") + 
+  theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplotGrob(legend_plot)
+legend <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+# Merge the plots
+guich_beh <- plot_grid(guich_move_acute, guich_move_chronic, guich_move_trials,
+                      guich_shelter_acute, guich_shelter_chronic, guich_shelter_trials,
+                      guich_emergence_acute, guich_emergence_chronic, guich_emergence_trials,
+                      ncol = 3, nrow = 3,
+                      rel_widths = c(0.75, 0.75, 1))
+#
+fig_guich_beh <- ggdraw() +
+  # Make background white
+  theme(plot.background = element_rect(fill = "white", color = NA)) +
+  # Insert legend 
+  draw_grob(legend, x = 0.49, y = 0.015, width = 0.01, height = 0.01) +
+  # Add the final plot but with reduced size for the legend and "title"
+  draw_plot(guich_beh, x = 0, y = 0.04, width = 1, height = 0.91) +
+  # Insert title for each plot
+  annotate("text", x = 0.02, y = 0.948, label = "A", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.948, label = "B", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.948, label = "C", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.02, y = 0.645, label = "D", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.645, label = "E", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.645, label = "F", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.02, y = 0.341, label = "G", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.318, y = 0.341, label = "H", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.615, y = 0.341, label = "I", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  # Insert name species
+  annotate("text", x = 0.58, y = 0.99, label = "Lampropholis guichenoti", hjust = 1, vjust = 1, size = 6, fontface = "italic")
+#
+ggsave("./output/figures/main/fig_guich_beh.png", plot=fig_guich_beh, width = 21, height = 25, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/main/fig_guich_beh.png")
 ```
@@
 ```{r, fig-results_mass}
+#| label: fig-results_mass
+#| fig.cap: "Change in mass of Lampropholis delicata and Lampropholis guichenoti after eight days of repeated stress exposure." 
+#
+#
+source(here("R", "func.R"))
+#
+## L delicata
+deli_mass_plot <- plotting(sp = "deli", res = "mass", type = "int") +
+  theme(legend.position = "none",
+        legend.title = element_blank())
+# Getting the legend
+legend_plot_deli <- plotting(sp = "deli", res = "mass", type = "int") + 
+  theme(legend.position = "right", legend.title = element_blank())
+gtable_deli <- ggplotGrob(legend_plot_deli)
+legend_deli <- gtable_deli$grobs[[which(sapply(gtable_deli$grobs, function(x) x$name) == "guide-box")]]
+#
+## L guichenoti
+guich_mass_plot <- plotting(sp = "guich", res = "mass", type = "int") +
+  theme(legend.position = "none",
+        legend.title = element_blank())
+# Getting the legend
+legend_plot_guich <- plotting(sp = "guich", res = "mass", type = "int") + 
+  theme(legend.position = "right", legend.title = element_blank())
+gtable_guich <- ggplotGrob(legend_plot_guich)
+legend_guich <- gtable_guich$grobs[[which(sapply(gtable_guich$grobs, function(x) x$name) == "guide-box")]]
+#
+mass_grid <- plot_grid(deli_mass_plot, NULL, guich_mass_plot, NULL,
+                       ncol = 4, nrow = 1,
+                       rel_widths = c(0.5, 0.5, 0.5, 0.45))
+#
+fig_mass <- ggdraw() +
+  # Make background white
+  theme(plot.background = element_rect(fill = "white", color = NA)) +
+  # Add the final plot but with reduced size for the legend and "title"
+  draw_plot(mass_grid, x = 0, y = 0, width = 1, height = 0.92) +
+  # Insert species name
+  annotate("text", x = 0.36, y = 0.99, label = "Lampropholis delicata", hjust = 1, vjust = 1, size = 5, fontface = "italic") +
+  annotate("text", x = 0.9, y = 0.99, label = "Lampropholis guichenoti", hjust = 1, vjust = 1, size = 5, fontface = "italic") +
+  # Insert title for each plot
+  annotate("text", x = 0.023, y = 0.91, label = "A", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  annotate("text", x = 0.535, y = 0.91, label = "B", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
+  # Insert picture species
+  draw_image(here("./Others/Deli.png"), x = 0.271, y = 0.65, width = 0.22, height = 0.22) +
+  draw_image(here("./Others/Guich.png"), x = 0.78, y = 0.65, width = 0.22, height = 0.22) +
+  # Insert legends
+  draw_grob(legend_deli, x = 0.375, y = 0.275, width = 0.01, height = 0.01) +
+  draw_grob(legend_guich, x = 0.88, y = 0.275, width = 0.01, height = 0.01)
+#
+ggsave("./output/figures/main/fig_mass.png", plot=fig_mass, width = 21, height = 8, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/main/fig_mass.png")
 ```
@@
 ```{r, clean_posteriors_tables}
-#| label: clean_posteriors
+#| label: clean_posteriors_tables
 source(here("R", "func.R"))
@@
 
-_Table **S1**.
+_Table S5.
 Estimates of each predictor for each variable modelled for L.
@@
   bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:8,10:16,18:24,26:32), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = c(2:8,10:16,18:24,26:29), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
@@
 
-_Table **S1**.
+_Table S6.
 Estimates of each predictor for each variable modelled for L.

### 2025-07-04 R/data_process.R

@@ -24,6 +24,7 @@ mass <- read.csv(here("data", "mass.csv")) %>%
 data.frame() #Read in mass data and transform lizard_id to character
 mass_clean <- merge(clean_ind, mass, by = "lizard_id", all = TRUE) %>%
   mutate(delta_mass_rescaled = (delta_mass - min(delta_mass, na.rm = TRUE) + 1)) %>%
+  mutate(food_ingested = food_ingested - mean(food_ingested, na.rm = TRUE))
 data.frame() #Merge mass data with clean_ind data and rescale delta_mass
 
 write.csv(mass_clean, here("./output/data_clean/mass_clean.csv")) 

### 2025-07-04 R/func.r

@@ -77,7 +77,7 @@ fit_m <- function(df, fam, sp, type, var) {
                             "~ day*trt + (1|lizard_id) + (1|clutch)"))
     } else if(type == "mass"){
       formula <- as.formula(paste(var, 
-                            "~ food_ingested*trt + (1|clutch)"))
+                            "~ trt + food_ingested + (1|clutch)"))
     }
     # Fit the model
     model <- brm(formula = formula,
@@ -152,7 +152,7 @@ format_p <- function(x, n, equal) {
 # Function to clean the posteriors and create a big df that includes
 # all variables and their predictors
 #' @title tidy_post
-#' @param df to select the posteriors to use. Options: choice or errors
+#' @param df to select the posteriors to use. 
 tidy_post <- function(df) {
   df <- as.data.frame(df)
   summary_df <- df %>%
@@ -165,7 +165,8 @@ tidy_post <- function(df) {
       CI_upper = format_dec(quantile(Estimate, 0.95, na.rm = TRUE), 3),
       pMCMC = format_dec(pmcmc(Estimate, null = 0, twotail = TRUE), 3),
       .groups = "drop"
-    )
+    ) %>%
+  data.frame()
   return(summary_df)
 }
 ####################
@@ -184,28 +185,30 @@ post_values <- function(df, var){
     slope_Control_Cold <- df$b_day
     slope_CORT_Cold <- df$b_day + df$`b_day:trtCORTMCold`
     slope_Control_Hot <- df$b_day + df$`b_day:trtControlMHot`
-    slope_CORT_Hot <- df$b_day + df$`b_day:trtCORTMHot`
+    slope_CORT_Hot <- df$b_day + df$`b_day:trtCORTMHot`    
+    # Create a data frame with the estimates
+    data_values <- data.frame(
+      int_Control_Cold = int_Control_Cold,
+      int_CORT_Cold = int_CORT_Cold,
+      int_Control_Hot = int_Control_Hot,
+      int_CORT_Hot = int_CORT_Hot,
+      slope_Control_Cold = slope_Control_Cold,
+      slope_CORT_Cold = slope_CORT_Cold,
+      slope_Control_Hot = slope_Control_Hot,
+      slope_CORT_Hot = slope_CORT_Hot)
+  
   } else if (var == "mass") {
     int_Control_Cold <- df$b_Intercept
     int_CORT_Cold <- df$b_Intercept + df$b_trtCORTMCold
[…]

### 2025-07-08 ms.qmd
updates

@@
 docx: 
+  number_sections: true
+  linestretch: 2
+  fontsize: 12pt
+  mainfont: Times New Roman
 bibliography: "./bib/refs.bib"
@@
 
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Mathéo Dupuy^1,5^, Daniel W.A. Noble^1^
+
+^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
+^2^ Flinders University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^3^ University of Wollongong, Wollongong, NSW 2500, Australia   
+^4^ Environmental Futures University of Wollongong, Wollongong, NSW 2500, Australia     
+^5^ Institute Universitaire de Technologie Paul Sabatier, Auch 32000, France  
+$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
+
+ORCID:
+
+Pablo Recio ORCID: 0000-0002-5890-0218  
+Dalton C. Leibold ORCID: 0000-0001-9645-2033  
+Ondi L. Crino ORCID: 0000-0001-5700-1387  
+Christopher R. Friesen ORCID: 0000-0001-5338-7454  
+Daniel W.A. Noble ORCID: 0000-0001-9460-8743
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 ## Introduction
 
-Animals facing stressors/disruptors and examples.  
-Mechanisms involved in the response to stressors/disruptors: endocrinology.
-Consequences of stressors/disruptors on animals: short-term and long-term.
+Animals constantly face events that can disrupt their homeostasis [**REFs**].
+These challenges can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [**REFs**].
+Regardless of the nature of the challenge, organisms must adjust to these alterations to maintain their physiological and behavioural integrity [**REFs**].
+In vertebrates, the ability to cope with allostasis is mediated by the hypothalamic-pituitary-adrenal (HPA) axis [**REF**] - the hypothalamic-pituitary-interrenal (HPI) axis in non-mammalians [**REFs**].
+This system elicits a physiological response that results in the secretion of glucocorticoids (GCs) [**REF**].
+Glucocorticoids are a group of hormones that help reallocate energetic resources to restore homeostasis following disruption [**REFs**].
+For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [**REFs**].
 
-Predation as a stressor/disruptor.
+In the short term, GCs help organisms cope with acute challenges and return to homeostasis.
+However, long-term exposure to GCs can carry significant metabolic costs [**REFs**].
+Prolonged elevations in GCs can lead to immune suppression, impaired growth, or permanent neural alterations [**REFs**].
+From a behavioural perspective, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [**REFs**].
+As such, animals are predicted to evolve mechanisms to appropriately respond to acute stressors while avoiding the costs in the long-term.
+For example, by regulating their responses to homotypic (repeated and similar) stressors, particularly those that are predictable and non-lethal [**REFs**; @grissom2009habituation].
+Habituation - the reduction of physiological responses elicited by exposure to a homotypic stressor [@pfister1979glucocorticosterone] - can be crucial to avoid the long-term costs of chronic stress without compromising the response to acute stressors.
 
-Can early environment influence the response to stressors/disruptors: physiological AND behavioural effects of GC exposure. Effects on antipredatory behaviour.
+Importantly, the process of physiological habituation shares key functional characteristics with the cognitive process of habituation - a form of non-associative learning where individuals reduce their responsiveness to a repeated stimulus [@thompson1966habituation].
+In both cases, organisms are repeatedly exposed to a stimulus/homotypic stressor and progressively decrease their response towards it [@grissom2009habituation].
+In fact, both behavioural and physiological habituation often co-occur.
+For example, rhesus macaques (_Macaca mulatta_) showed a decline in blood cortisol concentrations across consecutive days of restraint together with decreased vocalizations and behavioral agitation [@ruys2004behavioral].
+How both types of habituation interplay with each other depends on the nature of the stressor.
+Physiological habituation does not occur when animals are faced with systemic stressors like hypoglycemia or ether exposure whereas habituation to processive stressors — those involving primarily psychological qualities, such as restraint or simulated predation - is frequent [@grissom2009habituation].
+However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeteadly exposed to a homotypic processive stressor [@jaferi2006corticosterone; @grissom2009habituation].
+Adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
+While the HPA/HPI axis is paramount for physiological habituation, cognitive habituation also involve neural mechanisms related to sensory processing, learning, and behavioural control such as the limbic system, the superior colliculus, or the optic tectum [@grissom2009habituation; @dutta2014saliency].
+Habituation to processive stressors is therefore likely to involve complex interactions between physiological and cognitive processes.
 
-Can early environment influence the response to stressors/disruptors: other aspects. Why is temperature important? How is related to antipredatory behaviour?
+Crucially, the effectiveness of either system depends on how well the underlying neural structures develop and integrate.
+Many of these brain regions mature concurrently and can be highly plastic during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017].
+Early-life conditions can shape the structure and function of both neural systems, with long-term consequences for how individuals respond to acute and chronic stressors [@vargas2016early; @van2017modulation; @zito2017early; @crino_under_2020; @crino2024eggs].
+One of the key mediators of early-life effects is GC exposure.
+Elevations in GCs during development exert potent effects on HPA axis function, altering the response to stressors later in life [@vargas2016early; @zito2017early; @crino2024eggs].
+For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
+In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting cognitive processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004].
+Therefore, early-life exposure to GCs can have long-lasting effects on the response to both acute and chronic stressors, potentially leading to maladaptive responses in adulthood.
 
-Hypothesis PRESENTATION:
+Other environmental factors may also shape stress-related phenotypes, either independently or in combination with GCs.
+In ectotherms, early thermal conditions play a critical role in cognitive development [@amiel_effects_2017; @abayarathna_effects_2020].
+Incubation temperature affects cognitive abilities such as learning, memory, or perception, likely through changes in brain function and structure [@amiel_smart_2011; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017; @abayarathna_effects_2020].
+In addition, early thermal environment can also affect the development of the HPA/HPI axis [@crino_under_2020; @crino2024eggs].
+For example, in the delicata skink (_Lampropholis delicata_) individuals incubated at cooler temperatures showed increased baseline CORT concentration in the blood [@crino2024eggs].
+Although these differences are believed to be driven by size rather than direct endocrine effects [@crino2024eggs], they can still have important long-term consequences for how individuals respond to acute and chronic stressors.
+Importantly, incubation temperature and prenatal CORT may not act independently [see @Crino_2023], and their interaction could produce complex, trait-specific outcomes that shape both physiological and behavioural responses to stress.
+Yet, the interactive effects of these early-life conditions on stress responses remain largely unexplored.
 
-Both temperature and CORT influence the development of proactive-reactive personality types, associated at the same time with different HPA axis reactivity.
-In this sense, both increases in prenatal CORT and lower temperatures are related to more reactive individuals [**REFs**], which are more sensitive to stressors.
-As such, I predict individuals incubated at low temperature and/or treated with CORT to be more reactive, and thus to spent more time frozen after chasing them and not coming out from the shelter.
-At the same time I expect these individuals to have increased phisiological costs (mass lost) when submitted to a 'chronic' stressor.
+In this study, we investigate how prenatal CORT and incubation temperature shape the response to a stressor in two species of skinks: the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
+guichenoti_).
+We focus on the effects of these early-life conditions on the behavioural responses to a simulated predatory attack.
+Specifically, we examine how these factors influence the acute response to a stressor (i.e., the first exposure) and the chronic response (i.e., repeated exposures over time).
+Finally, we assessed body mass change across the stress period as a potential indirect cost, while controlling for food intake.
+We predicted that prenatal CORT exposure would heighten acute stress reactivity and slow habituation, while warmer incubation temperatures would reduce acute responses and enhance habituation.
+Furthermore, we predicted higher temperatures would buffer the effects of prenatal CORT on stress responses, potentially leading to more flexible behavioural adjustment.
+Finally, we expected that less flexible stress responses would lead to higher decreases in mass.
 
@@
 The lizard’s behaviour was then recorded for one hour.
-All videos were later analysed by **MATHEO**, who was blinded to the lizards’ treatment.
+All videos were later analysed by MD, who was blinded to the lizards’ treatment.
 From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile (frozen) immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
@@
 chronic stress response).
-Similarly, for mass change, we interpreted differences in the condition-by-ingestion interaction (i.e.
-variation in the slopes of mass change across treatment groups) as reflecting differences in energy assimilation efficiency with increased food intake.
-_pMCMC_ was used to test the hypothesis that posterior distributions of slopes and slope contrasts differed from zero.
-We considered an effect statistically significant when _pMCMC_ <0.05.
+For mass change, between treatments difference reflect differences in energetic cost of the chronic stressor.
+We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments or the slopes for each treatment were different from zero.
+
+Differences between species were tested by comparing the posterior distributions of the intercepts and slopes of the different treatments.
+We used the posterior distributions of the intercepts and slopes to test for differences between species in their response to an acute stressor (intercept) and a chronic stressor (slope).
 
@@
 However, our final sample size was 84 due to natural mortality (n = 10).
-The final sample sizes per treatment and species are listed in @fig-results_beh_deli, @fig-results_beh_guich and @fig-results_mass.
+The final sample sizes per treatment and species are listed in @fig-results_beh_deli, @fig-results_beh_guich, and @fig-results_mass.
 These animals came from a total of `r clutches_deli` clutches in _L.
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$mean_Temperature_int`; pMCMC `r deli_move_contrasts$pMCMC_Temperature_int`), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli and Table S1 in _Supplementary Material_).
-We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli and Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$median_Temperature_int`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
+We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
 
 _Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had any effect on the behavioural response to an acute stressor on _L.
-guichenoti_ (see @fig-results_beh_guich and Table S2 in _Supplementary Material_).
+guichenoti_ (see @fig-results_beh_guich A, D, G and Table S2 in _Supplementary Material_).
 
@@
 
-_Lampropholis delicata_: **HOW WERE THE SLOPES OVER TIME??**.
-We did not find any effects of prenatal conditions in _L.
-delicata_ response to a chronic stressor, as there were no significant differences in the slopes of the different treatments for any of the behavioural variables (see Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: CORT-treated lizards incubated at cold temperatures increased their latency to move over time (Slope (denoted hereafter as β) = `r format_dec(median(deli_move_posteriors$slope_CORT_Cold), 3)`; 95% HPDI = [`r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold, credMass = 0.95)[2], 3)`]), while the slopes were not different from zero in the rest of the treatments (see @fig-results_beh_deli B, C and Table S3 in _Supplementary Material_).
+However, constrasts between between treatments showed no effect of prenatal conditions on the latency to move througout the experiments (see Table S1 in _Supplementary Material_).
+Lizards from all treatments increased the probability of emerging from the shelter throughout the experiment (β > 0 in all cases, see @fig-results_beh_deli H, I and Table S3 in _Supplementary Material_), with no differences between treatments (see Table S1 in _Supplementary Material_).
+In contrast, the latency to shelter did not change over time (β = 0 in all cases, see @fig-results_beh_deli E, F and Table S3 in _Supplementary Material_).
+The contrasts between treatments showed no differences in the slopes (see Table S1 in _Supplementary Material_).
 
-_Lampropholis guichenoti_: **HOW WERE THE SLOPES OVER TIME??**.
+_Lampropholis guichenoti_: Behavioral responses towards a stressor did not change over time in _L.
+guichenoti_ (see @fig-results_beh_guich B, C, E, F, H, I and Table S3 in _Supplementary Material_).
 There were no significant differences in the behavioural response of _L.
-guichenoti_ to the repeated exposure to an stressor (see Table S2 in _Supplementary Material_).
+guichenoti_ to the repeated exposure to an stressor (see @fig-results_beh_guich B, C, E, F, H, I and Table S2 in _Supplementary Material_).
 
@@
 
-_Lampropholis delicata_:
+```{r, process_posteriors_mass}
+#| label: process_posteriors_mass
+# Code to make the posterior distribution of the mass models easier to call for comparissons of Δmass
+source(here("R", "func.R"))
+#
+# A) L. delicata
+mass_deli <- deli_mass_posteriors %>%
+  mutate(across(everything(), ~ exp(.) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
+  rename_with(~ str_remove(., "^int_"))
+# B) L. guichenoti
+mass_guich <- guich_mass_posteriors %>%
+  mutate(across(everything(), ~ exp(.) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
+  rename_with(~ str_remove(., "^int_"))
+```
 
-_Lampropholis guichenoti_:
+_Lampropholis delicata_: Overall, animals grew over time (Median Δmass = `r format_dec(median(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot)),3)` mg; 95% HDPI = [`r format_dec(hdi(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot), credMass = 0.95)[1],3)` ,`r format_dec(hdi(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot), credMass = 0.95)[2],3)`]), with no effect of temperature, CORT, or their interaction (see @fig-results_mass A and Table S4 in _Supplementary Material_).
+
+_Lampropholis guichenoti_: Mass of lizards increased over time (Median Δmass = `r format_dec(median(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot)),3)` mg; 95% HDPI = [`r format_dec(hdi(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot), credMass = 0.95)[1],3)` ,`r format_dec(hdi(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot), credMass = 0.95)[2],3)`]), and the contrasts between treatments showed no differences in the change in mass (see @fig-results_mass B and Table S4 in _Supplementary Material_).
 
@@
 
-#### Do both species differ in their behavioural response to an stressor?
+#### Do both species differ in their response to an stressor?
 
-- Acute stressor: differences between intercepts (put everything together or compare between incubation conditions?)
+_Acute stressor_: Latency to move after the first exposure in Control-Hot incubated _L.
+delicata_ was higher than for _L.
+guichenoti_ (I~L.
+deli~ - I~L.
+guich~ = `r format_dec(median(deli_move_posteriors$int_Control_Hot) - median(guich_move_posteriors$int_Control_Hot), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$int_Control_Hot - guich_move_posteriors$int_Control_Hot, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$int_Control_Hot - guich_move_posteriors$int_Control_Hot, credMass = 0.95)[2], 3)`]), but there were no differences in Latency to move between the rest of the treatments (see @table_contrasts_species).
+In contrast, the probability of emerging from the shelter after an acute stressor was lower in cold incubated _L.
+delicata_ than _L.
+guichenoti_ for both control (I~L.
+deli~ - I~L.
+guich~ = `r format_dec(median(deli_emergence_posteriors$int_Control_Cold) - median(guich_emergence_posteriors$int_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$int_Control_Cold - guich_emergence_posteriors$int_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$int_Control_Cold - guich_emergence_posteriors$int_Control_Cold, credMass = 0.95)[2], 3)`]) and CORT-treated (I~L.
+deli~ - I~L.
+guich~ = `r format_dec(median(deli_emergence_posteriors$int_CORT_Cold) - median(guich_emergence_posteriors$int_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$int_CORT_Cold - guich_emergence_posteriors$int_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$int_CORT_Cold - guich_emergence_posteriors$int_CORT_Cold, credMass = 0.95)[2], 3)`]) lizards, but not in hot incubated animals (see @table_contrasts_species).
+There were no differences between species in the latency until sheltering after an acute stressor (see @table_contrasts_species).
 
-- Chronic stressor: differences between slopes
+_Chronic stressor_: Over time, _L.
+delicata_ CORT-treated lizards incubated at cold temperatures increased latency to move more than _L.
+guichenoti_ (β~L.
+deli~ - β~L.
+guich~ = `r format_dec(median(deli_move_posteriors$slope_CORT_Cold) - median(guich_move_posteriors$slope_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
+The probability of emerging from the shelter increased over time in _L.
+delicata_ more than in _L.
+guichenoti_ (see @table_contrasts_species), but those differences were significant only in the Control-Cold treatment (β~L.
+deli~ - β~L.
+guich~ = `r format_dec(median(deli_emergence_posteriors$slope_Control_Cold) - median(guich_emergence_posteriors$slope_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
+There were no significant differences in the changes on latency to shelter between species (see @table_contrasts_species).
+
+_Mass change_: increases in mass were not different between species in any of the treatments (see @table_contrasts_species).
+
+_Table 1.
+Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L.
+guichenoti.
+Contrast represent differences in median parameters between species (L.
+delicata - L.
+guichenoti) for each treatment.
+95% Highest Posterior Density Intervals (95% HPDIs) indicate the probability that contrasts differ from zero._
+
+```{r, table_contrasts_species}
+#| label: table_contrasts_species
+#| table.caption: "Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L. guichenoti."
+source(here("R", "func.R"))
+#
+contrasts_sp <- bind_rows(move_deli = deli_move_posteriors,
+                          move_guich = guich_move_posteriors,
+                          shelter_deli = deli_shelter_posteriors,
+                          shelter_guich = guich_shelter_posteriors,
+                          emergence_deli = deli_emergence_posteriors,
+                          emergence_guich = guich_emergence_posteriors,
+                          mass_deli = deli_mass_posteriors,
+                          mass_guich = guich_mass_posteriors,
+                          .id = "data") %>%
+                pivot_longer(cols = -c(data),
+                            names_to = c("stress", "treatment"),
+                            names_pattern = "(int|slope)_(.*)") %>%
+                # Pivot to wide format
+                mutate(column_id = paste(data, stress, treatment, sep = "_")) %>%
+                group_by(column_id) %>%
+                mutate(draw = row_number()) %>%
+                ungroup() %>%
+                dplyr::select(column_id, draw, value) %>%
+                pivot_wider(names_from = column_id, values_from = value)
+# Make loop to get contrasts
+var <- c("move", "shelter", "emergence", "mass")
+stress <- c("int", "slope")
+treatment <- c("Control_Cold", "CORT_Cold", "Control_Hot", "CORT_Hot")
+contrasts_df <- data.frame()
+for(v in var){
+  for(st in stress){
+    for(t in treatment){
+      data_loop <- data.frame()
+
+      col_deli <- paste0(v, "_deli_", st, "_", t)
+      col_guich <- paste0(v, "_guich_", st, "_", t)
+      #
+      vector_deli <- contrasts_sp %>% pull(!!sym(col_deli))
+      vector_guich <- contrasts_sp %>% pull(!!sym(col_guich))
+
+      contrasts_median <- format_dec(median(vector_deli) - median(vector_guich), 3)
+      contrasts_lowlim <- format_dec(hdi(vector_deli - vector_guich, credMass = 0.95)[1], 3)
+      contrasts_uplim <- format_dec(hdi(vector_deli - vector_guich, credMass = 0.95)[2], 3)
+
+      data_loop <- data.frame(variable = v,
+                              stress = st,
+                              treatment = t,
+                              contrast_median = contrasts_median,
+                              low_lim = contrasts_lowlim,
+                              up_lim = contrasts_uplim)
+
+      contrasts_df <- bind_rows(contrasts_df, data_loop)
+    }
+  }
+}
+# Modify the df to get it ready for the table
+contrasts_df2 <- contrasts_df %>%
+  pivot_wider(names_from = stress,
+              values_from = c(contrast_median, low_lim, up_lim)) %>%
+  mutate(treatment = str_replace_all(treatment, "_", "-"),
+        treatment = factor(treatment,
+                          levels = c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot"))) %>%
+  mutate(HPDI_int = paste0("[", low_lim_int, ", ", up_lim_int, "]"),
+         HPDI_slope = paste0("[", low_lim_slope, ", ", up_lim_slope, "]")) %>%
+  dplyr::select(variable, treatment, contrast_median_int, HPDI_int, 
+                contrast_median_slope, HPDI_slope) %>%
+data.frame()       
+write.csv(contrasts_df2, here("./output/checking/contrasts_species.csv"))
+#
+# Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+table_contrasts_species <- flextable(contrasts_df2) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 2, value = as_paragraph("Treatment"), part = "header") %>% # Change the second column header
+  flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
+  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
+  flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
+  flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move"), part = "body") %>% 
+  flextable::compose(i = 5, j = 1, value = as_paragraph("Latency to shelter"), part = "body") %>% 
+  flextable::compose(i = 9, j = 1, value = as_paragraph("Emergence"), part = "body") %>% 
+  flextable::compose(i = 13, j = 1, value = as_paragraph("Δmass"), part = "body") %>% # Change the first column's names
+  flextable::compose(i = c(13:16), j = c(5,6), value = as_paragraph("-"), part = "body") %>%
+  flextable::hline(i = c(4,8,12), part = "body") %>% # Add horizontal lines
+  flextable::vline(j = c(2,4), part = "all") %>% # Add vertical lines
+  align(align = "center", j = c(3:6), part = "body") %>%
+  align(align = "center", i = c(1,2), j = c(1:6), part = "header") %>% 
+  flextable::bold(i = c(3,9,10), j = c(3,4), bold = TRUE, part = "body") %>% 
+  flextable::bold(i = c(2,9), j = c(5,6), bold = TRUE, part = "body") %>%
+autofit()
+#
+table_contrasts_species
+```
 
@@
 The contrasts are between the intercepts and slopes of the different treatments.
-Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
-In bold, the values that are significant at pMCMC < 0.05._
+95% Highest Posterior Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
 
@@
 
+                    rename_with(~ str_replace_all(., "X2\\.5\\.HDI_", "low_lim_")) %>%
+                    rename_with(~ str_replace_all(., "X97\\.5\\.HDI_", "up_lim_")) %>%
+
                     pivot_longer(cols = -variable,
                                 names_to = c("type", "factor", "stat"),
-                                names_pattern = "(mean|pMCMC)_(.*)_(int|slope)") %>%
+                                names_pattern = "(median|low_lim|up_lim)_(.*)_(int|slope)") %>%
 
@@
                                 values_from = value) %>%
+
+                    mutate(CI_int = paste0("[", low_lim_int, ", ", up_lim_int, "]"),
+                          CI_slope = paste0("[", low_lim_slope, ", ", up_lim_slope, "]")) %>%
+
+                    dplyr::select(variable, factor, median_int, CI_int,
+                                  median_slope, CI_slope) %>%
 data.frame()
@@
   flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
-  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("pMCMC"), part = "header") %>% # Change the fourth column header
+  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
   flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
@@
 The contrasts are between the intercepts and slopes of the different treatments.
-Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
-In bold, the values that are significant at pMCMC < 0.05._
+95% Highest Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
 
@@
 source(here("R", "func.R"))
+# 
 # Get the posterior contrasts for L. guichenoti and merge them into a single df
@@
                               .id = "variable") %>%
-                    mutate(across(-variable, as.character)) %>%
+                    mutate(across(everything(), as.character)) %>%
+
+                    rename_with(~ str_replace_all(., "X2\\.5\\.HDI_", "low_lim_")) %>%
+                    rename_with(~ str_replace_all(., "X97\\.5\\.HDI_", "up_lim_")) %>%
 
@@
                                 names_to = c("type", "factor", "stat"),
-                                names_pattern = "(mean|pMCMC)_(.*)_(int|slope)") %>%
+                                names_pattern = "(median|low_lim|up_lim)_(.*)_(int|slope)") %>%
 
@@
 Values of the slopes for the different treatments for each behavioural variable modelled for both species.
-The values are the posterior means and 95% credible intervals (CI).
-In bold, the values that are different from 0, tested by pMCMC < 0.05._
+95% Highest Density Intervals (95% HPDI) test the hypothesis that the slopes are different from zero._
 
-**INSERT TABLE FOR SLOPES**
+```{r, table_slopes}
+#| label: table_slopes
+#
+source(here("R", "func.R"))
+#
+# Get the slopes for both species in a common df
+sp <- c("deli", "guich")
+var <- c("move", "shelter", "emergence")
+#
+slopes_table_df <- data.frame()
+for(i in sp){
+  for(j in var){    
+    model_object <- get(paste0(i, "_", j, "_posteriors")) %>%      
+      dplyr::select(starts_with("slope_")) %>%
+      rename_with(~ str_replace_all(., "slope_", "")) %>%
+      pivot_longer(cols = everything(), names_to = "treatment", values_to = "slope") %>%
+      group_by(treatment) %>%
+      summarize(
+        Median = format_dec(median(slope), 3),
+        lowlim = format_dec(hdi(slope, credMass = 0.95)[1], 3),
+        uplim = format_dec(hdi(slope, credMass = 0.95)[2], 3)) %>%
+      mutate(species = i, variable = j)
+
+    slopes_table_df <- rbind(slopes_table_df, model_object)
+  }
+}
+#
+table_slopes_df <- slopes_table_df %>%
+  mutate(CI = paste0("[", lowlim, ", ", uplim, "]")) %>%
+  dplyr::select(species, variable, treatment, Median, CI) %>%
+  mutate(treatment = str_replace_all(treatment, "_", "-"),
+         treatment = factor(treatment, levels = c("Control-Cold",
+                                                  "CORT-Cold",
+                                                  "Control-Hot",
+                                                  "CORT-Hot")),
+        variable = factor(variable, levels = c("move", "shelter", "emergence"))) %>%
+  arrange(species, variable, treatment) %>%
+data.frame()
+write.csv(table_slopes_df, here("./output/checking/table_slopes.csv"))
+#
+# Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+table_slopes <- flextable(table_slopes_df) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Species"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 2, value = as_paragraph("Variable"), part = "header") %>% # Change the second column header
+  flextable::compose(i = 1, j = 3, value = as_paragraph("Treatment"), part = "header") %>% # Change the third column header
+  flextable::compose(i = 1, j = 5, value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fifth column header
+  flextable::compose(i = c(2:12,14:24), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = 1, j = 1, value = as_paragraph(as_i("L. delicata")), part = "body") %>% # Change the first row of the first column
+  flextable::compose(i = 13, j = 1, value = as_paragraph(as_i("L. guichenoti")), part = "body") %>% # Change the fourth row of the first column
+  flextable::compose(i = c(2:4,6:8,10:12,14:16,18:20,22:24), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
+  flextable::compose(i = c(1,13), j = 2, value = as_paragraph("Latency to move"), part = "body") %>% # Change the first row of the first column
+  flextable::compose(i = c(5,17), j = 2, value = as_paragraph("Latency to shelter"), part = "body") %>% # Change the fourth row of the first column
+  flextable::compose(i = c(9,21), j = 2, value = as_paragraph("Emergence"), part = "body") %>% # Change the seventh row of the first column
+  flextable::hline(i = c(4,8,16,20), j = c(2:5),part = "body") %>% # Add horizontal lines
+  flextable::hline(i = 12, part = "body") %>% # Add horizontal lines
+  bold(i = c(2,9,10,11,12), j = c(3:5), bold = TRUE, part = "body") %>% # Bold significant values
+  align(align = "center", j = c(4,5), part = "body") %>%
+  align(align = "center", j = c(1:5), part = "header") %>% 
+autofit()
+#
+table_slopes
+```
 
@@
 
-_Table S4.
-Contrasts for mass modelled for both species.
-The contrasts are between the slopes of the different treatments.
-Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
-In bold, the values that are significant at pMCMC < 0.05._
+_Table S4. Contrasts for mass modelled for both species. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
 
-**INSERT TABLE FOR MASS**
+```{r, table_contrasts_mass}
+#| label: table_contrasts_mass
+#
+source(here("R", "func.R"))
+#
+contrasts_mass <- bind_rows(deli = deli_mass_contrasts,
+                          guich = guich_mass_contrasts,
+                          .id = "Species") %>%
+                    mutate(across(everything(), as.character)) %>%
+
+                    rename_with(~ str_replace_all(., "X2\\.5\\.HDI_", "low_lim_")) %>%
+                    rename_with(~ str_replace_all(., "X97\\.5\\.HDI_", "up_lim_")) %>%
+
+                    pivot_longer(cols = -Species,
+                                names_to = c("type", "factor"),
+                                names_pattern = "(median|low_lim|up_lim)_(.*)") %>%
+
+                    pivot_wider(names_from = c(type),
+                                values_from = value) %>%
+
+                    mutate(CI = paste0("[", low_lim, ", ", up_lim, "]")) %>%
+
+                    dplyr::select(Species, factor, median, CI) %>%
+data.frame()
+write.csv(contrasts_mass, here("./output/checking/contrasts_mass.csv"))
+#
+# Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+table_mass <- flextable(contrasts_mass) %>%
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 2, value = as_paragraph("Predictor"), part = "header") %>% # Change the second column header
+  flextable::compose(i = 1, j = 3, value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
+  flextable::compose(i = 1, j = 4, value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
+  flextable::compose(i = c(2,3,5,6), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = 1, j = 1, value = as_paragraph(as_i("L. delicata")), part = "body") %>% # Change the first row of the first column
+  flextable::compose(i = 4, j = 1, value = as_paragraph(as_i("L. guichenoti")), part = "body") %>% # Change the fourth row of the first column
+  flextable::hline(i = 3, part = "body") %>% # Add horizontal lines
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>% 
+autofit()
+#
+table_mass
+```
+
+Contrasts were done by:  
+    - _Temperature_: β~Hot~ - β~Cold~  
+    - _CORT_: β~CORT~ - β~Control~  
+    - _Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
 
@@
 #
-deli_post <- dplyr::bind_rows(posterior_list$deli, .id = "variable")
-guich_post <- dplyr::bind_rows(posterior_list$guich, .id = "variable")
+deli_post <- dplyr::bind_rows(posterior_list$deli, .id = "variable") %>%
+  mutate(CI = paste0("[", low_lim, ", ", up_lim, "]")) %>%                  
+  dplyr::select(variable, Predictor, Median, CI) %>%
+data.frame()
+#
+guich_post <- dplyr::bind_rows(posterior_list$guich, .id = "variable") %>%
+  mutate(CI = paste0("[", low_lim, ", ", up_lim, "]")) %>%                  
+  dplyr::select(variable, Predictor, Median, CI) %>%
+data.frame()
 ```
 
-_Table S5.
-Estimates of each predictor for each variable modelled for L.
-delicata.
-Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
-In bold, the values that are significant at pMCMC < 0.05._
+_Table S5. Estimates of each predictor for each variable modelled for L. delicata. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
   flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
-  align(align = "center", j = c(3:6), part = "body") %>%
-  align(align = "center", j = c(1:6), part = "header") %>%
-  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>%
   flextable::compose(i = c(2:8,10:16,18:24,26:29), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
+  bold(i = c(1,3,7,9,17,18,25,26), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
   autofit()
@@
 
-_Table S6.
-Estimates of each predictor for each variable modelled for L.
-delicata.
-Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0.
-In bold, the values that are significant at pMCMC < 0.05._
+_Table S6. Estimates of each predictor for each variable modelled for L. guichenoti. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
   flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
-  align(align = "center", j = c(3:6), part = "body") %>%
-  align(align = "center", j = c(1:6), part = "header") %>% 
-  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Mean", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>% 
   flextable::compose(i = c(2:8,10:16,18:24,26:29), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
+  bold(i = c(1,9,23,24,25,26), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
   autofit()

### 2025-07-08 R/func.r

@@ -1,6 +1,7 @@
 #################### 
 ####################
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest,
+broom, tidybayes, ggh4x, HDInterval)
 #
 ####################
 ####################
@@ -160,10 +161,9 @@ tidy_post <- function(df) {
     pivot_longer(everything(), names_to = "Predictor", values_to = "Estimate") %>%
     group_by(Predictor) %>%
     summarise(
-      Mean = format_dec(mean(Estimate), 3),
-      CI_lower = format_dec(quantile(Estimate, 0.05, na.rm = TRUE), 3),
-      CI_upper = format_dec(quantile(Estimate, 0.95, na.rm = TRUE), 3),
-      pMCMC = format_dec(pmcmc(Estimate, null = 0, twotail = TRUE), 3),
+      Median = format_dec(median(Estimate), 3),
+      low_lim = format_dec(hdi(Estimate, credMass = 0.95)[1], 3),
+      up_lim = format_dec(hdi(Estimate, credMass = 0.95)[2], 3),
       .groups = "drop"
     ) %>%
   data.frame()
@@ -222,84 +222,112 @@ contrasts_func <- function(sp, res){
   df <- get(paste0(sp, "_", res, "_posteriors"))
   if(res == "mass"){
     # Contrasts
-    Temperature_int <- format_dec(mean(c(df$int_CORT_Hot, df$int_Control_Hot)) 
-                                  - mean(c(df$int_CORT_Cold, df$int_Control_Cold)), 3)
+    Temperature_int <- format_dec(median(c(df$int_CORT_Hot, df$int_Control_Hot)) 
+                                  - median(c(df$int_CORT_Cold, df$int_Control_Cold)), 3)
 
-    pMCMC_temp_int <- format_p(pmcmc(c(df$int_CORT_Hot, df$int_Control_Hot) 
-                                    - c(df$int_CORT_Cold, df$int_Control_Cold)), 3, equal = FALSE)
+    lowlim_temp_int <- format_dec(hdi(c(df$int_CORT_Hot, df$int_Control_Hot) 
[…]

### 2025-07-09 ms.qmd
updates

@@
 
-Animals constantly face events that can disrupt their homeostasis [**REFs**].
-These challenges can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [**REFs**].
-Regardless of the nature of the challenge, organisms must adjust to these alterations to maintain their physiological and behavioural integrity [**REFs**].
-In vertebrates, the ability to cope with allostasis is mediated by the hypothalamic-pituitary-adrenal (HPA) axis [**REF**] - the hypothalamic-pituitary-interrenal (HPI) axis in non-mammalians [**REFs**].
-This system elicits a physiological response that results in the secretion of glucocorticoids (GCs) [**REF**].
-Glucocorticoids are a group of hormones that help reallocate energetic resources to restore homeostasis following disruption [**REFs**].
-For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [**REFs**].
+Animals constantly face events that can disrupt their homeostasis.
+These challenges can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [@sapolsky_how_2000; @mcewen_concept_2003].
+Regardless of the nature of the challenge, organisms must adjust to these alterations to maintain their physiological and behavioural integrity [@wingfield2002endocrine].
+In vertebrates, the ability to cope with allostasis is mediated by the hypothalamic-pituitary-adrenal/interrenal (HPA/HPI) axis [@sapolsky_how_2000].
+This system elicits a physiological response that results in the secretion of glucocorticoids (GCs) [@sapolsky_how_2000].
+Glucocorticoids are a group of hormones that help reallocate energetic resources to restore homeostasis following disruption [@sapolsky_how_2000; @mcewen_concept_2003].
+For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [@trompeter2011invader].
 
 In the short term, GCs help organisms cope with acute challenges and return to homeostasis.
-However, long-term exposure to GCs can carry significant metabolic costs [**REFs**].
-Prolonged elevations in GCs can lead to immune suppression, impaired growth, or permanent neural alterations [**REFs**].
-From a behavioural perspective, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [**REFs**].
+However, long-term exposure to GCs can carry significant metabolic costs [@picard_mitochondria_2014].
+Prolonged elevations in GCs can lead to immune suppression, impaired growth, or permanent neural alterations [@mcewen2017neurobiological].
+From a behavioural perspective, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [@belliure2004behavioral; @martin2024blind].
 As such, animals are predicted to evolve mechanisms to appropriately respond to acute stressors while avoiding the costs in the long-term.
-For example, by regulating their responses to homotypic (repeated and similar) stressors, particularly those that are predictable and non-lethal [**REFs**; @grissom2009habituation].
+For example, by regulating their responses to homotypic (repeated and similar) stressors, particularly those that are predictable and non-lethal [@grissom2009habituation].
 Habituation - the reduction of physiological responses elicited by exposure to a homotypic stressor [@pfister1979glucocorticosterone] - can be crucial to avoid the long-term costs of chronic stress without compromising the response to acute stressors.
@@
 One of the key mediators of early-life effects is GC exposure.
-Elevations in GCs during development exert potent effects on HPA axis function, altering the response to stressors later in life [@vargas2016early; @zito2017early; @crino2024eggs].
+Elevations in GCs during development exert potent effects on HPA axis function, altering the response to stressors later in life [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs].
 For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
-In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting cognitive processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004].
+In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting cognitive processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta].
 Therefore, early-life exposure to GCs can have long-lasting effects on the response to both acute and chronic stressors, potentially leading to maladaptive responses in adulthood.
@@
 Although these differences are believed to be driven by size rather than direct endocrine effects [@crino2024eggs], they can still have important long-term consequences for how individuals respond to acute and chronic stressors.
-Importantly, incubation temperature and prenatal CORT may not act independently [see @Crino_2023], and their interaction could produce complex, trait-specific outcomes that shape both physiological and behavioural responses to stress.
+Importantly, incubation temperature and prenatal CORT may not act independently [see @wingfield_comparative_2008; @Crino_2023], and their interaction could produce complex, trait-specific outcomes that shape both physiological and behavioural responses to stress.
 Yet, the interactive effects of these early-life conditions on stress responses remain largely unexplored.
@@
 guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Invassivenes does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [**REFpers**], which may be linked to differences in their stress response [@koolhaas1999coping].
+Invassivenes does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their stress response [@koolhaas1999coping].
 
@@
 
-To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack once daily over multiple days [**insert relevant reference here**].
+To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack - a standardized processive stressor [@thaker2009acute; @trompeter2011invader] - once daily over multiple days.
 Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the shelter.
@@
 We used the posterior distributions of the intercepts and slopes to test for differences between species in their response to an acute stressor (intercept) and a chronic stressor (slope).
+95% HPDI were used to test the hypothesis that the contrasts differed from zero.
 
@@
 We started with 96 lizards, 48 per species and 12 per treatment per species.
-However, our final sample size was 84 due to natural mortality (n = 10).
+However, our final sample size was 85 due to natural mortality (n = 9).
 The final sample sizes per treatment and species are listed in @fig-results_beh_deli, @fig-results_beh_guich, and @fig-results_mass.

### 2025-07-09 ms.qmd
updates

@@
 title: "Effects of prenatal CORT and temperature on the response to a stressor"
+format:
   docx: 
-  number_sections: true
-  linestretch: 2
-  fontsize: 12pt
-  mainfont: Times New Roman
+    reference-doc: "./bib/tmpl.docx"
 bibliography: "./bib/refs.bib"
 csl: "./bib/animal-behaviour.csl"
-reference-doc: "./bib/tmpl.docx"
 execute:
@@
 Animals constantly face events that can disrupt their homeostasis.
-These challenges can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [@sapolsky_how_2000; @mcewen_concept_2003].
-Regardless of the nature of the challenge, organisms must adjust to these alterations to maintain their physiological and behavioural integrity [@wingfield2002endocrine].
-In vertebrates, the ability to cope with allostasis is mediated by the hypothalamic-pituitary-adrenal/interrenal (HPA/HPI) axis [@sapolsky_how_2000].
-This system elicits a physiological response that results in the secretion of glucocorticoids (GCs) [@sapolsky_how_2000].
-Glucocorticoids are a group of hormones that help reallocate energetic resources to restore homeostasis following disruption [@sapolsky_how_2000; @mcewen_concept_2003].
+These disruptions can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [@sapolsky_how_2000; @mcewen_concept_2003].
+Regardless of the nature of the challenge, organisms must adjust to allostasis to maintain their physiological and behavioural integrity [@wingfield2002endocrine].
+In vertebrates, the ability to cope with disruptions in homeostasis is carried out by the hypothalamic-pituitary-adrenal/interrenal (HPA/HPI) axis [@sapolsky_how_2000].
+This system elicits a physiological response, also known as the stress response, mediated by glucocorticoids (GCs), a group of hormones that facilitate the reallocation of energetic resources to restore homeostasis [@sapolsky_how_2000; @mcewen_concept_2003].
 For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [@trompeter2011invader].
-
-In the short term, GCs help organisms cope with acute challenges and return to homeostasis.
+In the short term, GCs help organisms cope with acute challenges.
 However, long-term exposure to GCs can carry significant metabolic costs [@picard_mitochondria_2014].
 Prolonged elevations in GCs can lead to immune suppression, impaired growth, or permanent neural alterations [@mcewen2017neurobiological].
-From a behavioural perspective, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [@belliure2004behavioral; @martin2024blind].
-As such, animals are predicted to evolve mechanisms to appropriately respond to acute stressors while avoiding the costs in the long-term.
+Furthermore, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [@belliure2004behavioral; @martin2024blind].
+As such, animals are predicted to evolve mechanisms to appropriately respond to acute stressors while avoiding the costs in the long term.
 For example, by regulating their responses to homotypic (repeated and similar) stressors, particularly those that are predictable and non-lethal [@grissom2009habituation].
@@
 How both types of habituation interplay with each other depends on the nature of the stressor.
-Physiological habituation does not occur when animals are faced with systemic stressors like hypoglycemia or ether exposure whereas habituation to processive stressors — those involving primarily psychological qualities, such as restraint or simulated predation - is frequent [@grissom2009habituation].
-However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeteadly exposed to a homotypic processive stressor [@jaferi2006corticosterone; @grissom2009habituation].
+Physiological habituation does not occur when animals are faced with systemic stressors like hypoglycemia or ether exposure, whereas habituation to processive stressors — those involving primarily psychological qualities, such as restraint or simulated predation - is frequent [@grissom2009habituation].
+However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeatedly exposed to a homotypic processive stressor [@jaferi2006corticosterone; @grissom2009habituation].
 Adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
-While the HPA/HPI axis is paramount for physiological habituation, cognitive habituation also involve neural mechanisms related to sensory processing, learning, and behavioural control such as the limbic system, the superior colliculus, or the optic tectum [@grissom2009habituation; @dutta2014saliency].
+While the HPA/HPI axis is paramount for physiological habituation, cognitive habituation also involves neural mechanisms related to sensory processing, learning, and behavioural control, such as the limbic system, the superior colliculus, or the optic tectum [@grissom2009habituation; @dutta2014saliency].
 Habituation to processive stressors is therefore likely to involve complex interactions between physiological and cognitive processes.
@@
 Elevations in GCs during development exert potent effects on HPA axis function, altering the response to stressors later in life [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs].
-For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
+For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post-hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
 In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting cognitive processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta].
@@
 In addition, early thermal environment can also affect the development of the HPA/HPI axis [@crino_under_2020; @crino2024eggs].
-For example, in the delicata skink (_Lampropholis delicata_) individuals incubated at cooler temperatures showed increased baseline CORT concentration in the blood [@crino2024eggs].
+For example, in the delicata skink (_Lampropholis delicata_), individuals incubated at cooler temperatures showed increased baseline CORT concentration in the blood [@crino2024eggs].
 Although these differences are believed to be driven by size rather than direct endocrine effects [@crino2024eggs], they can still have important long-term consequences for how individuals respond to acute and chronic stressors.
@@
 However, _L.
-delicata_ has been reported as a successful invader in New Zealand and Hawaii, while there are not such reports for _L.
+delicata_ has been reported as a successful invader in New Zealand and Hawaii, while there are no such reports for _L.
 guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Invassivenes does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their stress response [@koolhaas1999coping].
-
-### Subjects treatments, collection, and housing
+Invassiveness does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their stress response [@koolhaas1999coping].
+
+### Collection and housing
 
@@
 40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
-The crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperature was set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures at 34 ºC.
+The crickets were dusted with calcium weekly and multivitamins and calcium biweekly.
+Room temperature was set to 22-24 ºC, but we also provided the enclosures with a heat cord and a heat lamp following a 12 h light:12 h dark cycle, keeping the warm side of the enclosures at 34 ºC.
 
 *Eggs collection and incubation* - Eggs were collected between mid-October 2022 to the end of February 2023.
-We placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures to provide females with a place to lay the eggs.
+We placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite on one side of the communal enclosures to provide females with a place to lay the eggs.
 These boxes were checked three days a week.
-After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighted the eggs with a (OHAUS, Model spx123) digital scale with an accuracy of ± 0.001 g error.
+After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighed the eggs with a (OHAUS, Model spx123) digital scale with an accuracy of ± 0.001 g error.
 Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
@@
 C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
+We selected doses based on our previous study, where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
 Mean incubation temperatures represent the lower and upper limits of the natural range of nest temperatures in _L.
@@
 Both before and after the completion of the behavioural tests, lizard mass was measured to the nearest 0.001 g using a digital scale.
-Change in mass was calculated as the difference between final minus the initial mass and then rescaled by adding the smallest mass change observed in the dataset, to avoid negative values.
+Change in mass was calculated as the difference between the final minus the initial mass and then rescaled by adding the smallest mass change observed in the dataset, to avoid negative values.
 This rescaled change in mass was used as a proxy for the physiological cost of repeated stress exposure.
@@
 #| fig.cap: "Scheme of our experimental design.
-In panel A, the early environment manipulations resulting in four experimental conditions.
+In panel A, the early environment manipulations leading to four experimental conditions.
 In panel B, the timeline of the postnatal stress-response assay.
@@
 From Days 1–8, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
-Variables measured included freezing duration (latency to move), latency to shelter, and emergence from shelter.
+Variables measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor.
 Final mass was measured on Day 9."
@@
 
-We used as the response variables the latency to move, latency to shelter, and emergence from the shelter after the simulated attack, as well as the change in mass after the eight days of stress exposure.
+We used as the response variables the latency to move, the latency to shelter, and the emergence probability, as well as the change in mass after the eight days of stress exposure.
 We fitted separate models for each response variable.
 For the three behavioural variables, we included the prenatal condition (Control-Cold, CORT-Cold, Control-Hot, CORT-Hot), the day of the trial, and their interaction as fixed effects.
-To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e.
-the intercept reflects lizard behaviour in response to the first exposure to the stressor).
+To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e., the intercept reflects lizard behaviour in response to the first exposure to the stressor).
 For the variable mass, we included the prenatal condition, the total number of crickets ingested during the eight days of stress exposure, and their interaction as predictors.
@@
 Additionally, previous research has shown that clutches are generally sired by a single male, but sperm storage can occur [@kar2023heritability].
-Given our partial split-clutch design, and the fact that maternal effects are expected to be stronger than paternal effects in these species, including the clutch as a random factor should account for the effects of parental identity.
+Given our partial split-clutch design and the fact that maternal effects are expected to be stronger than paternal effects in these species, including the clutch as a random factor should account for the effects of parental identity.
 
 We used the posterior distributions of model parameters to test for differences between prenatal conditions.
-For behavioural variables, differences in the model intercepts between conditions were interpreted as differences in the acute response to a stressor (i.e.
-the first day of exposure), while differences in the slopes were interpreted as differences in behavioural adjustment across repeated exposures (i.e.
-chronic stress response).
-For mass change, between treatments difference reflect differences in energetic cost of the chronic stressor.
+For behavioural variables, differences in the model intercepts between conditions were interpreted as differences in the acute response to a stressor (i.e., the first day of exposure), while differences in the slopes were interpreted as differences in behavioural adjustment across repeated exposures (i.e., chronic stress response).
+For mass change, between-treatments differences reflect differences in the energetic cost of the chronic stressor.
 We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments or the slopes for each treatment were different from zero.
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r deli_move_contrasts$median_Temperature_int`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r format_dec(exp(deli_move_contrasts$median_Temperature_int), 3)`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
 We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
@@
 ## Discussion    
+Our study shows that early-life conditions can influence how skinks respond to a processive stressor, but the effects differ between species, the nature of th stressor, and the behaviour evaluated.
+_Lampropholis delicata_ incubated at warmer temperatures were less reactive to a simulated predator attack in the first trial, suggesting that early thermal conditions can shape the behavioural response to an acute stressor.
+In contrast, _L.
+guichenoti_ showed no treatment effects.
+Habituation patterns also differed.
+_Lampropholis delicata_ increased the probability of emerging from the shelter, suggesting some habituation to the stressor.
+However, none of the treatments showed habituation in latency to move or go to shelter.
+No behavioural change was detected in _L.
+guichenoti_ either.
+Finally, we found no evidence of physiological costs: individuals generally maintained or gained body mass, regardless of treatment.
+Together, these findings suggest that incubation temperature and prenatal hormone exposure shape stress-related behaviour in a species-specific manner, potentially reflecting differences in flexibility or stress physiology.
+
+#### Cold incubation temperature increases the response to an acute stressor in _L. delicata_ but not in _L. guichenoti_          
+Our results demonstrate that incubation temperature modulates the behavioural response to an acute stressor in _L.
+delicata_ but not in _L.
+guichenoti_.
+Hot-incubated _L.
+delicata_ spent less time immobile following a simulated predatory attack, suggesting a faster recovery from acute stress.
+**This result is consistent with the idea that shorter immobility duration (i.e., shorter freezing time) reflects a less reactive behavioural response, often interpreted as reduced perceived risk or more efficient coping [REFs].** Thus, the reduced latency to move of hot-incubated individuals may indicate reduced stress reactivity.
+In _L.
+delicata_, individuals incubated at 23 ºC exhibit elevated baseline CORT [@crino2024eggs], potentially enhancing sensitivity to stress.
+Therefore, the heightened acute response in cold-incubated individuals may be explained by elevated endogenous CORT levels.
+For example, **E.G.
+about CORT and response**.
+
+In contrast, early thermal environment had no effect on _L.
+guichenoti_'s response to an acute stressor.
+The species-specific response also warrants discussion.
+L.
+delicata exhibited treatment-dependent variation in response to acute stress, while L.
+guichenoti did not.
+This difference may reflect distinct developmental plasticity or stress-coping strategies between species.
+Given that L.
+delicata is a successful invasive species, greater sensitivity to early-life thermal environments and a more flexible behavioural response to acute threats may confer adaptive advantages in novel environments [REFs].
+In contrast, the more canalised stress response in L.
+guichenoti could reflect reduced behavioural plasticity, potentially limiting its adaptability in changing environments.
+
+Strikingly, we found no effects of prenatal CORT treatment on the acute stress response in either species.
+Glucocorticoid elevation during development are known to modulate stress-related phenotypes, including behavioural responsiveness to processive stressors [**REF...before??
+I might have another in the intro**].
+Previous studies have shown that
+
+#### Habituation to stressors varies between species and treatments
+Across repeated exposures to the simulated predatory attack, only L.
+delicata individuals from the CORT-Cold treatment increased their latency to move.
+Rather than indicating habituation, this pattern may reflect accumulating allostatic load.
+Habituation is typically inferred from a reduction in stress-related behaviours over time (e.g., decreased latency to resume activity), whereas an increase in latency could reflect mounting stress or impaired coping [@grissom2009habituation].
+Thus, the behavioural trajectory of the CORT-Cold group may suggest that these animals became less, rather than more, tolerant to the repeated stressor.
+
+Why did this pattern emerge only in the CORT-Cold L.
+delicata group?
+While our CORT treatment did not significantly elevate baseline hormone levels across groups, previous findings indicate that cold incubation alone raises baseline CORT [@crino2024eggs].
+The combination of prenatal CORT and cold temperature may therefore have produced a cumulative or synergistic effect on HPI axis development and reactivity.
+This effect might not have been detectable in our acute response measures, but could become apparent over repeated exposures as stress accumulates.
+
+Alternatively, cognitive explanations are also plausible.
+The cold-CORT combination may impair neural development in regions critical for behavioural flexibility and non-associative learning.
+Although our prior work found no significant impact of this combination on associative learning (e.g., colour discrimination and reversal learning; @recio2025cognitive; @recio2025early), cognitive processes underpinning habituation may rely on different neural substrates.
+The hippocampus and prefrontal cortex, implicated in behavioural inhibition and decision-making, may be differentially affected by early-life CORT and temperature exposure.
+Region-specific developmental trajectories could account for divergent effects across learning types.
+
+In contrast, L.
+guichenoti showed no evidence of behavioural change across repeated stress exposures, regardless of treatment.
+This further supports the notion of reduced behavioural flexibility or a more rigid stress-response phenotype in this species.
+It is noteworthy that L.
+delicata displayed treatment-specific changes even when these did not conform to classical habituation.
+Such responsiveness may reflect a greater capacity to modulate behaviour based on environmental cues, a trait often associated with invasive potential [REFs].
+
+Interestingly, across all treatments, L.
+delicata increased their probability of emerging from shelter within 40 minutes of the stressor as the experiment progressed, suggesting that overall risk perception decreased over time.
+This pattern is consistent with habituation at the level of decision-making rather than immediate post-threat responses (i.e., latency to move).
+L.
+guichenoti, in contrast, exhibited no such trend.
+This species difference in shelter-emergence behaviour provides further support for differential flexibility in stress evaluation and risk assessment.
+The divergence may again be rooted in differences in developmental plasticity or evolutionary history, with L.
+delicata potentially possessing a more labile stress-response system suited to fluctuating environments.
+
+The absence of time effects in latency to move or go to shelter for the other treatments in L.
+delicata suggests that habituation was limited and potentially context- or trait-dependent.
+Taken together, our findings highlight the complexity of stress habituation and the need to disentangle multiple behavioural and physiological components to fully understand coping strategies.
+
+#### Physiological costs of the stressor were low and did not differ between treatments or species
+Despite repeated exposure to a stressor over several days, we found no clear evidence of physiological costs in either species.
+Most individuals gained mass over the course of the experiment, and there were no treatment or species differences in mass change.
+This suggests that the imposed stressor, while sufficient to elicit behavioural responses, did not impair growth or energy balance under our conditions.
+This result is somewhat surprising, given the literature linking chronic stress to impaired growth, particularly in juvenile ectotherms [REFs].
+
+One possible explanation is that the stressor we employed, while ecologically relevant, was not sufficiently intense or frequent to produce measurable physiological effects.
+In natural settings, stressors may be more variable and compound with other challenges (e.g., food scarcity, thermal extremes), resulting in more pronounced physiological consequences.
+Our controlled environment, consistent food availability, and short stress exposure duration may have buffered such effects.
+
+Another consideration is that mass change may not fully capture the physiological costs of stress.
+Other endpoints, such as oxidative stress markers, telomere attrition, or immune function, may be more sensitive to sublethal chronic stress effects [REFs].
+Future studies incorporating these physiological markers would provide a more comprehensive picture of the costs associated with repeated stress.
+
+Finally, it is important to note that our behavioural findings suggest differences in stress evaluation and habituation across treatments and species, even in the absence of overt physiological changes.
+This dissociation underscores the importance of measuring multiple dimensions of the stress response and not relying solely on growth or condition as proxies for stress burden.
+
+#### Conclusion
+Our study demonstrates that early-life environments—specifically incubation temperature and prenatal glucocorticoid exposure—shape behavioural responses to acute and repeated stressors in skinks.
+These effects were species-specific, with L.
+delicata showing greater responsiveness to early-life conditions than L.
+guichenoti.
+Cold incubation increased acute stress reactivity in L.
+delicata, and the combination of cold and prenatal CORT impaired habituation, potentially via elevated baseline CORT or neural alterations.
+In contrast, L.
+guichenoti appeared less flexible in both acute and repeated stress responses.
+
+These findings contribute to our understanding of how early developmental environments shape adult stress phenotypes.
+They also highlight potential mechanisms underlying behavioural plasticity in invasive species, where adaptability to repeated stressors may facilitate persistence in novel environments.
+Future work should aim to link behavioural and physiological responses across ontogeny and explore neural mechanisms that underpin different forms of learning and habituation.
+
+#### Cold incubation temperature increases the response to an acute stressor in _L. delicata_ but not in _L. guichenoti_
+**CHECK PROPER UNITS**
+Hot-incubated _L.
+delicata_ spent less time frozen after the simulated attack.
+We did not see any differnces betwen treatments in any other variable recorded.
+No treatment effect on the response to an acute stressor in _L.
+guichenoti_.
+- Why do we consider less time to be less reactive?
+- Why differences in temp? --> Cold animals had higher baseline CORT levels...but CORT treated too...at least mean treatment but not high.
+- Why not differences based on CORT?
+- Why is temperature in _L. delicata_ more important for this than in _L. guichenoti_?? Invassiveness??
+
+#### Habituation to stressors varies between species and treatments
+Only CORT-Cold _L.
+delicata_ increase the latency to move over time, suggesting allostatic load more than habituation, as we would expect animals to resume their activity earlier with habituation.
+However, the analyses did not show significant differences between the contrasts.
+ - WHY those resuls? Power?
+ - Why CORT-Cold --> higher baseline CORT levels??
+CORT at the levels we used increases slightly baseline CORT blood levels in juveniles of _L.
+delicata_, this difference was not significant compared to the controls, but it could be enough for increasing at cold temps when CORT levels are already higher than at warm temps [@crino2024eggs].
+This seems to make CORT-treated _L.
+delicata_ incubated at cold temperatures as reactive to acute stress as Cold-Controls, but less prone to habituation.
+    Alternatively, this response could be due to lower cognitive abilities as expected by CORT + cold.
+However, this hasn't been seen in other learning abilities [@recio2025cognitive; @recio2025early].
+Nevertheless, the effects of temperature and CORT are expectd to be region-dependent, so there could be effects in non-associative learning while not in associative types of learning (i.e., colour association or its reveresal)
+ - Why not in _L. guichenoti_??
+The latency to move and to go to shelter did not change over time in any other treatment in _L. delicata_.
+All _L. delicata_ increase the probability of emerging in 40 min after the attack, suggesting that the risk perception decreased during the experiment. This did not happen in _L. guichenoti_   
+  - Risk perception seems the same both at the beginning and during the experiment for all treatments despite the frozen behaviour being higher in Cold-incubated lizards. Why??
+  - Why differences between species?? Is _L. delicata_ more flexible in evaluating risks?? --> Relationship with invasiveness
+
+  **CHECK TABLE FOR SPP COMPARISSONS**
+
+#### Physiological costs of the stressor were low and did not differ between treatments or species
+There were no "physiological costs"
+- Most of the individuals grew up...How's the growth compared to other studies where they were not subjected to a stressor?? What are the usual "costs"??
+- Caveats of the methods: other costs
 

### 2025-07-09 R/func.r

@@ -219,7 +219,18 @@ post_values <- function(df, var){
 #' @param res To select the response variable ("move", "shelter", "emergence", "mass")
 contrasts_func <- function(sp, res){
   data_table <- data.frame()
-  df <- get(paste0(sp, "_", res, "_posteriors"))
+  # Getting the df using real data (depending on the distribution of the response variable)
+  if(res == "emergence"){
+    df <- get(paste0(sp, "_", res, "_posteriors")) %>%
+          mutate(across(starts_with("int_"), ~ plogis(.x))) # Convert log-odds to probabilities
+      } else if(res == "mass"){
+    df <- get(paste0(sp, "_", res, "_posteriors"))  %>%
+          mutate(across(everything(), ~ exp(.x) + min(mass_clean$delta_mass, na.rm = TRUE) - 1))
+          # Convert log-transformed rescaled values back to original non-scaled values
+      } else {
+    df <- get(paste0(sp, "_", res, "_posteriors"))  %>%
+          mutate(across(starts_with("int_"), ~ exp(.x)))  # Convert log-transformed values back to original scale
+      }
   if(res == "mass"){
     # Contrasts
     Temperature_int <- format_dec(median(c(df$int_CORT_Hot, df$int_Control_Hot)) 

### 2025-07-10 ms.qmd
updates

@@
 In the short term, GCs help organisms cope with acute challenges.
-However, long-term exposure to GCs can carry significant metabolic costs [@picard_mitochondria_2014].
-Prolonged elevations in GCs can lead to immune suppression, impaired growth, or permanent neural alterations [@mcewen2017neurobiological].
+However, long-term exposure to GCs can carry significant metabolic costs, leading to immune suppression, impaired growth, or permanent neural alterations [@picard_mitochondria_2014; @mcewen2017neurobiological].
 Furthermore, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [@belliure2004behavioral; @martin2024blind].
@@
 Importantly, the process of physiological habituation shares key functional characteristics with the cognitive process of habituation - a form of non-associative learning where individuals reduce their responsiveness to a repeated stimulus [@thompson1966habituation].
-In both cases, organisms are repeatedly exposed to a stimulus/homotypic stressor and progressively decrease their response towards it [@grissom2009habituation].
+In both cases, organisms are repeatedly exposed to a stimulus/homotypic stressor and progressively decrease their response toward it [@grissom2009habituation].
 In fact, both behavioural and physiological habituation often co-occur.
-For example, rhesus macaques (_Macaca mulatta_) showed a decline in blood cortisol concentrations across consecutive days of restraint together with decreased vocalizations and behavioral agitation [@ruys2004behavioral].
+For example, rhesus macaques (_Macaca mulatta_) showed a decline in blood cortisol concentrations across consecutive days of restraint, accompanied by decreased vocalizations and behavioral agitation [@ruys2004behavioral].
 How both types of habituation interplay with each other depends on the nature of the stressor.
@@
 However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeatedly exposed to a homotypic processive stressor [@jaferi2006corticosterone; @grissom2009habituation].
-Adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
+For example, adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
 While the HPA/HPI axis is paramount for physiological habituation, cognitive habituation also involves neural mechanisms related to sensory processing, learning, and behavioural control, such as the limbic system, the superior colliculus, or the optic tectum [@grissom2009habituation; @dutta2014saliency].
@@
 One of the key mediators of early-life effects is GC exposure.
-Elevations in GCs during development exert potent effects on HPA axis function, altering the response to stressors later in life [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs].
+Elevations in GCs during development exert potent effects on HPA/HPI axis function, altering the response to stressors later in life [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs].
 For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post-hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
-In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting cognitive processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta].
+In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting neural processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta].
 Therefore, early-life exposure to GCs can have long-lasting effects on the response to both acute and chronic stressors, potentially leading to maladaptive responses in adulthood.
@@
 guichenoti_).
-We focus on the effects of these early-life conditions on the behavioural responses to a simulated predatory attack.
-Specifically, we examine how these factors influence the acute response to a stressor (i.e., the first exposure) and the chronic response (i.e., repeated exposures over time).
+We focus on the effects of these early-life conditions on the behavioural responses to a processive stressor.
+Specifically, we examine how these factors influence the acute (i.e., the first exposure) and the chronic response (i.e., repeated exposures over time) responses to a simulated predatory attack.
 Finally, we assessed body mass change across the stress period as a potential indirect cost, while controlling for food intake.
@@
 guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Invassiveness does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their stress response [@koolhaas1999coping].
+Invassiveness does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their ability to respond to acute and chronic processive stressors [@koolhaas1999coping].
 
@@
 Enclosures were provided with shelter, nonstick matting, and several small water dishes.
-The lizards were given water daily and were fed approx.
-40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
+The lizards were given water daily and were fed approximately 40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
 The crickets were dusted with calcium weekly and multivitamins and calcium biweekly.
@@
 Nonetheless, the higher temperature treatment (28 ºC) was above the thermal optima estimated for _L.
-delicata_ [T~opt~ = 25 ºC; @pettersen2023maternal] and reached more stressful temperatures daily as fluctuations occurred (i.e., 31 ºC).
+delicata_ [T~opt~ = 25 ºC; @pettersen2023maternal].
 
@@
 Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the shelter.
-The lizard was then chased for 60 seconds using a soft paintbrush, simulating a predatory attack.
+The lizard was then chased for 60 seconds using a soft paintbrush, simulating the predatory attack.
 Immediately after the chase, the shelter — but not the water dish — was returned to its original position, marking the start of the behavioural observation period.
@@
 All videos were later analysed by MD, who was blinded to the lizards’ treatment.
-From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile (frozen) immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
+From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
 A lizard was considered to have resumed movement if it moved continuously for at least five seconds, and to be under shelter if no limbs were visible.
-This procedure was repeated daily for eight consecutive days to assess the effects of mid-term exposure to a stressor.
-All trials were conducted between 1000–1400 h, when lizards are most active.
+This procedure was repeated daily for eight consecutive days to assess the effects of long-term exposure to a processive stressor.
+All trials were conducted between 1000–1400 h, when lizards were most active.
 
 Both before and after the completion of the behavioural tests, lizard mass was measured to the nearest 0.001 g using a digital scale.
-Change in mass was calculated as the difference between the final minus the initial mass and then rescaled by adding the smallest mass change observed in the dataset, to avoid negative values.
-This rescaled change in mass was used as a proxy for the physiological cost of repeated stress exposure.
+Change in mass was calculated as the difference between the final minus the initial mass.
+Change in mass was used as a proxy for the metabolic cost of repeated stress exposure, where mass loss or lower growth was considered as higher costs.
 To control for hunger levels and potential effects of food ingestion on mass change, we also recorded the number of crickets ingested by each lizard during the acclimation period.
@@
 In panel A, the early environment manipulations leading to four experimental conditions.
-In panel B, the timeline of the postnatal stress-response assay.
-On Day 0, lizard mass was recorded.
-From Days 1–8, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
-Variables measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor.
-Final mass was measured on Day 9."
+In panel B, the timeline of the postnatal stress-response assay: on Day 0 and 9, lizard mass was recorded; from Days 1–8, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
+Variables measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor."
 
@@
 We performed the analyses for each species separately.
-We fitted all models using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.4.0 [@R].
+We fitted a total of eight models using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.4.0 [@R].
 We ran four parallel MCMC chains of 10000 iterations for each model, with a warmup period of 4000 iterations.
@@
 To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e., the intercept reflects lizard behaviour in response to the first exposure to the stressor).
-For the variable mass, we included the prenatal condition, the total number of crickets ingested during the eight days of stress exposure, and their interaction as predictors.
-
-Lizard identity was included as a random factor in the behavioural models, but it was not included when mass was the response.
+For the variable mass, we included the prenatal condition and the total number of crickets ingested per each individual during the eight days of stress exposure as predictors.
+
+We included each lizard's random intercept and slope (trial) as a random factor in the behavioural models, but not when mass was the response.
 For all models, we included the clutch as a random factor.
@@
 
+All the estimated parameters in the results were transformed back to original units employing appropriate formulas depending on the distribution employed in the models.
+The error structure for latency to move, latency to shelter, or mass change followed a lognormal distribution [family = lognormal()], while the probability of emerging from the shelter was modeled using a Bernoulli distribution with a logit link function [family = Bernoulli(link = 'logit')].
+Change in mass was also rescaled by adding the minimum change in mass recorded before running the models to avoid negative values but transformed back to original values for the presentation of the results.
+
 ## Results
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I ~Cold~ = `r format_dec(exp(deli_move_contrasts$median_Temperature_int), 3)`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
+_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I~Cold~ = `r deli_move_contrasts$median_Temperature_int`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
 We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
@@
 #| label: fig-results_mass
-#| fig.cap: "Change in mass of Lampropholis delicata and Lampropholis guichenoti after eight days of repeated stress exposure." 
+#| fig.cap: "Change in mass of _Lampropholis delicata_ and _Lampropholis guichenoti_ after eight days of repeated stress exposure." 
 #
@@
 
-_Acute stressor_: Latency to move after the first exposure in Control-Hot incubated _L.
-delicata_ was higher than for _L.
-guichenoti_ (I~L.
-deli~ - I~L.
-guich~ = `r format_dec(median(deli_move_posteriors$int_Control_Hot) - median(guich_move_posteriors$int_Control_Hot), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$int_Control_Hot - guich_move_posteriors$int_Control_Hot, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$int_Control_Hot - guich_move_posteriors$int_Control_Hot, credMass = 0.95)[2], 3)`]), but there were no differences in Latency to move between the rest of the treatments (see @table_contrasts_species).
-In contrast, the probability of emerging from the shelter after an acute stressor was lower in cold incubated _L.
-delicata_ than _L.
-guichenoti_ for both control (I~L.
-deli~ - I~L.
-guich~ = `r format_dec(median(deli_emergence_posteriors$int_Control_Cold) - median(guich_emergence_posteriors$int_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$int_Control_Cold - guich_emergence_posteriors$int_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$int_Control_Cold - guich_emergence_posteriors$int_Control_Cold, credMass = 0.95)[2], 3)`]) and CORT-treated (I~L.
-deli~ - I~L.
-guich~ = `r format_dec(median(deli_emergence_posteriors$int_CORT_Cold) - median(guich_emergence_posteriors$int_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$int_CORT_Cold - guich_emergence_posteriors$int_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$int_CORT_Cold - guich_emergence_posteriors$int_CORT_Cold, credMass = 0.95)[2], 3)`]) lizards, but not in hot incubated animals (see @table_contrasts_species).
-There were no differences between species in the latency until sheltering after an acute stressor (see @table_contrasts_species).
-
-_Chronic stressor_: Over time, _L.
-delicata_ CORT-treated lizards incubated at cold temperatures increased latency to move more than _L.
-guichenoti_ (β~L.
-deli~ - β~L.
-guich~ = `r format_dec(median(deli_move_posteriors$slope_CORT_Cold) - median(guich_move_posteriors$slope_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
-The probability of emerging from the shelter increased over time in _L.
-delicata_ more than in _L.
-guichenoti_ (see @table_contrasts_species), but those differences were significant only in the Control-Cold treatment (β~L.
-deli~ - β~L.
-guich~ = `r format_dec(median(deli_emergence_posteriors$slope_Control_Cold) - median(guich_emergence_posteriors$slope_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
-There were no significant differences in the changes on latency to shelter between species (see @table_contrasts_species).
-
-_Mass change_: increases in mass were not different between species in any of the treatments (see @table_contrasts_species).
-
-_Table 1.
-Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L.
-guichenoti.
-Contrast represent differences in median parameters between species (L.
-delicata - L.
-guichenoti) for each treatment.
-95% Highest Posterior Density Intervals (95% HPDIs) indicate the probability that contrasts differ from zero._
-
-```{r, table_contrasts_species}
-#| label: table_contrasts_species
-#| table.caption: "Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L. guichenoti."
-source(here("R", "func.R"))
+```{r, contrasts_sp}
+#| label: contrasts_sp
+# Chunk created to make the df for the table below (Table 1) and also facilitate the comparisons between species below
 #
@@
                 dplyr::select(column_id, draw, value) %>%
-                pivot_wider(names_from = column_id, values_from = value)
+                pivot_wider(names_from = column_id, values_from = value) %>%
+
+                #Get the real values
+                mutate(across(matches("emergence.*_int"), ~ plogis(.x)),
+                      across(matches("(move|shelter).*_int"), ~ exp(.x)),
+                      across(matches("mass.*_int"), ~ exp(.x) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
+data.frame()
+```
+
+_Acute stressor_: The probability of emerging from the shelter after an acute stressor was lower in cold incubated _L.
+delicata_ than _L.
+guichenoti_ for both control (I~delicata~ - I~guichenoti~ = `r format_dec(median(contrasts_sp$emergence_deli_int_Control_Cold) - median(contrasts_sp$emergence_guich_int_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(contrasts_sp$emergence_deli_int_Control_Cold - contrasts_sp$emergence_guich_int_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(contrasts_sp$emergence_deli_int_Control_Cold - contrasts_sp$emergence_guich_int_Control_Cold, credMass = 0.95)[2], 3)`]) and CORT-treated (I~delicata~ - I~guichenoti~ = `r format_dec(median(contrasts_sp$emergence_deli_int_CORT_Cold) - median(contrasts_sp$emergence_guich_int_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(contrasts_sp$emergence_deli_int_CORT_Cold - contrasts_sp$emergence_guich_int_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(contrasts_sp$emergence_deli_int_CORT_Cold - contrasts_sp$emergence_guich_int_CORT_Cold, credMass = 0.95)[2], 3)`]) lizards, but not in hot incubated animals (see @table_contrasts_species).
+There were no differences between species in the latency to move or the latency until sheltering after an acute stressor (see @table_contrasts_species).
+
+_Chronic stressor_: Over time, _L.
+delicata_ CORT-treated lizards incubated at cold temperatures increased latency to move more than _L.
+guichenoti_ (β~delicata~ - β~guichenoti~ = `r format_dec(median(deli_move_posteriors$slope_CORT_Cold) - median(guich_move_posteriors$slope_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
+The probability of emerging from the shelter increased over time in _L.
+delicata_ more than in _L.
+guichenoti_ (see @table_contrasts_species), but those differences were significant only in the Control-Cold treatment (β~delicata~ - β~guichenoti~ = `r format_dec(median(deli_emergence_posteriors$slope_Control_Cold) - median(guich_emergence_posteriors$slope_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
+There were no significant differences in the changes on latency to shelter between species (see @table_contrasts_species).
+
+_Mass change_: increases in mass were not different between species in any of the treatments (see @table_contrasts_species).
+
+_Table 1.
+Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L.
+guichenoti.
+Contrast represent differences in median parameters between species (L.
+delicata - L.
+guichenoti) for each treatment.
+95% Highest Posterior Density Intervals (95% HPDIs) indicate the probability that contrasts differ from zero._
+
+```{r, table_contrasts_species}
+#| label: table_contrasts_species
+#| table.caption: "Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L. guichenoti."
+source(here("R", "func.R"))
+#
 # Make loop to get contrasts
@@
 }
-# Modify the df to get it ready for the table
+# Modify the contrast_df (above) to get it ready for the table
 contrasts_df2 <- contrasts_df %>%
@@
 write.csv(contrasts_df2, here("./output/checking/contrasts_species.csv"))
-#
 # Set defaults for the table
@@
   flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move"), part = "body") %>% 
-  flextable::compose(i = 5, j = 1, value = as_paragraph("Latency to shelter"), part = "body") %>% 
+  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move (s)"), part = "body") %>% 
+  flextable::compose(i = 5, j = 1, value = as_paragraph("Latency to shelter (s)"), part = "body") %>% 
   flextable::compose(i = 9, j = 1, value = as_paragraph("Emergence"), part = "body") %>% 
-  flextable::compose(i = 13, j = 1, value = as_paragraph("Δmass"), part = "body") %>% # Change the first column's names
+  flextable::compose(i = 13, j = 1, value = as_paragraph("Δmass (mg)"), part = "body") %>% # Change the first column's names
   flextable::compose(i = c(13:16), j = c(5,6), value = as_paragraph("-"), part = "body") %>%
@@
   align(align = "center", i = c(1,2), j = c(1:6), part = "header") %>% 
-  flextable::bold(i = c(3,9,10), j = c(3,4), bold = TRUE, part = "body") %>% 
+  flextable::bold(i = c(9,10), j = c(3,4), bold = TRUE, part = "body") %>% 
   flextable::bold(i = c(2,9), j = c(5,6), bold = TRUE, part = "body") %>%

### 2025-07-13 ms.qmd
updates

@@
 In vertebrates, the ability to cope with disruptions in homeostasis is carried out by the hypothalamic-pituitary-adrenal/interrenal (HPA/HPI) axis [@sapolsky_how_2000].
-This system elicits a physiological response, also known as the stress response, mediated by glucocorticoids (GCs), a group of hormones that facilitate the reallocation of energetic resources to restore homeostasis [@sapolsky_how_2000; @mcewen_concept_2003].
+This system elicits a physiological response, also known as the stress response, mediated by glucocorticoids (GCs) - a group of hormones that facilitates the reallocation of energetic resources to restore homeostasis [@sapolsky_how_2000; @mcewen_concept_2003].
 For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [@trompeter2011invader].
@@
 For example, rhesus macaques (_Macaca mulatta_) showed a decline in blood cortisol concentrations across consecutive days of restraint, accompanied by decreased vocalizations and behavioral agitation [@ruys2004behavioral].
-How both types of habituation interplay with each other depends on the nature of the stressor.
-Physiological habituation does not occur when animals are faced with systemic stressors like hypoglycemia or ether exposure, whereas habituation to processive stressors — those involving primarily psychological qualities, such as restraint or simulated predation - is frequent [@grissom2009habituation].
-However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeatedly exposed to a homotypic processive stressor [@jaferi2006corticosterone; @grissom2009habituation].
+However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeatedly exposed to a processive stressors - those involving primarily psychological qualities, such as restraint or simulated predation  [@jaferi2006corticosterone; @grissom2009habituation].
 For example, adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
@@
 
-Crucially, the effectiveness of either system depends on how well the underlying neural structures develop and integrate.
+However, the effectiveness of either system depends on how well the underlying neural structures develop and integrate.
 Many of these brain regions mature concurrently and can be highly plastic during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017].
@@
 Importantly, incubation temperature and prenatal CORT may not act independently [see @wingfield_comparative_2008; @Crino_2023], and their interaction could produce complex, trait-specific outcomes that shape both physiological and behavioural responses to stress.
-Yet, the interactive effects of these early-life conditions on stress responses remain largely unexplored.
-
-In this study, we investigate how prenatal CORT and incubation temperature shape the response to a stressor in two species of skinks: the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
+
+Because of the functional similarities between physiological and cognitive habituation, behavioural responses can offer insight into an animal’s ability to regulate stress over time, even in the absence of physiological measures.
+This is particularly relevant when investigating early-life programming, which can shape both HPA/I development and brain regions involved in behavioural inhibition and learning [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta; @amiel_hotter_2012].
+In this study, we investigated how prenatal CORT and incubation temperature shape the response to a stressor in two species of skinks: the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
 guichenoti_).
-We focus on the effects of these early-life conditions on the behavioural responses to a processive stressor.
-Specifically, we examine how these factors influence the acute (i.e., the first exposure) and the chronic response (i.e., repeated exposures over time) responses to a simulated predatory attack.
+We focused on the effects of these early-life conditions on the behavioural responses to a processive stressor.
+Specifically, we examined how these factors influence the acute (i.e., the first exposure) and the chronic (i.e., repeated exposures over time) responses by measuring freezing behaviour, shelter seeking, and shelter emergence after a simulated predatory attack.
 Finally, we assessed body mass change across the stress period as a potential indirect cost, while controlling for food intake.
@@
 
-### Subjects
-
-We used two species of skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
+#### Subjects
+
+We used two species of closely skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
 guichenoti_).
-These are oviparous, generalist, small (∼35-55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia [@chapple_know_2011; @chapple_biology_2014].
-However, _L.
-delicata_ has been reported as a successful invader in New Zealand and Hawaii, while there are no such reports for _L.
+Both species are oviparous, generalist, small (∼35-55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia [@chapple_know_2011; @chapple_biology_2014].
+However, while _L.
+delicata_ has been reported as a successful invader in New Zealand and Hawaii, there are no such reports for _L.
 guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Invassiveness does not seem to be related to differences in cognitive abilities [@bezzina2014does; @recio2025early; @recio2025early], but seems to be associated with differences in personality [@chapple_know_2011], which may be linked to differences in their ability to respond to acute and chronic processive stressors [@koolhaas1999coping].
-
-### Collection and housing
+Differences in the ability to colonise new environments are not related to cognitive abilities [@bezzina2014does; @recio2025cognitive; @recio2025early], but could be associated with differences in personality [@chapple_know_2011], which may be linked to different responses to acute and chronic processive stressors [@koolhaas1999coping].
+
+#### Collection and housing
 
@@
 We selected doses based on our previous study, where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
-Mean incubation temperatures represent the lower and upper limits of the natural range of nest temperatures in _L.
-delicata_ [mean nesting temperatures = 27.4 ºC; @cheetham2011embryonic].
-Nonetheless, the higher temperature treatment (28 ºC) was above the thermal optima estimated for _L.
-delicata_ [T~opt~ = 25 ºC; @pettersen2023maternal].
+Mean incubation temperatures represent the lower and upper limits of the natural range of nest temperatures in both species [@qualls2000post; @cheetham2011embryonic].
 
@@
 
-To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack - a standardized processive stressor [@thaker2009acute; @trompeter2011invader] - once daily over multiple days.
+To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack - a standardized processive stressor [@thaker2009acute; @trompeter2011invader; @iglesias2024maternal] - once daily over multiple days.
 Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the shelter.
@@
 
-### Statistical analyses
+#### Statistical analyses
 
@@
 #| label: fig-results_beh_deli
-#| fig.cap: "Behavioural response of Lampropholis delicata to an stressor."
-
-#### Panels A and D show the predicted probability of choosing the correct ramp in the first trial (Decision first trial).
-In panels B and E, the distribution of the estimates of slopes per each treatment.
-In all A, B, D, and E the x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
+#| fig.cap: "Behavioural response of _Lampropholis delicata_ to a stressor.
+Panels A, D, and G show the distribution of the estimates for all the behavioural responses after the first simulated attack per each treatment (i.e., response to acute stressor).
+In panels B, E, and H the distribution of the estimates of slopes per each treatment (i.e., response to chronic stressor).
+In all A, B, D, E, G and H the x-axis represents the estimated values, and in the y-axis the density of the estimates.
 The different colors indicate the different treatments.
-Points and bars represent the mean and standard deviation of the mean of the estimates, respectively.
-Dashed lines indicate value 0.33 (the probability of choosing correctly by chance) in panels A, D, and 0 in panels B, E.
-Asterisks indicate significant results (p~mcmc~ < 0.05).
-Panels C and F show the predicted probability of choosing the correct ramp first over trials.
-The lines represent the mean predicted probability of choosing the correct ramp for each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained using the slope and intercept estimates from the posterior distributions.
-The different colors indicate the different treatments.
+Points and bars represent the median and 95% Highest Posterior Density Intervals (95% HPDI), respectively.
+Dashed lines in panels B, E, and H indicate values equal 0 (i.e.
+no response to chronic stress).
+Panels C, F, and I show the behavioural estimated values over trials.
+The lines represent the median estimated value for each trial, and the shaded areas indicate the 95% HPDI; both were obtained using the slope and intercept estimates from the posterior distributions.
+The different colors indicate the different treatments."
+#  
 source(here("R", "func.R"))
@@
 There were no significant differences in the behavioural response of _L.
-guichenoti_ to the repeated exposure to an stressor (see @fig-results_beh_guich B, C, E, F, H, I and Table S2 in _Supplementary Material_).
+guichenoti_ to the repeated exposure to a stressor (see @fig-results_beh_guich B, C, E, F, H, I and Table S2 in _Supplementary Material_).
 
@@
 #| label: fig-results_beh_guich
-#| fig.cap: "Behavioural response of Lampropholis guichenoti to an stressor." 
+#| fig.cap: "Behavioural response of _Lampropholis guichenoti_ to a stressor.
+Panels A, D, and G show the distribution of the estimates for all the behavioural responses after the first simulated attack per each treatment (i.e., response to acute stressor).
+In panels B, E, and H the distribution of the estimates of slopes per each treatment (i.e., response to chronic stressor).
+In all A, B, D, E, G and H the x-axis represents the estimated values, and in the y-axis the density of the estimates.
+The different colors indicate the different treatments.
+Points and bars represent the median and 95% Highest Posterior Density Intervals (95% HPDI), respectively.
+Dashed lines in panels B, E, and H indicate values equal 0 (i.e.
+no response to chronic stress).
+Panels C, F, and I show the behavioural estimated values over trials.
+The lines represent the median estimated value for each trial, and the shaded areas indicate the 95% HPDI; both were obtained using the slope and intercept estimates from the posterior distributions.
+The different colors indicate the different treatments."
 #
@@
 #| label: fig-results_mass
-#| fig.cap: "Change in mass of _Lampropholis delicata_ and _Lampropholis guichenoti_ after eight days of repeated stress exposure." 
+#| fig.cap: "Change in mass of _Lampropholis delicata_ (Panel A) and _Lampropholis guichenoti_ (Panel B)after eight days of repeated stress exposure." 
 #
@@
 
-#### Do both species differ in their response to an stressor?
+#### Do both species differ in their response to a stressor?
 
@@
 ## Discussion    
-Our study shows that early-life conditions can influence how skinks respond to a processive stressor, but the effects differ between species, the nature of th stressor, and the behaviour evaluated.
+We predicted that elevated CORT levels in eggs or incubating them at colder temperatures would increase lizards reactivity to a stressor both in the first encounter or during repeated exposure.
+Higher reactivity would imply larger latencies to move after the simulated attack, lower times until sheltering, and lower probabilities of emerging from the shelter.
+
+Our study shows that early-life conditions can influence how skinks respond to a processive stressor, but the effects differ between species, the nature of the stressor, and the behaviour evaluated.
 _Lampropholis delicata_ incubated at warmer temperatures were less reactive to a simulated predator attack in the first trial, suggesting that early thermal conditions can shape the behavioural response to an acute stressor.
@@
 #### Cold incubation temperature increases the response to an acute stressor in _L. delicata_ but not in _L. guichenoti_    
-Our results demonstrate that incubation temperature modulates the behavioural response to an acute stressor in _L.
+We found that the early-life conditions can impact lizards' response to an acute stressor, but this effects were species- and behaviour-dependent.
+Our results demonstrate that incubation temperature can modulate the behavioural response to an acute stressor in _L.
 delicata_ but not in _L.
@@
 delicata_ spent less time immobile following a simulated predatory attack, suggesting a faster recovery from acute stress.
-**This result is consistent with the idea that shorter immobility duration (i.e., shorter freezing time) reflects a less reactive behavioural response, often interpreted as reduced perceived risk or more efficient coping [REFs].** Thus, the reduced latency to move of hot-incubated individuals may indicate reduced stress reactivity.
+This result is consistent with the idea that shorter immobility duration (i.e., shorter freezing time) reflects a less reactive behavioural response, often associated to lower HPA/HPI activity [@koolhaas1999coping].
 In _L.
 delicata_, individuals incubated at 23 ºC exhibit elevated baseline CORT [@crino2024eggs], potentially enhancing sensitivity to stress.
-Therefore, the heightened acute response in cold-incubated individuals may be explained by elevated endogenous CORT levels.
-For example, **E.G.
-about CORT and response**.
-
-In contrast, early thermal environment had no effect on _L.
-guichenoti_'s response to an acute stressor.
-The species-specific response also warrants discussion.
-L.
-delicata exhibited treatment-dependent variation in response to acute stress, while L.
-guichenoti did not.
-This difference may reflect distinct developmental plasticity or stress-coping strategies between species.
-Given that L.
-delicata is a successful invasive species, greater sensitivity to early-life thermal environments and a more flexible behavioural response to acute threats may confer adaptive advantages in novel environments [REFs].
-In contrast, the more canalised stress response in L.
-guichenoti could reflect reduced behavioural plasticity, potentially limiting its adaptability in changing environments.
-
-Strikingly, we found no effects of prenatal CORT treatment on the acute stress response in either species.
-Glucocorticoid elevation during development are known to modulate stress-related phenotypes, including behavioural responsiveness to processive stressors [**REF...before??
-I might have another in the intro**].
-Previous studies have shown that
-
-#### Habituation to stressors varies between species and treatments
-Across repeated exposures to the simulated predatory attack, only L.
-delicata individuals from the CORT-Cold treatment increased their latency to move.
-Rather than indicating habituation, this pattern may reflect accumulating allostatic load.
-Habituation is typically inferred from a reduction in stress-related behaviours over time (e.g., decreased latency to resume activity), whereas an increase in latency could reflect mounting stress or impaired coping [@grissom2009habituation].
-Thus, the behavioural trajectory of the CORT-Cold group may suggest that these animals became less, rather than more, tolerant to the repeated stressor.
-
-Why did this pattern emerge only in the CORT-Cold L.
-delicata group?
-While our CORT treatment did not significantly elevate baseline hormone levels across groups, previous findings indicate that cold incubation alone raises baseline CORT [@crino2024eggs].
-The combination of prenatal CORT and cold temperature may therefore have produced a cumulative or synergistic effect on HPI axis development and reactivity.
-This effect might not have been detectable in our acute response measures, but could become apparent over repeated exposures as stress accumulates.
-
-Alternatively, cognitive explanations are also plausible.
-The cold-CORT combination may impair neural development in regions critical for behavioural flexibility and non-associative learning.
-Although our prior work found no significant impact of this combination on associative learning (e.g., colour discrimination and reversal learning; @recio2025cognitive; @recio2025early), cognitive processes underpinning habituation may rely on different neural substrates.
-The hippocampus and prefrontal cortex, implicated in behavioural inhibition and decision-making, may be differentially affected by early-life CORT and temperature exposure.
-Region-specific developmental trajectories could account for divergent effects across learning types.
-
-In contrast, L.
-guichenoti showed no evidence of behavioural change across repeated stress exposures, regardless of treatment.
-This further supports the notion of reduced behavioural flexibility or a more rigid stress-response phenotype in this species.
-It is noteworthy that L.
-delicata displayed treatment-specific changes even when these did not conform to classical habituation.
-Such responsiveness may reflect a greater capacity to modulate behaviour based on environmental cues, a trait often associated with invasive potential [REFs].
-
-Interestingly, across all treatments, L.
-delicata increased their probability of emerging from shelter within 40 minutes of the stressor as the experiment progressed, suggesting that overall risk perception decreased over time.
-This pattern is consistent with habituation at the level of decision-making rather than immediate post-threat responses (i.e., latency to move).
-L.
-guichenoti, in contrast, exhibited no such trend.
-This species difference in shelter-emergence behaviour provides further support for differential flexibility in stress evaluation and risk assessment.
-The divergence may again be rooted in differences in developmental plasticity or evolutionary history, with L.
-delicata potentially possessing a more labile stress-response system suited to fluctuating environments.
-
-The absence of time effects in latency to move or go to shelter for the other treatments in L.
-delicata suggests that habituation was limited and potentially context- or trait-dependent.
-Taken together, our findings highlight the complexity of stress habituation and the need to disentangle multiple behavioural and physiological components to fully understand coping strategies.
+Therefore, the heightened acute response in cold-incubated individuals may be explained by elevated endogenous CORT.
+Alternatively, the higher latency times to move may reflect impaired neural development caused by cold incubation.
+Colder incubation temperatures can disrupt the formation and connectivity of brain regions involved in sensory processing and behavioural control [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017], ultimately leading to slower perception and action planning during freezing [@roelofs2022freezing].
+
+In contrast, incubation temperature had no effect on _L.
+guichenoti_'s response to an acute stressor, suggesting that stress reactivity in this species is more robust to early thermal environment.
+Our findings align with previous studies showing that incubation temperature impacts the antipredatory behaviour of _L.
+delicata_ but not _L.
+guichenoti_ [@iglesias2024maternal].
+Between-species differences on the effect of incubation temperature could reflect differences in history-traits such as egg size.
+Smaller eggs in _L.
+delicata_ may make this species more sensitive to environmental alterations during development, with consequences for their physiology and behaviour later in life [@crino2024eggs; @iglesias2024maternal].
+Although heightened plasticity may enable _L.
+delicata_ to fine-tune its phenotype to local conditions, it could also increase vulnerability to suboptimal environments.
+In contrast, the apparent robustness of _L.
+guichenoti_ might buffer against certain environmental inputs but limit flexibility [@fox2019beyond].
+Future studies should explore how these differing developmental strategies influence fitness across different environments.
+
+#### Prenatal CORT did not influence behavioural response to an acute stressor     
+Given the significant effects of prenatal GCs on the development of the HPA/HPI axis reported in other species [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs], we predicted prenatal CORT elevations to increase lizards' reactivity to acute stressors.
+However, we found no effects of prenatal CORT elevations on any of the behaviours analysed, and these results were consistent across species.
+These findings suggest that the response to acute stressors is robust to CORT elevations.
+
+Alternatively, our CORT manipulation may have failed to elicit significant effects.
+Glucocorticoids are known to have hormetic effects on different traits [@mcewen_concept_2003].
+While our CORT treatment was efficient in elevating yolk CORT concentrations, they did not result in increased baseline CORT in juveniles[@crino2024eggs].
+In contrast, doses at lower concentration did [@crino2024eggs].
+Previous reports show that enhanced responsiveness to acute stressors are usually accompanied by higher baseline GC levels [@koolhaas1999coping; @ruys2004behavioral; @thaker2009acute].
+Therefore, the absence of CORT effects on the acute response may reflect the lack of endocrine differences between treatment groups in _L.
+delicata_.
+However, this mechanism remains speculative for _L.
+guichenoti_, as comparable endocrine data are currently lacking.
+Further experiments would benefit from studying how exposing eggs to lower CORT concentrations influence stress-related physiology and behaviours in both species.
+
+#### The effects of prenatal conditions on the response to a repeated stressor are species-dependent   
+We predicted that, over time, lizards would decrease the latency to move while increasing both the latency to shelter and the probability of emerging from the shelter.
+These patterns would reflect a more relaxed response to a processive stressor, indicative of cognitive habituation [@koolhaas1999coping; @grissom2009habituation].
+We also expected that prenatal CORT exposure or cold incubation would impair habituation, given their known effects on the HPI axis and cognitive development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta; @amiel_hotter_2012; @vargas2016early; @zito2017early; @crino2024eggs].
+Nonetheless, our results show complex patterns were habituation differs across species, treatments, and behavioural measures.
+
+Only _L.
+delicata_ showed evidence that early-life conditions influenced the response to a chronic stressor.
+Although slope contrasts did no reveal significant differences between treatments, the CORT-Cold group was the only one where the latency to move changed significantly over time.
+However, instead of the predicted reduction, we found an increase in the time spent frozen after the simulated attack, suggesting sensititazion to the stressor rather than habituation.
+This result may reflect a cumulative effect of prenatal CORT and cold incubation temperatures, even in the absence of a statistically significant interaction.
+In _L.
+delicata_, cold incubation alone increases baseline CORT levels in juveniles [@crino2024eggs].
+Although the high CORT dose used in our experiment did not significantly elevate baseline CORT on its own, it produced intermediate values between controls and a lower dose — the latter of which did significantly elevate baseline levels [@crino2024eggs].
+Therefore, it is possible that the combination of cold incubation and the intermediate endocrine effect of the high CORT dose could together increase HPI axis activity enough to enlarge behavioural reactivity over time [@koolhaas1999coping].
+
+Alternatively, the combination of CORT and cold temperatures could be impairing habituation trough effects on neural development [see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017].
+Although previous studies found no influence of prenatal CORT or temperature on associative and reversal learning [@recio2025cognitive; @recio2025early], the effect of prenatal conditions can vary by brain region [@coomber_independent_1997; @sakata_neural_2000].
+Possibly, neural structure of regions involved cognitive habituation are more sensitive to early-life conditions than those related to other cognitive domains.
+Further work is needed to test this possibility and identify the specific neural mechanisms linking developmental conditions to stress-related behavioural flexibility.
+
+#### _L. delicata_, but not _L. guichenoti_, shows habituation to a processive stressor    
+We found that _L.
+delicata_ increased the probability of emergence from the shelter, which suggest that risk perception decreased over time.
+In contrast, _L.
+guichenoti_ showed no sign of habituation.
+Although the slopes did not significantly differ between species for most of the treatments (see @table_contrasts_species), contrasts show that change over time was overall larger in _L.
+delicata_.
+Futhermore, only in _L.
+delicata_ slopes were different from zero, suggesting that there are differences between species in the habituation to a processive stressor.
+Differences in habituation may contribute to _L.
+delicata_'s success as an invasive species, as greater flexiblility in responding to processive stressors could facilitate adaptation to new environments [@chapple_biology_2014; @chapple2015deliinvLHI].
+For instance, learning to reassess risk based on previous exposure can reduce the costs of the antipredatory response [@lima1990behavioral].
+Moreover, cognitive habituation may reflect and individual's capacity to cope with the physiological toll of chronic stress [@grissom2009habituation], a potentially advantageous trait when colonising new habitats [@wingfield2013comparative; @surkova2024flexible].
+
+Nevertheless, the relative contribution of cognitive versus physiological mechanisms to habituate remains unclear.
+Previous studies found no differences in associative learning or behavioural flexibility between _L.
+delicata_ and _L.
+guichenoti_ [@bezzina2014does; @recio2025cognitive; @recio2025early], suggesting that cognitive processes alone may not explain the between-species differences observed here.
+Likewise, baseline CORT levels do not fully account for the patterns either: hot-incubated _L.
+delicata_, which exhibit lower baseline CORT [@crino2024eggs], did not habituate faster than cold-incubated lizards.
+Instead, our results point to a complex interaction between cognitive and physiological habituation.
+Future studies that measure behavioural traits and CORT dynamics concurrently should be performed to disentangle the cognitive and physiological components of habituation.
 
 #### Physiological costs of the stressor were low and did not differ between treatments or species      
-Despite repeated exposure to a stressor over several days, we found no clear evidence of physiological costs in either species.
-Most individuals gained mass over the course of the experiment, and there were no treatment or species differences in mass change.
-This suggests that the imposed stressor, while sufficient to elicit behavioural responses, did not impair growth or energy balance under our conditions.
-This result is somewhat surprising, given the literature linking chronic stress to impaired growth, particularly in juvenile ectotherms [REFs].
-
-One possible explanation is that the stressor we employed, while ecologically relevant, was not sufficiently intense or frequent to produce measurable physiological effects.
-In natural settings, stressors may be more variable and compound with other challenges (e.g., food scarcity, thermal extremes), resulting in more pronounced physiological consequences.
-Our controlled environment, consistent food availability, and short stress exposure duration may have buffered such effects.
-
-Another consideration is that mass change may not fully capture the physiological costs of stress.
-Other endpoints, such as oxidative stress markers, telomere attrition, or immune function, may be more sensitive to sublethal chronic stress effects [REFs].
-Future studies incorporating these physiological markers would provide a more comprehensive picture of the costs associated with repeated stress.
-
-Finally, it is important to note that our behavioural findings suggest differences in stress evaluation and habituation across treatments and species, even in the absence of overt physiological changes.
-This dissociation underscores the importance of measuring multiple dimensions of the stress response and not relying solely on growth or condition as proxies for stress burden.
+Given the metabolic toll that enhanced GC levels impose on animals [@sapolsky_how_2000; @picard_mitochondria_2014], we predicted that repeated exposure to a stressor would influence body mass.
+Furthermore, we expected those costs to be related to alterarions in stress reactivity shaped by prenatal conditions.
+In contrast, we found treatment effects in mass change, and most of the lizards gained rather than lost mass.
+Notably, mass change did not differ between species despite differences in their flexibility to respond to a chronic stressor - a trait oftenly associated with more effective coping [@grissom2009habituation].
+These findings suggest that repeated exposure to a processive stressor imposes minimal metabolic costs on _L.
+delicata_ and _L.
+guichenoti_.
+
+Nevertheless, we cannot exclude the possibility that other, subtler physiological costs occurred but were not captured by overall changes in mass.
+Other studies report impairments in immune suppression, impaired growth, or permanent neural alterations following repeated exposure to systemic and processive stressors  [@picard_mitochondria_2014; @mcewen2017neurobiological].
+Future research should investigate how chronic stress impacts other physiological traits, and whether developmental conditions modulate the associated costs.
 
 #### Conclusion
-Our study demonstrates that early-life environments—specifically incubation temperature and prenatal glucocorticoid exposure—shape behavioural responses to acute and repeated stressors in skinks.
-These effects were species-specific, with L.
-delicata showing greater responsiveness to early-life conditions than L.
-guichenoti.
-Cold incubation increased acute stress reactivity in L.
-delicata, and the combination of cold and prenatal CORT impaired habituation, potentially via elevated baseline CORT or neural alterations.
-In contrast, L.
-guichenoti appeared less flexible in both acute and repeated stress responses.
-
-These findings contribute to our understanding of how early developmental environments shape adult stress phenotypes.
-They also highlight potential mechanisms underlying behavioural plasticity in invasive species, where adaptability to repeated stressors may facilitate persistence in novel environments.
-Future work should aim to link behavioural and physiological responses across ontogeny and explore neural mechanisms that underpin different forms of learning and habituation.
-
-#### Cold incubation temperature increases the response to an acute stressor in _L. delicata_ but not in _L. guichenoti_
-**CHECK PROPER UNITS**
-Hot-incubated _L.
-delicata_ spent less time frozen after the simulated attack.
-We did not see any differnces betwen treatments in any other variable recorded.
-No treatment effect on the response to an acute stressor in _L.
-guichenoti_.
-- Why do we consider less time to be less reactive?
-- Why differences in temp? --> Cold animals had higher baseline CORT levels...but CORT treated too...at least mean treatment but not high.
-- Why not differences based on CORT?
-- Why is temperature in _L. delicata_ more important for this than in _L. guichenoti_?? Invassiveness??
-
-#### Habituation to stressors varies between species and treatments
-Only CORT-Cold _L.
-delicata_ increase the latency to move over time, suggesting allostatic load more than habituation, as we would expect animals to resume their activity earlier with habituation.
-However, the analyses did not show significant differences between the contrasts.
- - WHY those resuls? Power?
- - Why CORT-Cold --> higher baseline CORT levels??
-CORT at the levels we used increases slightly baseline CORT blood levels in juveniles of _L.
-delicata_, this difference was not significant compared to the controls, but it could be enough for increasing at cold temps when CORT levels are already higher than at warm temps [@crino2024eggs].
-This seems to make CORT-treated _L.
-delicata_ incubated at cold temperatures as reactive to acute stress as Cold-Controls, but less prone to habituation.
-    Alternatively, this response could be due to lower cognitive abilities as expected by CORT + cold.
-However, this hasn't been seen in other learning abilities [@recio2025cognitive; @recio2025early].
-Nevertheless, the effects of temperature and CORT are expectd to be region-dependent, so there could be effects in non-associative learning while not in associative types of learning (i.e., colour association or its reveresal)
- - Why not in _L. guichenoti_??
-The latency to move and to go to shelter did not change over time in any other treatment in _L. delicata_.
-All _L. delicata_ increase the probability of emerging in 40 min after the attack, suggesting that the risk perception decreased during the experiment. This did not happen in _L. guichenoti_   
-  - Risk perception seems the same both at the beginning and during the experiment for all treatments despite the frozen behaviour being higher in Cold-incubated lizards. Why??
-  - Why differences between species?? Is _L. delicata_ more flexible in evaluating risks?? --> Relationship with invasiveness
-
-  **CHECK TABLE FOR SPP COMPARISSONS**
-
-#### Physiological costs of the stressor were low and did not differ between treatments or species
-There were no "physiological costs"
-- Most of the individuals grew up...How's the growth compared to other studies where they were not subjected to a stressor?? What are the usual "costs"??
-- Caveats of the methods: other costs
+Our study demonstrates that incubation temperature and prenatal CORT can shape behavioural responses to stress in a species- and context-dependent manner.
+_Lampropholis delicata_ showed greater plasticity in response to both acute and repeated exposure to a processive stressor, while _L.
+guichenoti_ appeared robust to prenatal conditions.
+These species differences may reflect divergent developmental strategies, with potential implications for ecological resilience and invasion success.
+Despite measurable behavioural differences, we found no evidence of metabolic costs - measured as changes in mass - associated with stress exposure under our experimental conditions.
+Together, our findings highlight the value of behavioural measures as accessible indicators of stress response and developmental plasticity.
+Future research integrating hormonal and neural data will be key to understanding the mechanisms underlying behavioural flexibility and its fitness consequences across environments.
 
 ## References
+<div id="refs"></div>
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
 

### 2025-10-17 R/data_process.R
updates

@@ -18,23 +18,13 @@ clean_ind <- read.csv(here("data", "general_data.csv")) %>%
 data.frame()
 
 ####################################
-# B) Merging mass.csv with clean individual data (clean_ind)
-mass <- read.csv(here("data", "mass.csv")) %>%
-  mutate(lizard_id = as.character(lizard_id))%>%
-data.frame() #Read in mass data and transform lizard_id to character
-mass_clean <- merge(clean_ind, mass, by = "lizard_id", all = TRUE) %>%
-  mutate(delta_mass_rescaled = (delta_mass - min(delta_mass, na.rm = TRUE) + 1)) %>%
-  mutate(food_ingested = food_ingested - mean(food_ingested, na.rm = TRUE))
-data.frame() #Merge mass data with clean_ind data and rescale delta_mass
-
-write.csv(mass_clean, here("./output/data_clean/mass_clean.csv")) 
-
-####################################
-# C) Merging behaviour.csv with clean individual data (clean_ind)
+# B) Merging behaviour.csv with clean individual data (clean_ind)
 behav <- read.csv(here("data", "behaviour.csv"))%>%
   mutate(lizard_id = as.character(lizard_id))%>%
   mutate(day = day - 1) %>% # Adjust day to start from 0
 data.frame() #Read in behaviour data and transform lizard_id to character
-behav_clean <- merge(clean_ind, behav, by = "lizard_id", all = TRUE) 
 
+behav_clean <- merge(clean_ind, behav, by = "lizard_id", all = TRUE) %>%
+  mutate(svl = scale(as.numeric(svl), center = TRUE, scale =FALSE)) %>%
+data.frame()
 write.csv(behav_clean, here("./output/data_clean/behav_clean.csv"))

### 2025-10-17 R/exploratory.R

@@ -7,26 +7,7 @@ ggh4x, cowplot, fitdistrplus, MASS, goftest)
 source(here("R", "func.R"))
 
 ####################################
-# A) Exploring distribution of the main variables for mass_clean 
-data_expl <- mass_clean %>%
-       mutate(treatment = paste(cort, temp, sep = "-"))
-# Histogram
-bin_width <- (max(data_expl$delta_mass_rescaled, na.rm = TRUE) - min(data_expl$delta_mass_rescaled, na.rm = TRUE)) / sqrt(length(data_expl$delta_mass_rescaled))
-hist_plot_mass <- ggplot(data_expl, aes(x = delta_mass_rescaled)) +
-  geom_histogram(binwidth = bin_width, fill = "#062d00", color = "black", alpha = 0.5) +
-  theme_classic() +
-  labs(x = "delta_mass", y = "Counts") +
-  theme(axis.title = element_text(size = 12, family = "Times"),
-        axis.text = element_text(size = 10, family = "Times"))
-ggsave(here("./output/figures/exploratory/hist_mass.png"), hist_plot_mass, width = 20, height = 20, dpi = 300)
-
-# Distribution
-qq_plot_mass <- qq_plots_single(data_expl, "delta_mass_rescaled", "mass")
-ggsave(here("./output/figures/exploratory/qq_plot_mass.png"), qq_plot_mass, width = 20, height = 20, dpi = 300)
-#### Delta mass considered as log-normal distribution
-
-####################################
-# B) Exploring distribution of the main variables for behav_clean
+# A) Exploring distribution of the main variables
 data_expl <- behav_clean %>%
   mutate(treatment = paste(cort, temp, sep = "-")) %>%
   mutate(

### 2025-10-17 R/func.r

@@ -68,18 +68,12 @@ sample <- function(df, corti, therm){
 #' @param fam To specify the family of the model
 #' @param sp To specify the species
 #' @param var To specify the response variable
-#' @param type To specify which type of model/formula
 #' @return Raw posteriors of fitted brm model for each treatment, species, and group (df)
-fit_m <- function(df, fam, sp, type, var) {
+fit_m <- function(df, fam, sp, var) {
   #Fit the model only if it has not been fit yet (if refit=TRUE)
   if(refit){
-    if(type == "beh"){
-      formula <- as.formula(paste(var, 
-                            "~ day*trt + (1|lizard_id) + (1|clutch)"))
-    } else if(type == "mass"){
-      formula <- as.formula(paste(var, 
-                            "~ trt + food_ingested + (1|clutch)"))
-    }
+    formula <- as.formula(paste(var, 
+                            "~ day*trt + svl + (1|lizard_id) + (1|clutch)"))
     # Fit the model
     model <- brm(formula = formula,
                 data = df,

### 2025-10-18 ms.qmd
updates

@@
 ---
-title: "Effects of prenatal CORT and temperature on the response to a stressor"
+title: "Effects of prenatal CORT and temperature on the antipredatory behaviour in a lizard"
 format:
@@
 bibliography: "./bib/refs.bib"
-csl: "./bib/animal-behaviour.csl"
+csl: "./bib/behavioral-ecology-and-sociobiology.csl"
 execute:
@@
 
-  data_n <- mass_clean %>% 
+  data_n <- behav_clean %>%
+    filter(day == 0) %>%
     filter(species == s)
@@
 #
-clutches_deli <- mass_clean %>% 
+clutches_deli <- behav_clean %>% 
   filter(species == "delicata") %>%
+  filter(day == 0) %>%
   distinct(clutch) %>% 
@@
 #
-clutches_guich <- mass_clean %>% 
+clutches_guich <- behav_clean %>% 
   filter(species == "guichenoti") %>%
+  filter(day == 0) %>%
   distinct(clutch) %>% 
@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Mathéo Dupuy^1,5^, Daniel W.A. Noble^1^
+Pablo Recio^1,2^,$\ddagger$ , Daniel W.A. Noble^1^, Dalton C. Leibold^1^, Christopher R. Friesen^3,4^, Mathéo Dupuy^1,5^, Ondi L. Crino^1,2^
 
@@
 Pablo Recio ORCID: 0000-0002-5890-0218  
+Daniel W.A. Noble ORCID: 0000-0001-9460-8743     
 Dalton C. Leibold ORCID: 0000-0001-9645-2033  
+Christopher R. Friesen ORCID: 0000-0001-5338-7454  
+Mathéo Dupuy ORCID:         
 Ondi L. Crino ORCID: 0000-0001-5700-1387
-Christopher R. Friesen ORCID: 0000-0001-5338-7454  
-Daniel W.A. Noble ORCID: 0000-0001-9460-8743
 
@@
 guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Differences in the ability to colonise new environments are not related to cognitive abilities [@bezzina2014does; @recio2025cognitive; @recio2025early], but could be associated with differences in personality [@chapple_know_2011], which may be linked to different responses to acute and chronic processive stressors [@koolhaas1999coping].
+Differences in the ability to colonise new environments are not related to cognitive abilities [@bezzina2014does; @recio2025cognitive; @recio2025early], but could be associated with differences in personality [@chapple_know_2011], which may be linked to different responses to  predatory stimuli [**REFS**].
 
@@
 
-#### Response to acute and long-term stressors
+#### Antipredatory response
 
 Three weeks before the start of the behavioural tests, we adapted the lizard enclosures to the experimental setup.
-The enclosures used for the stress response trials were the same as those where the lizards had previously been housed, but modified to include only a single shelter (9 × 6 × 1.5 cm) placed at one end and a water dish in the centre.
+The enclosures used for the antipredatory response trials were the same as those where the lizards had previously been housed, but modified to include only a single shelter (9 × 6 × 1.5 cm) placed at one end and a water dish in the centre.
 All other shelter materials and the matting substrate were removed to allow continuous monitoring.
@@
 
-To test the effects of prenatal conditions on the response to a stressor, we simulated a predatory attack - a standardized processive stressor [@thaker2009acute; @trompeter2011invader; @iglesias2024maternal] - once daily over multiple days.
-Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the shelter.
+To test the effects of prenatal conditions on the short- and long-term antipredatory response, we simulated a predatory attack once daily over multiple days.
+Each trial began with one of the researchers (PR) removing the water dish from the enclosure, followed by the removal of the shelter.
 The lizard was then chased for 60 seconds using a soft paintbrush, simulating the predatory attack.
@@
 All videos were later analysed by MD, who was blinded to the lizards’ treatment.
-From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) emergence — whether the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
+From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) probability to emerge — whether or not the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
 A lizard was considered to have resumed movement if it moved continuously for at least five seconds, and to be under shelter if no limbs were visible.
-This procedure was repeated daily for eight consecutive days to assess the effects of long-term exposure to a processive stressor.
 All trials were conducted between 1000–1400 h, when lizards were most active.
-
-Both before and after the completion of the behavioural tests, lizard mass was measured to the nearest 0.001 g using a digital scale.
-Change in mass was calculated as the difference between the final minus the initial mass.
-Change in mass was used as a proxy for the metabolic cost of repeated stress exposure, where mass loss or lower growth was considered as higher costs.
-To control for hunger levels and potential effects of food ingestion on mass change, we also recorded the number of crickets ingested by each lizard during the acclimation period.
-To do so, we placed a known number of crickets in each enclosure after each of the trials, and counted the number of crickets remaining after 24 hours.
-The total number of crickets ingested during the eight days of tests was then used as a covariate in the analyses of mass change.
-The experimental design is summarised in @fig-Methods B.
 
@@
 In panel A, the early environment manipulations leading to four experimental conditions.
-In panel B, the timeline of the postnatal stress-response assay: on Day 0 and 9, lizard mass was recorded; from Days 1–8, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
-Variables measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor."
+In panel B, the postnatal antpredatory response assay, for 8 days, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
+During this recording, we measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor."
 
@@
 We performed the analyses for each species separately.
-We fitted a total of eight models using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.4.0 [@R].
+We fitted a total of six models using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.5.1 [@R].
 We ran four parallel MCMC chains of 10000 iterations for each model, with a warmup period of 4000 iterations.
 
-We used as the response variables the latency to move, the latency to shelter, and the emergence probability, as well as the change in mass after the eight days of stress exposure.
-We fitted separate models for each response variable.
-For the three behavioural variables, we included the prenatal condition (Control-Cold, CORT-Cold, Control-Hot, CORT-Hot), the day of the trial, and their interaction as fixed effects.
+We used as the response variables the latency to move, the latency to shelter, and the probability to emerge, modelled separately.
+We included the prenatal condition (Control-Cold, CORT-Cold, Control-Hot, CORT-Hot), the day of the trial, and their interaction as fixed effects.
 To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e., the intercept reflects lizard behaviour in response to the first exposure to the stressor).
-For the variable mass, we included the prenatal condition and the total number of crickets ingested per each individual during the eight days of stress exposure as predictors.
-
-We included each lizard's random intercept and slope (trial) as a random factor in the behavioural models, but not when mass was the response.
-For all models, we included the clutch as a random factor.
-_L.
+We also included snout-vent length - measured one day before the beginning of the trials - to control for potential effects of size on the antipredatory behaviour [**REFs**].
+SVL was centered to the mean to facilitate interpretation of the intercepts.
+
+We included lizard identity and clutch identity as a random factors.
+The former was included to account for individual performance over time, while clutch identity was included to control for parental conditions.
+In this sense, _L.
 delicata_ lays one clutch per year, while _L.
 guichenoti_ lays two [@chapple_know_2011; @chapple2015deliinvLHI].
-Since eggs were collected during half of the breeding season, clutches likely come from different mothers.
-Additionally, previous research has shown that clutches are generally sired by a single male, but sperm storage can occur [@kar2023heritability].
-Given our partial split-clutch design and the fact that maternal effects are expected to be stronger than paternal effects in these species, including the clutch as a random factor should account for the effects of parental identity.
+Since eggs were collected during half of the breeding season, each clutch likely comes from a different mother.
+Additionally, previous research has shown that clutches are generally sired by a single male [@kar2023heritability].
+Although sperm storage can occur in these species [@kar2023heritability], given our partial split-clutch design and the fact that maternal effects are expected to be stronger than paternal effects in these species, including the clutch as a random factor should account for the effects of parental identity.
 
 We used the posterior distributions of model parameters to test for differences between prenatal conditions.
-For behavioural variables, differences in the model intercepts between conditions were interpreted as differences in the acute response to a stressor (i.e., the first day of exposure), while differences in the slopes were interpreted as differences in behavioural adjustment across repeated exposures (i.e., chronic stress response).
-For mass change, between-treatments differences reflect differences in the energetic cost of the chronic stressor.
+Differences in the model intercepts between conditions were interpreted as differences in the immediate antipredatory behaviour, while differences in the slopes were interpreted as differences in antipredatory adjustments.
 We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments or the slopes for each treatment were different from zero.
@@
 Differences between species were tested by comparing the posterior distributions of the intercepts and slopes of the different treatments.
-We used the posterior distributions of the intercepts and slopes to test for differences between species in their response to an acute stressor (intercept) and a chronic stressor (slope).
 95% HPDI were used to test the hypothesis that the contrasts differed from zero.
@@
 All the estimated parameters in the results were transformed back to original units employing appropriate formulas depending on the distribution employed in the models.
-The error structure for latency to move, latency to shelter, or mass change followed a lognormal distribution [family = lognormal()], while the probability of emerging from the shelter was modeled using a Bernoulli distribution with a logit link function [family = Bernoulli(link = 'logit')].
-Change in mass was also rescaled by adding the minimum change in mass recorded before running the models to avoid negative values but transformed back to original values for the presentation of the results.
+The error structure for latency to move and latency to shelter followed a lognormal distribution [family = lognormal()], while the probability of emerging from the shelter was modelled using a Bernoulli distribution with a logit link function [family = Bernoulli(link = 'logit')].
 
@@
 However, our final sample size was 85 due to natural mortality (n = 9).
-The final sample sizes per treatment and species are listed in @fig-results_beh_deli, @fig-results_beh_guich, and @fig-results_mass.
+The final sample sizes per treatment and species are listed in @fig-results_beh_deli and @fig-results_beh_guich.
 These animals came from a total of `r clutches_deli` clutches in _L.
@@
 #
-deli_move <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_move")
-deli_shelter <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_shelter")
-deli_emergence <- fit_m(df = deli_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "deli", var = "emergence")
+deli_move <- fit_m(df = deli_beh, fam = lognormal(), sp = "deli", var = "lat_move")
+deli_shelter <- fit_m(df = deli_beh, fam = lognormal(), sp = "deli", var = "lat_shelter")
+deli_emergence <- fit_m(df = deli_beh, fam = bernoulli(link = "logit"), sp = "deli", var = "emergence")
 #
@@
 #
-guich_move <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_move")
-guich_shelter <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_shelter")
-guich_emergence <- fit_m(df = guich_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "guich", var = "emergence")
-```
-
-```{r, models_mass}
-#| label: models_mass
-source(here("R", "func.R"))
-#
-refit <- FALSE
-#
-# A) L. delicata
-deli_mass <- mass_clean %>%
-  filter(species == "delicata")
-#
-deli_mass <- fit_m(df = deli_mass, fam = lognormal(), type = "mass", sp = "deli", var = "delta_mass_rescaled")
-# 
-#
-# B) L. guichenoti
-guich_mass <- mass_clean %>%
-  filter(species == "guichenoti")
-#
-guich_mass <- fit_m(df = guich_mass, fam = lognormal(), type = "mass", sp = "guich", var = "delta_mass_rescaled")
+guich_move <- fit_m(df = guich_beh, fam = lognormal(), sp = "guich", var = "lat_move")
+guich_shelter <- fit_m(df = guich_beh, fam = lognormal(), sp = "guich", var = "lat_shelter")
+guich_emergence <- fit_m(df = guich_beh, fam = bernoulli(link = "logit"), sp = "guich", var = "emergence")
 ```
@@
 ## A) L. delicata
-deli_move_posteriors <- post_values(deli_move, var = "beh")
-deli_shelter_posteriors <- post_values(deli_shelter, var = "beh")
-deli_emergence_posteriors <- post_values(deli_emergence, var = "beh")
-deli_mass_posteriors <- post_values(deli_mass, var = "mass")
+deli_move_posteriors <- post_values(deli_move)
+deli_shelter_posteriors <- post_values(deli_shelter)
+deli_emergence_posteriors <- post_values(deli_emergence)
 #
 ## B) L. guichenoti
-guich_move_posteriors <- post_values(guich_move, var = "beh")
-guich_shelter_posteriors <- post_values(guich_shelter, var = "beh")
-guich_emergence_posteriors <- post_values(guich_emergence, var = "beh")
-guich_mass_posteriors <- post_values(guich_mass, var = "mass")
+guich_move_posteriors <- post_values(guich_move)
+guich_shelter_posteriors <- post_values(guich_shelter)
+guich_emergence_posteriors <- post_values(guich_emergence)
 #
@@
 deli_emergence_contrasts <- contrasts_func(sp = "deli", res = "emergence")
-deli_mass_contrasts <- contrasts_func(sp = "deli", res = "mass")
 #
@@
 guich_emergence_contrasts <- contrasts_func(sp = "guich", res = "emergence")
-guich_mass_contrasts <- contrasts_func(sp = "guich", res = "mass")
-```
-
-#### Do prenatal conditions affect the behavioural response to an acute stressor?
+```
+
+#### Do prenatal conditions affect the immediate antipredatory response?
 
 _Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I~Cold~ = `r deli_move_contrasts$median_Temperature_int`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
-We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability after an acute stressor (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
-
-_Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had any effect on the behavioural response to an acute stressor on _L.
+We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability on the first day of the assay (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
+SVL had no significant effect on any of the behavioural responses (see Table S4 in _Supplementary Material_).
+
+_Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had a significant effect on the immediate antipredatory response on _L.
 guichenoti_ (see @fig-results_beh_guich A, D, G and Table S2 in _Supplementary Material_).
+We did not see any significant effects of SVL on any of the behavioural responses (see Table S5 in _Supplementary Material_).
 
@@
 ggsave("./output/figures/main/fig_deli_beh.png", plot=fig_deli_beh, width = 21, height = 25, units = "cm", dpi = 600)
+ggsave("./BEHPROC/submitted_material/fig_deli_beh.pdf", plot=fig_deli_beh, width = 22, height = 26, units = "cm", dpi = 1200)
 knitr::include_graphics("./output/figures/main/fig_deli_beh.png")
@@
 
-#### Do prenatal conditions affect the behavioural response to a chronic stressor?
+#### Do prenatal conditions affect the long-term antipredatory response?
 
@@
 ggsave("./output/figures/main/fig_guich_beh.png", plot=fig_guich_beh, width = 21, height = 25, units = "cm", dpi = 600)
+ggsave("./BEHPROC/submitted_material/fig_guich_beh.pdf", plot=fig_guich_beh, width = 22, height = 26, units = "cm", dpi = 1200)
 knitr::include_graphics("./output/figures/main/fig_guich_beh.png")
-```
-
-#### Do prenatal conditions influence the costs of repeated stress exposure?
-
-```{r, process_posteriors_mass}
-#| label: process_posteriors_mass
-# Code to make the posterior distribution of the mass models easier to call for comparissons of Δmass
-source(here("R", "func.R"))
-#
-# A) L. delicata
-mass_deli <- deli_mass_posteriors %>%
-  mutate(across(everything(), ~ exp(.) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
-  rename_with(~ str_remove(., "^int_"))
-# B) L. guichenoti
-mass_guich <- guich_mass_posteriors %>%
-  mutate(across(everything(), ~ exp(.) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
-  rename_with(~ str_remove(., "^int_"))
-```
-
-_Lampropholis delicata_: Overall, animals grew over time (Median Δmass = `r format_dec(median(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot)),3)` mg; 95% HDPI = [`r format_dec(hdi(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot), credMass = 0.95)[1],3)` ,`r format_dec(hdi(c(mass_deli$ControlCold, mass_deli$CORT_Cold, mass_deli$Control_Hot, mass_deli$CORT_Hot), credMass = 0.95)[2],3)`]), with no effect of temperature, CORT, or their interaction (see @fig-results_mass A and Table S4 in _Supplementary Material_).
-
-_Lampropholis guichenoti_: Mass of lizards increased over time (Median Δmass = `r format_dec(median(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot)),3)` mg; 95% HDPI = [`r format_dec(hdi(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot), credMass = 0.95)[1],3)` ,`r format_dec(hdi(c(mass_guich$ControlCold, mass_guich$CORT_Cold, mass_guich$Control_Hot, mass_guich$CORT_Hot), credMass = 0.95)[2],3)`]), and the contrasts between treatments showed no differences in the change in mass (see @fig-results_mass B and Table S4 in _Supplementary Material_).
-
-```{r, fig-results_mass}
-#| label: fig-results_mass
-#| fig.cap: "Change in mass of _Lampropholis delicata_ (Panel A) and _Lampropholis guichenoti_ (Panel B)after eight days of repeated stress exposure." 
-#
-#
-source(here("R", "func.R"))
-#
-## L delicata
-deli_mass_plot <- plotting(sp = "deli", res = "mass", type = "int") +
-  theme(legend.position = "none",
-        legend.title = element_blank())
-# Getting the legend
-legend_plot_deli <- plotting(sp = "deli", res = "mass", type = "int") + 
-  theme(legend.position = "right", legend.title = element_blank())
-gtable_deli <- ggplotGrob(legend_plot_deli)
-legend_deli <- gtable_deli$grobs[[which(sapply(gtable_deli$grobs, function(x) x$name) == "guide-box")]]
-#
-## L guichenoti
-guich_mass_plot <- plotting(sp = "guich", res = "mass", type = "int") +
-  theme(legend.position = "none",
-        legend.title = element_blank())
-# Getting the legend
-legend_plot_guich <- plotting(sp = "guich", res = "mass", type = "int") + 
-  theme(legend.position = "right", legend.title = element_blank())
-gtable_guich <- ggplotGrob(legend_plot_guich)
-legend_guich <- gtable_guich$grobs[[which(sapply(gtable_guich$grobs, function(x) x$name) == "guide-box")]]
-#
-mass_grid <- plot_grid(deli_mass_plot, NULL, guich_mass_plot, NULL,
-                       ncol = 4, nrow = 1,
-                       rel_widths = c(0.5, 0.5, 0.5, 0.45))
-#
-fig_mass <- ggdraw() +
-  # Make background white
-  theme(plot.background = element_rect(fill = "white", color = NA)) +
-  # Add the final plot but with reduced size for the legend and "title"
-  draw_plot(mass_grid, x = 0, y = 0, width = 1, height = 0.92) +
-  # Insert species name
-  annotate("text", x = 0.36, y = 0.99, label = "Lampropholis delicata", hjust = 1, vjust = 1, size = 5, fontface = "italic") +
-  annotate("text", x = 0.9, y = 0.99, label = "Lampropholis guichenoti", hjust = 1, vjust = 1, size = 5, fontface = "italic") +
-  # Insert title for each plot
-  annotate("text", x = 0.023, y = 0.91, label = "A", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
-  annotate("text", x = 0.535, y = 0.91, label = "B", hjust = 1, vjust = 1, size = 6, fontface = "bold") +
-  # Insert picture species
-  draw_image(here("./Others/Deli.png"), x = 0.271, y = 0.65, width = 0.22, height = 0.22) +
-  draw_image(here("./Others/Guich.png"), x = 0.78, y = 0.65, width = 0.22, height = 0.22) +
-  # Insert legends
-  draw_grob(legend_deli, x = 0.375, y = 0.275, width = 0.01, height = 0.01) +
-  draw_grob(legend_guich, x = 0.88, y = 0.275, width = 0.01, height = 0.01)
-#
-ggsave("./output/figures/main/fig_mass.png", plot=fig_mass, width = 21, height = 8, units = "cm", dpi = 600)
-knitr::include_graphics("./output/figures/main/fig_mass.png")
-```
-
-#### Do both species differ in their response to a stressor?
-
-```{r, contrasts_sp}
-#| label: contrasts_sp
-# Chunk created to make the df for the table below (Table 1) and also facilitate the comparisons between species below
-#
-contrasts_sp <- bind_rows(move_deli = deli_move_posteriors,
-                          move_guich = guich_move_posteriors,
-                          shelter_deli = deli_shelter_posteriors,
-                          shelter_guich = guich_shelter_posteriors,
-                          emergence_deli = deli_emergence_posteriors,
-                          emergence_guich = guich_emergence_posteriors,
-                          mass_deli = deli_mass_posteriors,
-                          mass_guich = guich_mass_posteriors,
-                          .id = "data") %>%
-                pivot_longer(cols = -c(data),
-                            names_to = c("stress", "treatment"),
-                            names_pattern = "(int|slope)_(.*)") %>%
-                # Pivot to wide format
-                mutate(column_id = paste(data, stress, treatment, sep = "_")) %>%
-                group_by(column_id) %>%
-                mutate(draw = row_number()) %>%
-                ungroup() %>%
-                dplyr::select(column_id, draw, value) %>%
-                pivot_wider(names_from = column_id, values_from = value) %>%
-
-                #Get the real values
-                mutate(across(matches("emergence.*_int"), ~ plogis(.x)),
-                      across(matches("(move|shelter).*_int"), ~ exp(.x)),
-                      across(matches("mass.*_int"), ~ exp(.x) + min(mass_clean$delta_mass, na.rm = TRUE) - 1)) %>%
-data.frame()
-```
-
-_Acute stressor_: The probability of emerging from the shelter after an acute stressor was lower in cold incubated _L.
-delicata_ than _L.
-guichenoti_ for both control (I~delicata~ - I~guichenoti~ = `r format_dec(median(contrasts_sp$emergence_deli_int_Control_Cold) - median(contrasts_sp$emergence_guich_int_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(contrasts_sp$emergence_deli_int_Control_Cold - contrasts_sp$emergence_guich_int_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(contrasts_sp$emergence_deli_int_Control_Cold - contrasts_sp$emergence_guich_int_Control_Cold, credMass = 0.95)[2], 3)`]) and CORT-treated (I~delicata~ - I~guichenoti~ = `r format_dec(median(contrasts_sp$emergence_deli_int_CORT_Cold) - median(contrasts_sp$emergence_guich_int_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(contrasts_sp$emergence_deli_int_CORT_Cold - contrasts_sp$emergence_guich_int_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(contrasts_sp$emergence_deli_int_CORT_Cold - contrasts_sp$emergence_guich_int_CORT_Cold, credMass = 0.95)[2], 3)`]) lizards, but not in hot incubated animals (see @table_contrasts_species).
-There were no differences between species in the latency to move or the latency until sheltering after an acute stressor (see @table_contrasts_species).
-
-_Chronic stressor_: Over time, _L.
-delicata_ CORT-treated lizards incubated at cold temperatures increased latency to move more than _L.
-guichenoti_ (β~delicata~ - β~guichenoti~ = `r format_dec(median(deli_move_posteriors$slope_CORT_Cold) - median(guich_move_posteriors$slope_CORT_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_move_posteriors$slope_CORT_Cold - guich_move_posteriors$slope_CORT_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
-The probability of emerging from the shelter increased over time in _L.
-delicata_ more than in _L.
-guichenoti_ (see @table_contrasts_species), but those differences were significant only in the Control-Cold treatment (β~delicata~ - β~guichenoti~ = `r format_dec(median(deli_emergence_posteriors$slope_Control_Cold) - median(guich_emergence_posteriors$slope_Control_Cold), 3)`; 95% HDPI = [`r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[1], 3)`, `r format_dec(hdi(deli_emergence_posteriors$slope_Control_Cold - guich_emergence_posteriors$slope_Control_Cold, credMass = 0.95)[2], 3)`]), while the rest of the treatments did not differ between species (see @table_contrasts_species).
-There were no significant differences in the changes on latency to shelter between species (see @table_contrasts_species).
-
-_Mass change_: increases in mass were not different between species in any of the treatments (see @table_contrasts_species).
-
-_Table 1.
-Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L.
-guichenoti.
-Contrast represent differences in median parameters between species (L.
-delicata - L.
-guichenoti) for each treatment.
-95% Highest Posterior Density Intervals (95% HPDIs) indicate the probability that contrasts differ from zero._
-
-```{r, table_contrasts_species}
-#| label: table_contrasts_species
-#| table.caption: "Contrasts of posterior distributions for intercepts and slopes of behavioral variables across treatments, comparing Lampropholis delicata and L. guichenoti."
-source(here("R", "func.R"))
-#
-# Make loop to get contrasts
-var <- c("move", "shelter", "emergence", "mass")
-stress <- c("int", "slope")
-treatment <- c("Control_Cold", "CORT_Cold", "Control_Hot", "CORT_Hot")
-contrasts_df <- data.frame()
-for(v in var){
-  for(st in stress){
-    for(t in treatment){
-      data_loop <- data.frame()
-
-      col_deli <- paste0(v, "_deli_", st, "_", t)
-      col_guich <- paste0(v, "_guich_", st, "_", t)
-      #
-      vector_deli <- contrasts_sp %>% pull(!!sym(col_deli))
-      vector_guich <- contrasts_sp %>% pull(!!sym(col_guich))
-
-      contrasts_median <- format_dec(median(vector_deli) - median(vector_guich), 3)
-      contrasts_lowlim <- format_dec(hdi(vector_deli - vector_guich, credMass = 0.95)[1], 3)
-      contrasts_uplim <- format_dec(hdi(vector_deli - vector_guich, credMass = 0.95)[2], 3)
-
-      data_loop <- data.frame(variable = v,
-                              stress = st,
-                              treatment = t,
-                              contrast_median = contrasts_median,
-                              low_lim = contrasts_lowlim,
-                              up_lim = contrasts_uplim)
-
-      contrasts_df <- bind_rows(contrasts_df, data_loop)
-    }
-  }
-}
-# Modify the contrast_df (above) to get it ready for the table
-contrasts_df2 <- contrasts_df %>%
-  pivot_wider(names_from = stress,
-              values_from = c(contrast_median, low_lim, up_lim)) %>%
-  mutate(treatment = str_replace_all(treatment, "_", "-"),
-        treatment = factor(treatment,
-                          levels = c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot"))) %>%
-  mutate(HPDI_int = paste0("[", low_lim_int, ", ", up_lim_int, "]"),
-         HPDI_slope = paste0("[", low_lim_slope, ", ", up_lim_slope, "]")) %>%
-  dplyr::select(variable, treatment, contrast_median_int, HPDI_int, 
-                contrast_median_slope, HPDI_slope) %>%
-data.frame()       
-write.csv(contrasts_df2, here("./output/checking/contrasts_species.csv"))
-# Set defaults for the table
-set_flextable_defaults(
-  font.family = "Times New Roman",
-  font.size = 10)
-#
-# Make Table
-table_contrasts_species <- flextable(contrasts_df2) %>%
-  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
-  flextable::compose(i = 1, j = 2, value = as_paragraph("Treatment"), part = "header") %>% # Change the second column header
-  flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
-  flextable::compose(i = 1, j = c(4,6), value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
-  flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
-  flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::compose(i = 1, j = 1, value = as_paragraph("Latency to move (s)"), part = "body") %>% 
-  flextable::compose(i = 5, j = 1, value = as_paragraph("Latency to shelter (s)"), part = "body") %>% 
-  flextable::compose(i = 9, j = 1, value = as_paragraph("Emergence"), part = "body") %>% 
-  flextable::compose(i = 13, j = 1, value = as_paragraph("Δmass (mg)"), part = "body") %>% # Change the first column's names
-  flextable::compose(i = c(13:16), j = c(5,6), value = as_paragraph("-"), part = "body") %>%
-  flextable::hline(i = c(4,8,12), part = "body") %>% # Add horizontal lines
-  flextable::vline(j = c(2,4), part = "all") %>% # Add vertical lines
-  align(align = "center", j = c(3:6), part = "body") %>%
-  align(align = "center", i = c(1,2), j = c(1:6), part = "header") %>% 
-  flextable::bold(i = c(9,10), j = c(3,4), bold = TRUE, part = "body") %>% 
-  flextable::bold(i = c(2,9), j = c(5,6), bold = TRUE, part = "body") %>%
-autofit()
-#
-table_contrasts_species
 ```
@@
 
-_Table S4. Contrasts for mass modelled for both species. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
-
-```{r, table_contrasts_mass}
-#| label: table_contrasts_mass
-#
-source(here("R", "func.R"))
-#
-contrasts_mass <- bind_rows(deli = deli_mass_contrasts,
-                          guich = guich_mass_contrasts,
-                          .id = "Species") %>%
-                    mutate(across(everything(), as.character)) %>%
-
-                    rename_with(~ str_replace_all(., "X2\\.5\\.HDI_", "low_lim_")) %>%
-                    rename_with(~ str_replace_all(., "X97\\.5\\.HDI_", "up_lim_")) %>%
-
-                    pivot_longer(cols = -Species,
-                                names_to = c("type", "factor"),
-                                names_pattern = "(median|low_lim|up_lim)_(.*)") %>%
-
-                    pivot_wider(names_from = c(type),
-                                values_from = value) %>%
-
-                    mutate(CI = paste0("[", low_lim, ", ", up_lim, "]")) %>%
-
-                    dplyr::select(Species, factor, median, CI) %>%
-data.frame()
-write.csv(contrasts_mass, here("./output/checking/contrasts_mass.csv"))
-#
-# Set defaults for the table
-set_flextable_defaults(
-  font.family = "Times New Roman",
-  font.size = 10)
-#
-# Make Table
-table_mass <- flextable(contrasts_mass) %>%
-  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
-  flextable::compose(i = 1, j = 2, value = as_paragraph("Predictor"), part = "header") %>% # Change the second column header
-  flextable::compose(i = 1, j = 3, value = as_paragraph("Contrast"), part = "header") %>% # Change the third column header
-  flextable::compose(i = 1, j = 4, value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
-  flextable::compose(i = c(2,3,5,6), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::compose(i = 1, j = 1, value = as_paragraph(as_i("L. delicata")), part = "body") %>% # Change the first row of the first column
-  flextable::compose(i = 4, j = 1, value = as_paragraph(as_i("L. guichenoti")), part = "body") %>% # Change the fourth row of the first column
-  flextable::hline(i = 3, part = "body") %>% # Add horizontal lines
-  align(align = "center", j = c(3,4), part = "body") %>%
-  align(align = "center", j = c(1:4), part = "header") %>% 
-autofit()
-#
-table_mass
-```
-
-Contrasts were done by:  
-    - _Temperature_: β~Hot~ - β~Cold~  
-    - _CORT_: β~CORT~ - β~Control~  
-    - _Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
 ```{r, clean_posteriors_tables}
@@
 #
-variables <- c("move", "shelter", "emergence", "mass")
+variables <- c("move", "shelter", "emergence")
 species <- c("deli", "guich")
@@
 
-_Table S5. Estimates of each predictor for each variable modelled for L. delicata. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
+_Table S4. Estimates of each predictor for each variable modelled for L. delicata. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
   align(align = "center", j = c(1:4), part = "header") %>%
-  flextable::compose(i = c(2:8,10:16,18:24,26:29), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
-  bold(i = c(1,3,7,9,17,18,25,26), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
+  flextable::compose(i = c(2:9,11:18,20:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(9,18), part = "body") %>% # Add horizontal lines
+  bold(i = c(1,3,8,10,19,20,22), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
   autofit()
@@
 
-_Table S6. Estimates of each predictor for each variable modelled for L. guichenoti. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
+_Table S5. Estimates of each predictor for each variable modelled for L. guichenoti. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
   align(align = "center", j = c(1:4), part = "header") %>% 
-  flextable::compose(i = c(2:8,10:16,18:24,26:29), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(8,16,24), part = "body") %>% # Add horizontal lines
-  bold(i = c(1,9,23,24,25,26), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
+  flextable::compose(i = c(2:9,11:18,20:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(9,18), part = "body") %>% # Add horizontal lines
+  bold(i = c(1,10,26,27), j = c(2:4), bold = TRUE, part = "body") %>% # Bold significant values
   autofit()

### 2025-10-18 R/func.r

@@ -168,41 +168,30 @@ tidy_post <- function(df) {
 # Function to get estimates to make the contrasts between treatments
 #' @title post_values
 #' @param df To select the df
-#' @param var To indicate the kind of variable to use (beh or mass)
-post_values <- function(df, var){
+post_values <- function(df){
   df <- as.data.frame(df)
-  if(var == "beh"){
-    int_Control_Cold <- df$b_Intercept
-    int_CORT_Cold <- df$b_Intercept + df$b_trtCORTMCold
-    int_Control_Hot <- df$b_Intercept + df$b_trtControlMHot
-    int_CORT_Hot <- df$b_Intercept + df$b_trtCORTMHot
-    slope_Control_Cold <- df$b_day
-    slope_CORT_Cold <- df$b_day + df$`b_day:trtCORTMCold`
-    slope_Control_Hot <- df$b_day + df$`b_day:trtControlMHot`
-    slope_CORT_Hot <- df$b_day + df$`b_day:trtCORTMHot`    
-    # Create a data frame with the estimates
-    data_values <- data.frame(
-      int_Control_Cold = int_Control_Cold,
-      int_CORT_Cold = int_CORT_Cold,
-      int_Control_Hot = int_Control_Hot,
-      int_CORT_Hot = int_CORT_Hot,
-      slope_Control_Cold = slope_Control_Cold,
-      slope_CORT_Cold = slope_CORT_Cold,
-      slope_Control_Hot = slope_Control_Hot,
-      slope_CORT_Hot = slope_CORT_Hot)
+
+  int_Control_Cold <- df$b_Intercept
+  int_CORT_Cold <- df$b_Intercept + df$b_trtCORTMCold
+  int_Control_Hot <- df$b_Intercept + df$b_trtControlMHot
+  int_CORT_Hot <- df$b_Intercept + df$b_trtCORTMHot
   
-  } else if (var == "mass") {
-    int_Control_Cold <- df$b_Intercept
-    int_CORT_Cold <- df$b_Intercept + df$b_trtCORTMCold
-    int_Control_Hot <- df$b_Intercept + df$b_trtControlMHot
-    int_CORT_Hot <- df$b_Intercept + df$b_trtCORTMHot
-    # Create a data frame with the estimates
-    data_values <- data.frame(
-      int_Control_Cold = int_Control_Cold,
-      int_CORT_Cold = int_CORT_Cold,
[…]

### 2025-10-19 ms.qmd
updates

@@
 ---
-title: "Effects of prenatal CORT and temperature on the antipredatory behaviour in a lizard"
+title: "Antipredatory strategies are robust to prenatal CORT and temperature in two skink species"
 format:
@@
 
-Animals constantly face events that can disrupt their homeostasis.
-These disruptions can be acute, such as a sudden predator attack, or long-term, like prolonged droughts or habitat degradation [@sapolsky_how_2000; @mcewen_concept_2003].
-Regardless of the nature of the challenge, organisms must adjust to allostasis to maintain their physiological and behavioural integrity [@wingfield2002endocrine].
-In vertebrates, the ability to cope with disruptions in homeostasis is carried out by the hypothalamic-pituitary-adrenal/interrenal (HPA/HPI) axis [@sapolsky_how_2000].
-This system elicits a physiological response, also known as the stress response, mediated by glucocorticoids (GCs) - a group of hormones that facilitates the reallocation of energetic resources to restore homeostasis [@sapolsky_how_2000; @mcewen_concept_2003].
-For example, when an animal faces a predatory attack, GCs mobilise energy to support increased locomotor activity, vigilance, or attention to avoid the threat [@trompeter2011invader].
-In the short term, GCs help organisms cope with acute challenges.
-However, long-term exposure to GCs can carry significant metabolic costs, leading to immune suppression, impaired growth, or permanent neural alterations [@picard_mitochondria_2014; @mcewen2017neurobiological].
-Furthermore, reacting to stressors can reduce the time and energy available for essential activities such as foraging, thermoregulation, or social interactions [@belliure2004behavioral; @martin2024blind].
-As such, animals are predicted to evolve mechanisms to appropriately respond to acute stressors while avoiding the costs in the long term.
-For example, by regulating their responses to homotypic (repeated and similar) stressors, particularly those that are predictable and non-lethal [@grissom2009habituation].
-Habituation - the reduction of physiological responses elicited by exposure to a homotypic stressor [@pfister1979glucocorticosterone] - can be crucial to avoid the long-term costs of chronic stress without compromising the response to acute stressors.
+Predation risk is a prominent ecological pressure that can shape the behavior, physiology, and life histories of prey species [@lima1990behavioral; @brown1999ecology].
+To reduce the likelihood of being detected, attacked or captured, prey animals have developed a suite of antipredatory strategies, including morphological defences, behavioral tactics, and physiological adjustments [@stankowich2005fear].
+However, these strategies are usually costly, leading to trade-offs between antipredatory avoidance and other fitness-related traits [@lima1990behavioral; @stankowich2005fear].
+To optimize benefits over costs, animals must carefully evaluate predation risks and respond accurately to them, adjusting their responses to changing risk levels [@lima1990behavioral; @chen2025zebra].
+This plasticity allows animals to balance the costs and benefits of avoiding predation in dynamic environments [@amo2005flexibility; @chen2025zebra].
 
-Importantly, the process of physiological habituation shares key functional characteristics with the cognitive process of habituation - a form of non-associative learning where individuals reduce their responsiveness to a repeated stimulus [@thompson1966habituation].
-In both cases, organisms are repeatedly exposed to a stimulus/homotypic stressor and progressively decrease their response toward it [@grissom2009habituation].
-In fact, both behavioural and physiological habituation often co-occur.
-For example, rhesus macaques (_Macaca mulatta_) showed a decline in blood cortisol concentrations across consecutive days of restraint, accompanied by decreased vocalizations and behavioral agitation [@ruys2004behavioral].
-However, physiological habituation alone is insufficient to explain the behavioural patterns observed when animals are repeatedly exposed to a processive stressors - those involving primarily psychological qualities, such as restraint or simulated predation  [@jaferi2006corticosterone; @grissom2009habituation].
-For example, adrenalectomised rats - incapable of physiological habituation - still exhibit cognitive habituation [@jaferi2006corticosterone], suggesting a degree of independence between the two processes.
-While the HPA/HPI axis is paramount for physiological habituation, cognitive habituation also involves neural mechanisms related to sensory processing, learning, and behavioural control, such as the limbic system, the superior colliculus, or the optic tectum [@grissom2009habituation; @dutta2014saliency].
-Habituation to processive stressors is therefore likely to involve complex interactions between physiological and cognitive processes.
+However, there is considerable variation in how animals cope with predation risk [@stankowich2005fear; @thaker2009alternative; @horvath2020ain].
+Some individuals exhibit strong antipredatory responses or adjust more accurately their strategies under changing risk levels than others [@stankowich2005fear; @thaker2009alternative].
+Cognition and the stress response system can be key factors predicting individual differences in the antipredatory response [@griffin2001learning; @kelley2003learned; @shier2007effects; @thaker2009acute; @thaker2010stress].
+First, cognitive proceeses like predator detection and accurate decision-making are essential for avoiding predation, while behaving flexibly is fundamental to optimize antipredatory responses under different circumstances [@griffin2001learning; @kelley2003learned].
+Second, the stress response system - mediated by the hypothalamic-pituitary-adrenal (HPA) or hypothalamic-pituitary-interrenal (HPI) axis - plays a crucial role in mediating behavioral and physiological responses to predation [@thaker2009acute; @thaker2010stress].
+Under threat, the HPA/HPI axis triggers the release of glucocorticoids (GCs), which help mobilize resources to cope with imminent risk [@sapolsky_how_2000].
+Therefore, cognitive processes and the stress response system likely interact to determine how animals can cope with predation risk.
+As such, investigating the factors that influence cognitive development and HPA/HPI function is fundamental to understanding individual differences in antipredatory responses and their plasticity.
 
-However, the effectiveness of either system depends on how well the underlying neural structures develop and integrate.
-Many of these brain regions mature concurrently and can be highly plastic during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017].
-Early-life conditions can shape the structure and function of both neural systems, with long-term consequences for how individuals respond to acute and chronic stressors [@vargas2016early; @van2017modulation; @zito2017early; @crino_under_2020; @crino2024eggs].
-One of the key mediators of early-life effects is GC exposure.
-Elevations in GCs during development exert potent effects on HPA/HPI axis function, altering the response to stressors later in life [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs].
-For example, zebra finches (_Taeniopygia guttata_) nutritionally stressed during early-life had higher levels of CORT than controls at post-hatching day 25, while periodic maternal separation increased basal corticosterone levels in rats (_Rattus norvegicus_) [@vargas2016early; @zito2017early].
-In parallel, developmental GC exposure can impact brain regions involved in learning, memory, and behavioural flexibility, affecting neural processes relevant to cognitive habituation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta].
-Therefore, early-life exposure to GCs can have long-lasting effects on the response to both acute and chronic stressors, potentially leading to maladaptive responses in adulthood.
+Early-life environment is critical for the development of both cognitive function and the HPA/HPI axis [@lemaire_prenatal_2000; @jafari2017prenatal; @abayarathna_effects_2020; @crino2024eggs].
+Stressful conditions during development can lead to long-lasting alterations in cognitive abilities and GCs regulation [@lemaire_prenatal_2000; @zhu_prenatal_2004; @jafari2017prenatal; @crino2024eggs].
+Prenatal stress or GC exposure can impact neural processes relevant to perception or behavioral flexibility [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta; @meunier2020olfactory], or alter the HPA/HPI axis function leading to changes in stress responsiveness and coping strategies [@koolhaas1999coping; @jafari2017prenatal].
+Although evidence is mixed, most studies suggest that prenatal GC elevations produce individuals that are more reactive to threats and less behaviorally flexible [@koolhaas1999coping; but see @recio2025early].
 
-Other environmental factors may also shape stress-related phenotypes, either independently or in combination with GCs.
-In ectotherms, early thermal conditions play a critical role in cognitive development [@amiel_effects_2017; @abayarathna_effects_2020].
-Incubation temperature affects cognitive abilities such as learning, memory, or perception, likely through changes in brain function and structure [@amiel_smart_2011; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017; @abayarathna_effects_2020].
-In addition, early thermal environment can also affect the development of the HPA/HPI axis [@crino_under_2020; @crino2024eggs].
-For example, in the delicata skink (_Lampropholis delicata_), individuals incubated at cooler temperatures showed increased baseline CORT concentration in the blood [@crino2024eggs].
-Although these differences are believed to be driven by size rather than direct endocrine effects [@crino2024eggs], they can still have important long-term consequences for how individuals respond to acute and chronic stressors.
-Importantly, incubation temperature and prenatal CORT may not act independently [see @wingfield_comparative_2008; @Crino_2023], and their interaction could produce complex, trait-specific outcomes that shape both physiological and behavioural responses to stress.
+However, other aspects of the early-life environment can also influence cognitive processes or modulate the effects of GC exposure [@amiel_smart_2011; @dayananda_incubation_2017; @abayarathna_effects_2020].
+Factors that, as prenatal GC exposure, can alter metabolic function during development, typically have significant long-lasting effects on brain function and cognition [@amiel_effects_2017].
+In this regard, temperature is a prominent environmental factor influencing metabolic rates in ectotherms [@schulte2015effects; @noble_developmental_2018].
+As such, prenatal thermal environment shapes multiple traits in ectotherms, including cognitive processes [@amiel_smart_2011; @dayananda_incubation_2017; @noble_developmental_2018; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019; @abayarathna_effects_2020].
+For example, incubation at higher temperatures in several species of lizards is associated with improved cognition and more dense neural structures [@amiel_smart_2011; @amiel_hotter_2012; @amiel_effects_2017].
+Therefore, prenatal thermal environment can influence antipredatory behavior and its flexibility through effects on cognitive processes [see @burger_antipredator_1998].
 
-Because of the functional similarities between physiological and cognitive habituation, behavioural responses can offer insight into an animal’s ability to regulate stress over time, even in the absence of physiological measures.
-This is particularly relevant when investigating early-life programming, which can shape both HPA/I development and brain regions involved in behavioural inhibition and learning [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta; @amiel_hotter_2012].
-In this study, we investigated how prenatal CORT and incubation temperature shape the response to a stressor in two species of skinks: the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
+Since both temperature and GC exposure can influence the antipredatory response through similar mechanisms, we could expect prenatal GCs and temperature have interactive effects on antipredatory strategies.
+Importantly, this interaction is likely ecologically relevant, as extreme environmental temperatures experienced by mothers can increase GCs transmitted to offspring [@racic_effects_2020; @Crino_2023].
+However, the combined effects of prenatal GC exposure and incubation temperature on antipredatory strategies remain poorly understood.
+
+The aim of the present study is to examine how prenatal conditions - particularly prenatal GCs exposure and thermal environment - can shape the antipredatory response and its flexibility in ectotherms.
+We investigated the combined effects of prenatal corticosterone (CORT) - the primary GC in reptiles - and incubation temperature on the antipredatory behavior and the antipredatory behavioral flexibility in two closely related skink species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
 guichenoti_).
-We focused on the effects of these early-life conditions on the behavioural responses to a processive stressor.
-Specifically, we examined how these factors influence the acute (i.e., the first exposure) and the chronic (i.e., repeated exposures over time) responses by measuring freezing behaviour, shelter seeking, and shelter emergence after a simulated predatory attack.
-Finally, we assessed body mass change across the stress period as a potential indirect cost, while controlling for food intake.
-We predicted that prenatal CORT exposure would heighten acute stress reactivity and slow habituation, while warmer incubation temperatures would reduce acute responses and enhance habituation.
-Furthermore, we predicted higher temperatures would buffer the effects of prenatal CORT on stress responses, potentially leading to more flexible behavioural adjustment.
-Finally, we expected that less flexible stress responses would lead to higher decreases in mass.
+We experimentally manipulated CORT levels in eggs and incubated them at either cold or hot temperatures.
+After hatching, we assessed the lizards' antipredatory response and their ability to adjust this behavior over repeated exposures to a simulated predator attack.
+Given the known effects of prenatal GCs and temperature on brain development and HPA/HPI function [@jafari2017prenatal; @lemaire_prenatal_2000; @abayarathna_effects_2020; @crino2024eggs], we hypothesized that prenatal CORT exposure would enhance antipredatory response.
+However, it may reduce antipredatory behavioral flexibility.
+Additionally, we predicted that hot incubation temperatures would improve behavioral adjustments and buffer the effects of prenatal CORT on antipredatory behavior [@amiel_smart_2011; @amiel_hotter_2012; @amiel_effects_2017].
+
+Finally, we explored potential species differences in the effects of prenatal conditions on antipredatory strategies.
+In this sense, although both species share similar life-histories, habitats, and distributions [@chapple_know_2011; @chapple_biology_2014], only _L.
+delicata_ has been reported as a successful coloniser outside their natural range [@chapple_know_2011; @chapple_biology_2014].
+These differences in the ability to colonise new environments are not related to cognitive abilities [@chapple_know_2011; @recio2025cognitive; @recio2025early], but could be associated with differences in personality [@chapple_know_2011], which may be linked to different antipredatory strategies [@amo2005flexibility; @horvath2020ain].
 
 ## Methods
-
-#### Subjects
-
-We used two species of closely skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
-guichenoti_).
-Both species are oviparous, generalist, small (∼35-55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia [@chapple_know_2011; @chapple_biology_2014].
-However, while _L.
-delicata_ has been reported as a successful invader in New Zealand and Hawaii, there are no such reports for _L.
-guichenoti_ [@chapple_know_2011; @chapple_biology_2014].
-Differences in the ability to colonise new environments are not related to cognitive abilities [@bezzina2014does; @recio2025cognitive; @recio2025early], but could be associated with differences in personality [@chapple_know_2011], which may be linked to different responses to  predatory stimuli [**REFS**].
 
@@
 
-Three weeks before the start of the behavioural tests, we adapted the lizard enclosures to the experimental setup.
+Three weeks before the start of the behavioral tests, we adapted the lizard enclosures to the experimental setup.
 The enclosures used for the antipredatory response trials were the same as those where the lizards had previously been housed, but modified to include only a single shelter (9 × 6 × 1.5 cm) placed at one end and a water dish in the centre.
@@
 The lizard was then chased for 60 seconds using a soft paintbrush, simulating the predatory attack.
-Immediately after the chase, the shelter — but not the water dish — was returned to its original position, marking the start of the behavioural observation period.
-The lizard’s behaviour was then recorded for one hour.
-All videos were later analysed by MD, who was blinded to the lizards’ treatment.
-From each video, we recorded three behavioural variables: i) latency to move — the time spent immobile immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) probability to emerge — whether or not the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
+Immediately after the chase, the shelter — but not the water dish — was returned to its original position, marking the start of the behavioral observation period.
+The lizard’s behavior was then recorded for one hour.
+All videos were later analyzed by MD, who was blinded to the lizards’ treatment.
+From each video, we recorded three behavioral variables: i) latency to move — the time spent immobile immediately after the simulated attack; ii) latency to shelter — the time it took the lizard to hide under the shelter after resuming movement; and iii) probability to emerge — whether or not the lizard emerged from the shelter during the 40 minutes following hiding (1 = emerged, 0 = did not emerge).
 A lizard was considered to have resumed movement if it moved continuously for at least five seconds, and to be under shelter if no limbs were visible.
@@
 In panel A, the early environment manipulations leading to four experimental conditions.
-In panel B, the postnatal antpredatory response assay, for 8 days, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioural recording.
+In panel B, the postnatal antpredatory response assay, for 8 days, lizards were subjected daily to a simulated predator attack (chased with a soft paintbrush for 1 minute), followed by a 1-hour behavioral recording.
 During this recording, we measured the latency to move, the latency to shelter, and the probability of emergening from the shelter in 40 minutes after the stressor."
@@
 We included the prenatal condition (Control-Cold, CORT-Cold, Control-Hot, CORT-Hot), the day of the trial, and their interaction as fixed effects.
-To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model (i.e., the intercept reflects lizard behaviour in response to the first exposure to the stressor).
-We also included snout-vent length - measured one day before the beginning of the trials - to control for potential effects of size on the antipredatory behaviour [**REFs**].
+To facilitate interpretation of the model intercepts, trial day was transformed so that day 1 in the experiment corresponds to day 0 in the model.
+We also included snout-vent length - measured one day before the beginning of the trials - to control for potential effects of size on the antipredatory behavior [@stankowich2005fear].
 SVL was centered to the mean to facilitate interpretation of the intercepts.
@@
 We used the posterior distributions of model parameters to test for differences between prenatal conditions.
-Differences in the model intercepts between conditions were interpreted as differences in the immediate antipredatory behaviour, while differences in the slopes were interpreted as differences in antipredatory adjustments.
+Differences in the model intercepts between conditions were interpreted as differences in the immediate antipredatory behavior, while differences in the slopes were interpreted as differences in antipredatory flexibility.
 We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments or the slopes for each treatment were different from zero.
@@
 
-All the estimated parameters in the results were transformed back to original units employing appropriate formulas depending on the distribution employed in the models.
+All the estimated parameters in the results were transformed back to original units unless otherwise indicated.
+To back-transformed the results, we employed appropriate formulas depending on the distribution used in the models.
 The error structure for latency to move and latency to shelter followed a lognormal distribution [family = lognormal()], while the probability of emerging from the shelter was modelled using a Bernoulli distribution with a logit link function [family = Bernoulli(link = 'logit')].
@@
 
-```{r, models_behaviour}
-#| label: models_behaviour
+```{r, models_behavior}
+#| label: models_behavior
 source(here("R", "func.R"))
@@
 
-_Lampropholis delicata_: Incubation at cold temperatures decreased the latency to move after a simulated attack on the first day (Intercept (denoted hereafter as I)~Hot~ - I~Cold~ = `r deli_move_contrasts$median_Temperature_int`; 95% Highest Posterior Density Intervals (95% HPDI) = [`r deli_move_contrasts$X2.5.HDI_Temperature_int`, `r deli_move_contrasts$X97.5.HDI_Temperature_int`]), while neither CORT or the interaction have any effect on latency to shelter or emergence (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
-We did not find any effects of prenatal temperature, CORT or their interaction on latency to shelter or emergence probability on the first day of the assay (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
-SVL had no significant effect on any of the behavioural responses (see Table S4 in _Supplementary Material_).
+_Lampropholis delicata_: Neither prenatal CORT, temperature, nor their interaction had a significant effect on the antipredatory response on the antipredatory behavior of _L.
+delicata_ (see @fig-results_beh_deli A, D, G and Table S1 in _Supplementary Material_).
+SVL had no significant effect on any of the behavioral responses (see Table S4 in _Supplementary Material_).
 
-_Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had a significant effect on the immediate antipredatory response on _L.
+_Lampropholis guichenoti_: Neither prenatal CORT, temperature, nor their interaction had a significant effect on the antipredatory response on _L.
 guichenoti_ (see @fig-results_beh_guich A, D, G and Table S2 in _Supplementary Material_).
-We did not see any significant effects of SVL on any of the behavioural responses (see Table S5 in _Supplementary Material_).
+We did not see any significant effects of SVL on any of the behavioral responses (see Table S5 in _Supplementary Material_).
 
@@
 #| label: fig-results_beh_deli
-#| fig.cap: "Behavioural response of _Lampropholis delicata_ to a stressor.
-Panels A, D, and G show the distribution of the estimates for all the behavioural responses after the first simulated attack per each treatment (i.e., response to acute stressor).
+#| fig.cap: "behavioral response of _Lampropholis delicata_ to a stressor.
+Panels A, D, and G show the distribution of the estimates for all the behavioral responses after the first simulated attack per each treatment (i.e., response to acute stressor).
 In panels B, E, and H the distribution of the estimates of slopes per each treatment (i.e., response to chronic stressor).
@@
 no response to chronic stress).
-Panels C, F, and I show the behavioural estimated values over trials.
+Panels C, F, and I show the behavioral estimated values over trials.
 The lines represent the median estimated value for each trial, and the shaded areas indicate the 95% HPDI; both were obtained using the slope and intercept estimates from the posterior distributions.
@@
 
-#### Do prenatal conditions affect the long-term antipredatory response?
+#### Do prenatal conditions affect the antipredatory flexibility?
 
@@
 
-_Lampropholis guichenoti_: Behavioral responses towards a stressor did not change over time in _L.
-guichenoti_ (see @fig-results_beh_guich B, C, E, F, H, I and Table S3 in _Supplementary Material_).
-There were no significant differences in the behavioural response of _L.
-guichenoti_ to the repeated exposure to a stressor (see @fig-results_beh_guich B, C, E, F, H, I and Table S2 in _Supplementary Material_).
+_Lampropholis guichenoti_: We did not see any antipredatory flexibility in _L.
+guichenoti_, reflected by slopes being no different from 0 (see @fig-results_beh_guich B, C, E, F, H, I and Table S3 in _Supplementary Material_).
 
@@
 #| label: fig-results_beh_guich
-#| fig.cap: "Behavioural response of _Lampropholis guichenoti_ to a stressor.
-Panels A, D, and G show the distribution of the estimates for all the behavioural responses after the first simulated attack per each treatment (i.e., response to acute stressor).
+#| fig.cap: "behavioral response of _Lampropholis guichenoti_ to a stressor.
+Panels A, D, and G show the distribution of the estimates for all the behavioral responses after the first simulated attack per each treatment (i.e., response to acute stressor).
 In panels B, E, and H the distribution of the estimates of slopes per each treatment (i.e., response to chronic stressor).
@@
 
+#### Do species differ in their antipredatory response and flexibility?
+
+Neither the immediate antipredatory response nor the antipredatory flexibility differed between species for any of the treatments and behavioral responses analyzed (see @table-species-contrasts).
+
+_Table 1.
+Contrasts between species for the intercepts (Response) and slopes (Flexibility) of all the behavioral responses analyzed.
+Values represent the median (Contrast) and 95% Highest Posterior Density Intervals (95% HPDI) of the posterior distributions of the contrasts.
+For all contrasts, we calculated the differences betwee Lampropholis delicata - Lampropholis guichenoti.
+We did not transform the values back to original units.
+Significant differences are indicated in bold._
+```{r, table-species-contrasts}
+#| label: table-species-contrasts
+#| tbl-cap: "Contrasts between species for the intercepts (Response) and slopes (Flexibility) of all the behavioral responses analyzed.
+Values represent the median (Contrast) and 95% Highest Posterior Density Intervals (95% HPDI) of the posterior distributions of the contrasts.
+For all contrasts, we calculated Lampropholis delicata - Lampropholis guichenoti.
+We did not transform the values back to original units.
+Significant differences are indicated in bold."
+source(here("R", "func.R"))
+#
+# Getting the contrasts between species (all contrasts are L. delicata - L. guichenoti)
+var <- c("move", "shelter", "emergence")
+type <- c("int", "slope")
+
+species_contrasts_df <- data.frame()
+
+for(v in var){
+  for(t in type){
+
+    value_c1 <- c(
+      get(paste0("deli_", v, "_posteriors"))[[paste0(t, "_Control_Cold")]],
+      get(paste0("deli_", v, "_posteriors"))[[paste0(t, "_CORT_Cold")]],
+      get(paste0("deli_", v, "_posteriors"))[[paste0(t, "_Control_Hot")]],
+      get(paste0("deli_", v, "_posteriors"))[[paste0(t, "_CORT_Hot")]]
+    )
+
+    value_c2 <- c(
+      get(paste0("guich_", v, "_posteriors"))[[paste0(t, "_Control_Cold")]],
+      get(paste0("guich_", v, "_posteriors"))[[paste0(t, "_CORT_Cold")]],
+      get(paste0("guich_", v, "_posteriors"))[[paste0(t, "_Control_Hot")]],
+      get(paste0("guich_", v, "_posteriors"))[[paste0(t, "_CORT_Hot")]]
+    )
+
+    contrast_values <- value_c1 - value_c2
+    median_contrast <- format_dec(median(contrast_values), 3)
+    hdi_low <- format_dec(hdi(contrast_values, credMass = 0.95)[1], 3)
+    hdi_high <- format_dec(hdi(contrast_values, credMass = 0.95)[2], 3)
+
+    temp_df <- data.frame(
+      Behavior = v,
+      Type = t,
+      Median = median_contrast,
+      HPDI_Lower = hdi_low,
+      HPDI_Upper = hdi_high
+    )
+
+    species_contrasts_df <- rbind(species_contrasts_df, temp_df)
+  }
+} 
+#
+# Improving database
+species_contrasts_clean <- species_contrasts_df %>%
+  mutate(
+    Behavior = case_when(
+      Behavior == "move" ~ "Latency to move",
+      Behavior == "shelter" ~ "Latency to shelter",
+      Behavior == "emergence" ~ "Probability to emerge"),
+    Type = case_when(
+      Type == "int" ~ "Response",
+      Type == "slope" ~ "Flexibility"),
+    `95% HPDI` = paste0(" [", HPDI_Lower, ", ", HPDI_Upper, "]")) %>%
+  dplyr::select(Type, Behavior, Median, `95% HPDI`) %>%
+  arrange(factor(Type, levels = c("Response", "Flexibility")),
+          factor(Behavior, levels = c("Latency to move", "Latency to shelter", "Probability to emerge"))) %>%
+data.frame()
+# # Set defaults for the table
+set_flextable_defaults(
+  font.family = "Times New Roman",
+  font.size = 10)
+#
+# Make Table
+sp_contrasts_table <-  flextable(species_contrasts_clean) %>%
+  flextable::compose(i = 1, j = 3, value = as_paragraph("Contrast"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 4, value = as_paragraph("95% HPDI"), part = "header") %>% # Change the second column header
+  flextable::compose(i = c(2,3,5,6), j = 1, value = as_paragraph(""), part = "body") %>% # Make some blanks
+  flextable::hline(i = 3, part = "body") %>% # Add horizontal line
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", part = "header") %>%
+autofit()
+#
+sp_contrasts_table
+```
+
 ## Discussion    
-We predicted that elevated CORT levels in eggs or incubating them at colder temperatures would increase lizards reactivity to a stressor both in the first encounter or during repeated exposure.
-Higher reactivity would imply larger latencies to move after the simulated attack, lower times until sheltering, and lower probabilities of emerging from the shelter.
+We predicted that elevating CORT levels in eggs or incubating them at colder temperatures would increase lizards antipredatory response and decrease antipredatory flexibility.
+Higher antipredatory response would imply larger latencies to move after the simulated attack, lower times until sheltering, and lower probabilities of emerging from the shelter.
+And lower antipredatory flexibility would be reflected in a lack of changes in these behaviors over repeated exposures to the simulated attack.
+Additionally, we predicted between species differences in both the the antipredatory behavior and flexibility, and how early conditions would impact their behavior.
+However, our results did not support our predictions.
+Neither prenatal CORT exposure nor incubation temperature influenced the antipredatory response or flexibility in either species.
+Moreover, we did not find significant differences between species in any of the behavioral responses analyzed.
 
-Our study shows that early-life conditions can influence how skinks respond to a processive stressor, but the effects differ between species, the nature of the stressor, and the behaviour evaluated.
-_Lampropholis delicata_ incubated at warmer temperatures were less reactive to a simulated predator attack in the first trial, suggesting that early thermal conditions can shape the behavioural response to an acute stressor.
-In contrast, _L.
-guichenoti_ showed no treatment effects.
-Habituation patterns also differed.
-_Lampropholis delicata_ increased the probability of emerging from the shelter, suggesting some habituation to the stressor.
-However, none of the treatments showed habituation in latency to move or go to shelter.
-No behavioural change was detected in _L.
-guichenoti_ either.
-Finally, we found no evidence of physiological costs: individuals generally maintained or gained body mass, regardless of treatment.
-Together, these findings suggest that incubation temperature and prenatal hormone exposure shape stress-related behaviour in a species-specific manner, potentially reflecting differences in flexibility or stress physiology.
+#### Antipredatory response is robust to prenatal conditions      
+We hypothesized that elevated prenatal CORT and cold incubation temperatures would increase latency to move while decreasing the latency to shelter and the probabilities of emerging from the shelter after a simulated attack.
+However, neither prenatal CORT exposure or incubation temperature influenced the antipredatory response in either species.
 
-#### Cold incubation temperature increases the response to an acute stressor in _L. delicata_ but not in _L. guichenoti_    
-We found that the early-life conditions can impact lizards' response to an acute stressor, but this effects were species- and behaviour-dependent.
-Our results demonstrate that incubation temperature can modulate the behavioural response to an acute stressor in _L.
-delicata_ but not in _L.
+Given the significant effects of prenatal GCs on the development of the HPA/HPI axis reported in other species [@costantini2011meta; @vargas2016early; @jafari2017prenatal; @zito2017early; @crino2024eggs], we predicted prenatal CORT elevations to increase lizards’ reactivity to a predatory attempt.
+However, we found no effects of prenatal CORT elevations on any of the behaviors analyzed, and these results were consistent across species.
+These findings suggest that the antipredatory response is robust to CORT elevations in _L.
+delicata_ and _L.
 guichenoti_.
-Hot-incubated _L.
-delicata_ spent less time immobile following a simulated predatory attack, suggesting a faster recovery from acute stress.
-This result is consistent with the idea that shorter immobility duration (i.e., shorter freezing time) reflects a less reactive behavioural response, often associated to lower HPA/HPI activity [@koolhaas1999coping].
-In _L.
-delicata_, individuals incubated at 23 ºC exhibit elevated baseline CORT [@crino2024eggs], potentially enhancing sensitivity to stress.
-Therefore, the heightened acute response in cold-incubated individuals may be explained by elevated endogenous CORT.
-Alternatively, the higher latency times to move may reflect impaired neural development caused by cold incubation.
-Colder incubation temperatures can disrupt the formation and connectivity of brain regions involved in sensory processing and behavioural control [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017], ultimately leading to slower perception and action planning during freezing [@roelofs2022freezing].
-
-In contrast, incubation temperature had no effect on _L.
-guichenoti_'s response to an acute stressor, suggesting that stress reactivity in this species is more robust to early thermal environment.
-Our findings align with previous studies showing that incubation temperature impacts the antipredatory behaviour of _L.
-delicata_ but not _L.
-guichenoti_ [@iglesias2024maternal].
-Between-species differences on the effect of incubation temperature could reflect differences in history-traits such as egg size.
-Smaller eggs in _L.
-delicata_ may make this species more sensitive to environmental alterations during development, with consequences for their physiology and behaviour later in life [@crino2024eggs; @iglesias2024maternal].
-Although heightened plasticity may enable _L.
-delicata_ to fine-tune its phenotype to local conditions, it could also increase vulnerability to suboptimal environments.
-In contrast, the apparent robustness of _L.
-guichenoti_ might buffer against certain environmental inputs but limit flexibility [@fox2019beyond].
-Future studies should explore how these differing developmental strategies influence fitness across different environments.
-
-#### Prenatal CORT did not influence behavioural response to an acute stressor     
-Given the significant effects of prenatal GCs on the development of the HPA/HPI axis reported in other species [@costantini2011meta; @vargas2016early; @zito2017early; @crino2024eggs], we predicted prenatal CORT elevations to increase lizards' reactivity to acute stressors.
-However, we found no effects of prenatal CORT elevations on any of the behaviours analysed, and these results were consistent across species.
-These findings suggest that the response to acute stressors is robust to CORT elevations.
-
 Alternatively, our CORT manipulation may have failed to elicit significant effects.
 Glucocorticoids are known to have hormetic effects on different traits [@mcewen_concept_2003].
-While our CORT treatment was efficient in elevating yolk CORT concentrations, they did not result in increased baseline CORT in juveniles[@crino2024eggs].
+While our CORT treatment was efficient in elevating yolk CORT concentrations, they did not result in increased baseline CORT in juveniles [@crino2024eggs].
 In contrast, doses at lower concentration did [@crino2024eggs].
-Previous reports show that enhanced responsiveness to acute stressors are usually accompanied by higher baseline GC levels [@koolhaas1999coping; @ruys2004behavioral; @thaker2009acute].
-Therefore, the absence of CORT effects on the acute response may reflect the lack of endocrine differences between treatment groups in _L.
+Previous reports show that enhanced responsiveness to predatory attempts are usually accompanied by higher baseline GC levels [@koolhaas1999coping; @ruys2004behavioral; @thaker2009acute].
+Therefore, the absence of CORT effects on the antipredatory response may reflect the lack of endocrine differences between treatment groups in _L.
 delicata_.
@@
 guichenoti_, as comparable endocrine data are currently lacking.
-Further experiments would benefit from studying how exposing eggs to lower CORT concentrations influence stress-related physiology and behaviours in both species.
+Further experiments would benefit from studying how exposing eggs to lower CORT concentrations influence antipredatory behaviors and stress-related physiology in both species.
 
-#### The effects of prenatal conditions on the response to a repeated stressor are species-dependent   
-We predicted that, over time, lizards would decrease the latency to move while increasing both the latency to shelter and the probability of emerging from the shelter.
-These patterns would reflect a more relaxed response to a processive stressor, indicative of cognitive habituation [@koolhaas1999coping; @grissom2009habituation].
-We also expected that prenatal CORT exposure or cold incubation would impair habituation, given their known effects on the HPI axis and cognitive development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @costantini2011meta; @amiel_hotter_2012; @vargas2016early; @zito2017early; @crino2024eggs].
-Nonetheless, our results show complex patterns were habituation differs across species, treatments, and behavioural measures.
+Incubation temperature impacts antipredatory behavior in other reptiles [@burger_antipredator_1998; @iglesias2024maternal].
+For example, pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were less responsive to a predator - less hissing or striking - than hot incubated ones [@burger_antipredator_1998].
+In contrast, we did not see any effects of the prenatal thermal environment on the antipredatory response in either _L.
+delicata_ or _L.
+guichenoti_.
+Similarly, incubation temperature alone did not affect antipredatory behavior in _L.
+delicata_ or _L.
+guichenoti_ in previous studies [@iglesias2024maternal].
+However, when cold incubation temperature was combined with low maternal investment - simulated by extracting 12.5% of the egg yolk - lizards showed increased antipredatory responses [@iglesias2024maternal].
+These results suggest that incubation temperature alone may not be sufficient to alter antipredatory behavior in these species, but it could interact with other early-life conditions to shape antipredatory strategies.
+Future research should explore how incubation temperature interacts with other prenatal factors to influence antipredatory behavior in these skinks.
 
-Only _L.
-delicata_ showed evidence that early-life conditions influenced the response to a chronic stressor.
-Although slope contrasts did no reveal significant differences between treatments, the CORT-Cold group was the only one where the latency to move changed significantly over time.
-However, instead of the predicted reduction, we found an increase in the time spent frozen after the simulated attack, suggesting sensititazion to the stressor rather than habituation.
-This result may reflect a cumulative effect of prenatal CORT and cold incubation temperatures, even in the absence of a statistically significant interaction.
-In _L.
-delicata_, cold incubation alone increases baseline CORT levels in juveniles [@crino2024eggs].
-Although the high CORT dose used in our experiment did not significantly elevate baseline CORT on its own, it produced intermediate values between controls and a lower dose — the latter of which did significantly elevate baseline levels [@crino2024eggs].
-Therefore, it is possible that the combination of cold incubation and the intermediate endocrine effect of the high CORT dose could together increase HPI axis activity enough to enlarge behavioural reactivity over time [@koolhaas1999coping].
+#### Antipredatory flexibility is robust to prenatal conditions
+We did not find significant effects of prenatal CORT exposure or incubation temperature on antipredatory flexibility.
+Either in _L.
+delicata_, where there was some evidence of flexibility, or in _L.
+guichenoti_, where no flexibility was observed, prenatal conditions had no significant effects on how lizards modified their antipredatory behavior over repeated exposures to the simulated attack.
 
-Alternatively, the combination of CORT and cold temperatures could be impairing habituation trough effects on neural development [see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017].
-Although previous studies found no influence of prenatal CORT or temperature on associative and reversal learning [@recio2025cognitive; @recio2025early], the effect of prenatal conditions can vary by brain region [@coomber_independent_1997; @sakata_neural_2000].
-Possibly, neural structure of regions involved cognitive habituation are more sensitive to early-life conditions than those related to other cognitive domains.
-Further work is needed to test this possibility and identify the specific neural mechanisms linking developmental conditions to stress-related behavioural flexibility.
+Both prenatal CORT exposure and cold incubation temperatures have been shown to impair cognitive abilities in other species [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @costantini2011meta].
+Given that complex cognitive processes like behavioral flexibility can be involved in adjusting antipredatory behavior to changing risk levels [@amo2005flexibility], we expected our prenatal conditions to influence antipredatory flexibility.
+Instead, our study aligns with previous research showing that reversal learning - a standardized proxy of behavioral flexibility - was also unaffected by prenatal CORT and incubation temperature in both species [@recio2025early].
+Our findings suggest that some cognitive process may be robust to prenatal CORT exposure and incubation temperature in these species.
+However, the effects of prenatal GCs and thermal environment on brain development are region-dependent, with some brain nuclei being more impacted than others [@coomber_independent_1997; @lemaire_prenatal_2000].
+Hence, different cognitive processes, encoded in different brain regions, may be more sensitive to the effects of early conditions.
+Future research should explore the combined effects of prenatal CORT and temperature on brain function and other cognitive domains.
 
-#### _L. delicata_, but not _L. guichenoti_, shows habituation to a processive stressor    
-We found that _L.
-delicata_ increased the probability of emergence from the shelter, which suggest that risk perception decreased over time.
-In contrast, _L.
-guichenoti_ showed no sign of habituation.
-Although the slopes did not significantly differ between species for most of the treatments (see @table_contrasts_species), contrasts show that change over time was overall larger in _L.
-delicata_.
-Futhermore, only in _L.
-delicata_ slopes were different from zero, suggesting that there are differences between species in the habituation to a processive stressor.
-Differences in habituation may contribute to _L.
-delicata_'s success as an invasive species, as greater flexiblility in responding to processive stressors could facilitate adaptation to new environments [@chapple_biology_2014; @chapple2015deliinvLHI].
-For instance, learning to reassess risk based on previous exposure can reduce the costs of the antipredatory response [@lima1990behavioral].
-Moreover, cognitive habituation may reflect and individual's capacity to cope with the physiological toll of chronic stress [@grissom2009habituation], a potentially advantageous trait when colonising new habitats [@wingfield2013comparative; @surkova2024flexible].
+#### Antipredatory response and flexibility do not differ between species
+Both _L.
+delicata_ and _L.
+guichenoti_ have similar life-histories and ecological niches [@chapple_know_2011; @chapple_biology_2014].
+However, while _L.
+delicata_ is a successful invader in several parts of the world, _L.
+guichenoti_ has not established populations outside its native range [@chapple2015deliinvLHI].
+These differences in invasion success are likely related to differences in behavioral traits, with _L.
+delicata_ being more exploratory than _L.
+guichenoti_ [@chapple_know_2011].
+Given these differences, we expected _L.
+delicata_ to show lower antipredatory responses and higher antipredatory flexibility than _L.
+guichenoti_.
+However, we did not find any indication of species differences in their antipredatory strategies.
+Although _L.
+delicata_ showed some degree of antipredatory flexibility - reflected by a positive slope in the probability of emerging from shelter - the between-species differences were not significant.
+Our findings suggest that antipredatory strategies do not seem to be a good predictor of the differences in invasion success between these species.
 
-Nevertheless, the relative contribution of cognitive versus physiological mechanisms to habituate remains unclear.
-Previous studies found no differences in associative learning or behavioural flexibility between _L.
+In addition, both species responded similarly to early-life conditions, with neither prenatal CORT or incubation temperature affecting their antipredatory response or flexibility.
+Some studies show that invasive species are more plastic and responsive to environmental conditions than non-invasive species [@cordeschi2022plastic; @iglesias2024maternal].
+However, we did not find any differences in plasticity between _L.
 delicata_ and _L.
-guichenoti_ [@bezzina2014does; @recio2025cognitive; @recio2025early], suggesting that cognitive processes alone may not explain the between-species differences observed here.
-Likewise, baseline CORT levels do not fully account for the patterns either: hot-incubated _L.
-delicata_, which exhibit lower baseline CORT [@crino2024eggs], did not habituate faster than cold-incubated lizards.
-Instead, our results point to a complex interaction between cognitive and physiological habituation.
-Future studies that measure behavioural traits and CORT dynamics concurrently should be performed to disentangle the cognitive and physiological components of habituation.
-
-#### Physiological costs of the stressor were low and did not differ between treatments or species      
-Given the metabolic toll that enhanced GC levels impose on animals [@sapolsky_how_2000; @picard_mitochondria_2014], we predicted that repeated exposure to a stressor would influence body mass.
-Furthermore, we expected those costs to be related to alterarions in stress reactivity shaped by prenatal conditions.
-In contrast, we found treatment effects in mass change, and most of the lizards gained rather than lost mass.
-Notably, mass change did not differ between species despite differences in their flexibility to respond to a chronic stressor - a trait oftenly associated with more effective coping [@grissom2009habituation].
-These findings suggest that repeated exposure to a processive stressor imposes minimal metabolic costs on _L.
-delicata_ and _L.
-guichenoti_.
-
-Nevertheless, we cannot exclude the possibility that other, subtler physiological costs occurred but were not captured by overall changes in mass.
-Other studies report impairments in immune suppression, impaired growth, or permanent neural alterations following repeated exposure to systemic and processive stressors  [@picard_mitochondria_2014; @mcewen2017neurobiological].
-Future research should investigate how chronic stress impacts other physiological traits, and whether developmental conditions modulate the associated costs.
+guichenoti_, at least on their antipredatory strategies [see also @recio2025cognitive; @recio2025early].
+Future research should explore other behavioral traits that could explain the differences in invasion success between these species and how this traits can be shaped by developmental conditions.
 
 #### Conclusion
-Our study demonstrates that incubation temperature and prenatal CORT can shape behavioural responses to stress in a species- and context-dependent manner.
-_Lampropholis delicata_ showed greater plasticity in response to both acute and repeated exposure to a processive stressor, while _L.
-guichenoti_ appeared robust to prenatal conditions.
-These species differences may reflect divergent developmental strategies, with potential implications for ecological resilience and invasion success.
-Despite measurable behavioural differences, we found no evidence of metabolic costs - measured as changes in mass - associated with stress exposure under our experimental conditions.
-Together, our findings highlight the value of behavioural measures as accessible indicators of stress response and developmental plasticity.
-Future research integrating hormonal and neural data will be key to understanding the mechanisms underlying behavioural flexibility and its fitness consequences across environments.
+Our study shows that early-life conditions did not influence how skinks respond to a simulated attack.
+The antipredatory behavior and antipredatory flexibility in both _Lampropholis delicata_ and _L.
+guichenoti_ were not affected by prenatal CORT and incubation temperature.
+We also found that only _Lampropholis delicata_ showed some flexibility in their antipredatory strategies, reflected by the increased the probability of emerging from the shelter over time.
+However, when both species were compared, the differences were not significant.
+Together, these findings suggest that antipredatory response and antipredatory flexibility in both species are robust to incubation temperature and prenatal GC exposure, and that both species follow similar antipredatory strategies.
+Future studies should evaluate whether more complex developmental environments can shape antipredatory phenotypes.
+Besides, it is fundamental to integrate behavioral assays with neurological measures to understand the mechanisms underlying antipredatory response and flexibility.
+Finally, including other reptile species in this type of studies will help determine whether the robustness of antipredatory behavior to prenatal conditions is unique to _Lampropholis_ or represents a broader pattern.
 
@@
 _Table S1.
-Contrasts of the posterior distributions of the predictors for each behavioural variable modelled for L.
+Contrasts of the posterior distributions of the predictors for each behavioral variable modelled for L.
 delicata.
-The contrasts are between the intercepts and slopes of the different treatments.
+The contrasts are between the intercepts (Response) and slopes (Flexibility) of the different treatments.
 95% Highest Posterior Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
@@
 # Make Table
-table_deli_behaviour <- flextable(contrast_deli_beh) %>%
+table_deli_behavior <- flextable(contrast_deli_beh) %>%
   flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
@@
   flextable::compose(i = 1, j = c(4,6), value = as_paragraph("95% HPDI"), part = "header") %>% # Change the fourth column header
-  flextable::add_header_row(values = c("", "Acute (Intercept)", "Chronic (Slope)"), colwidths = c(2, 2, 2)) %>% # Add a header row
+  flextable::add_header_row(values = c("", "Response", "Flexibility"), colwidths = c(2, 2, 2)) %>% # Add a header row
   flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
@@
 #
-table_deli_behaviour
+table_deli_behavior
 ```
@@
 _Table S2.
-Contrasts of the posterior distributions of the predictors for each behavioural variable modelled for L.
+Contrasts of the posterior distributions of the predictors for each behavioral variable modelled for L.
 guichenoti.
-The contrasts are between the intercepts and slopes of the different treatments.
-95% Highest Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
+The contrasts are between the intercepts (Response) and slopes (Flexibility) of the different treatments.
+95% Highest Posterior Density Intervals (95% HPDI) test the hypothesis that the contrasts are different from zero._
 
@@
 # Make Table
-table_guich_behaviour <- flextable(contrast_guich_beh) %>%
+table_guich_behavior <- flextable(contrast_guich_beh) %>%
   flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
@@
 #
-table_guich_behaviour
+table_guich_behavior
 ```
@@
 _Table S3.
-Values of the slopes for the different treatments for each behavioural variable modelled for both species.
+Values of the slopes (indicator of Flexibility) for the different treatments for each behavioral variable modelled for both species.
+Median shows the median values of the posterior distribution.
 95% Highest Density Intervals (95% HPDI) test the hypothesis that the slopes are different from zero._
@@
 
-_Table S4. Estimates of each predictor for each variable modelled for L. delicata. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
+_Table S4.
+Estimates of each predictor for each variable modelled for L.
+delicata.
+Median indicates the median of the posterior distribution for each estimate.
+95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
   flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
+  flextable::compose(i = 1, j = 4, value = as_paragraph("95% HPDI"), part = "header") %>% # Change the last column header
   align(align = "center", j = c(3,4), part = "body") %>%
@@
 
-_Table S5. Estimates of each predictor for each variable modelled for L. guichenoti. 95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
+_Table S5.
+Estimates of each predictor for each variable modelled for L.
+guichenoti.
+Median indicates the median of the posterior distribution for each estimate.
+95% Highest Density Intervals (95% HPDI) test the hypothesis that the estimates are different from zero._
 
@@
 ```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```

### 2025-10-19 R/func.r

@@ -297,7 +297,7 @@ plotting <- function(sp, res, type){
 
     # Labels for the y-axis
   if(type == "slop"){
-    label <- "Estimated slopes"
+    label <- "Flexibility (slope)"
   } else {   
     if(res == "move"){
       label <- "Latency to move (s)"
